### 2015-09-27 MANUSCRIPT/manuscript.tex
Add the first manuscript draft

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

%\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

%\usepackage{graphicx}

%\draft % marks overfull lines with a black rule on the right
%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{Quantitative quality of lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@aalto.fi}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{Aalto University}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
% insert abstract here
The quantitative quality of lipid-cholesterol interactions in atomistic resolution models will be determined against 
NMR and scattering data.
\end{abstract}

%\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands

%\label{}
\section{Introduction}
The quantitative quality of lipid-cholesterol interactions in atomistic resolution models will be determined against 
NMR and scattering data.

\section{comparison of acyl chain order parameters between experiments and simulations}
Order parameters from simulations and experiments for acyl chains of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC)
are shown in Fig. \ref{OrderParametersCHOL}.

\todo{Why are the order parameters of CHARMM36 too large compared to simulations even without cholesterol?}

\todo{Why there is decrease in order parameters towards beginning of acyl chain in MacRog model without cholesterol?}

\todo{Do the results suggests that condensation effect is too strong in MacRog model?}

 \begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}

  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}

\end{figure*}

\section{comparison of structure factors between experiments and simulations}

\todo{Structure factors should be calculated from the data with cholesterol
from CHARMM and MacRog simulations (the trajectories are in Zenodo). The calculations from Berger model should be checked
(data in https://github.com/NMRLipids/NmrLipidsCholXray/tree/master/scratch/POPCberger)}

\todo{The experimental data delivered by Georg Pabst and Peter Heftberger should be plotted
together with the simulation results. The data is already in https://github.com/NMRLipids/NmrLipidsCholXray/tree/master/DATA/POPC-Cholesterol}

\section{Conclusions}

% Tables may be be put in the text as floats.
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
%
% \begin{table}
% \caption{\label{} }
% \begin{tabular}{}
% \end{tabular}
% \end{table}

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
\end{acknowledgments}

% Create the reference section using BibTe
\bibliography{refs.bib}

%\newpage
%\section{APPENDIX: The NMR results reported by Tiago Ferreira}

\listoftodos

\end{document}
%
% ****** End of file aiptemplate.tex ******

### 2016-07-01 MANUSCRIPT/manuscript.tex
Add preliminary results

@@
 
+\section{Methods}
+
+\subsection{X-ray scattering experiments}
+
+This folder contains SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations.
+Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27°C.
+Data were analyzed in terms of the SDP-GAP model described in Heftberger et al., J.
+Appl.
+Cryst.
+2013 and Heftberger et al.
+Biophys.
+J.
+2015.
+Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor.
+By fitting the scattered intensity data we obtain both contributions.
+Here we posted only form factors (ASCII format).
+For information on the quality of the fit we also give plots of the fitted intensity data.
+The electron density profile has been modelled in terms of the SDP model (see papers by Kucerka and coworkers), that is volume distribution functions are modelled by individual Gaussians or error functions.
+Cholesterol is also accounted for by two Gaussians.
+This model has been proposed by Jianjun Pan (USF, Tampa, FL), but is to the best of our knowledge not published (see also PhD Thesis by Peter Heftberger).
+Additional figures show the volume distribution functions and the resulting electron density profiles.
+
+Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
+\subsection{Molecular dynamics simulations}
+\begin{table*}[]
+\centering
+\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
+$^a$ The number of lipid molecules
+$^b$ The number of cholesterol molecules
+$^c$ Cholesterol concentration (mol\%)
+$^d$ The number of water molecules
+$^e$ Simulation temperature
+$^f$ The total simulation time
+$^g$ Time frames used in the analysis
+$^h$ Reference link for the downloadable simulation files
+$^i$ Reference for the full simulation details
+}\label{systemsCHOL}
+\begin{tabular}{c c c c c c c c c c c}
+%\hline
+Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
+\hline
+Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
+/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
+                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
+                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
+                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
+                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
+                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
+CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  &  [\citenum{botan15}]  \\
+%                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+%                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+%                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & [\citenum{botan15}] \\
+%                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+%                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+%                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+%                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & [\citenum{botan15}] \\
+MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}] \\ 
+                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+\end{tabular}
+\end{table*}
+
 \section{comparison of acyl chain order parameters between experiments and simulations}
@@
 are shown in Fig. \ref{OrderParametersCHOL}.
-
-\todo{Why are the order parameters of CHARMM36 too large compared to simulations even without cholesterol?}
-
-\todo{Why there is decrease in order parameters towards beginning of acyl chain in MacRog model without cholesterol?}
-
-\todo{Do the results suggests that condensation effect is too strong in MacRog model?}
-
  \begin{figure*}[]
@@
 
+  \todo{Why are the order parameters of CHARMM36 too large compared to simulations even without cholesterol? 
+    Discussion in https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}
+
+  \todo{Why there is decrease in order parameters towards beginning of acyl chain in MacRog model without cholesterol?}
+
+  \todo{Do the results suggests that condensation effect is too strong in MacRog and Berger models?
+  Discussion in https://github.com/NMRLipids/NmrLipidsCholXray/issues/5}
 \end{figure*}
 
-\section{comparison of structure factors between experiments and simulations}
+The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
+(currently only sn-1).
 
-\todo{Structure factors should be calculated from the data with cholesterol
-from CHARMM and MacRog simulations (the trajectories are in Zenodo). The calculations from Berger model should be checked
-(data in https://github.com/NMRLipids/NmrLipidsCholXray/tree/master/scratch/POPCberger)}
+ \begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}
 
-\todo{The experimental data delivered by Georg Pabst and Peter Heftberger should be plotted
-together with the simulation results. The data is already in https://github.com/NMRLipids/NmrLipidsCholXray/tree/master/DATA/POPC-Cholesterol}
+  \caption{\label{OrderParametersCHOLchanges}
+    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}
+
+\end{figure}
+
+\section{Comparison of form factors between experiments and simulations}
+The form factors calculated from different simulations with different cholesterol content are shown in 
+Fig.~\ref{FormFactors}.
+
+ \begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
+
+  \caption{\label{FormFactors}
+    Form factors from simulations and experiments.
+  }
+
+  \todo{Form factor calculation method should be double checked.}
+
+  \todo{Experimental form factor amplitudes are not scaled to match with simulations, as done usually}
+
+  \todo{Not all experimental and simulation data is here.}
+
+\end{figure}
 
@@
 \end{acknowledgments}
-
+\newpage
 % Create the reference section using BibTe

### 2016-08-26 scripts/makeOPchangesPLOTdata.sh
Add Slipid results from data reported by Matti Javanainen (http://dx.doi.org/10.5281/zenodo.60607) in http://nmrlipids.blogspot.com/2016/07/nmrlipids-iii-preliminary-observations.html?showComment=1471869325195#c6926781927722407348

@@ -10,17 +10,25 @@
 #psc -r < ../DATA/OPcholBERGERsn-1_1.dat | sc -W% - > ../DATA/OPcholBERGERsn-1_1transpose.dat
 
 #MACROG
-POPCsn1DATA=../scratch/POPCmacrog/OrderParamSN1.dat
-POPCchol11percSN1data=../scratch/POPCmacrog/OrderParamSN1chol10.dat
-POPCchol44percSN1data=../scratch/POPCmacrog/OrderParamSN1chol40.dat
-POPCchol50percSN1data=../scratch/POPCmacrog/OrderParamSN1chol50.dat
-POPCchol56percSN1data=../scratch/POPCmacrog/OrderParamSN1chol60.dat
-paste $POPCsn1DATA $POPCchol11percSN1data $POPCchol44percSN1data $POPCchol50percSN1data $POPCchol56percSN1data | awk 'BEGIN{print "0   11   44    50    56"}{print $2"   "$5"   "$8"    "$11"   "$14}' > ../DATA/OPcholMACROGsn-1_1.dat
-psc -r < ../DATA/OPcholMACROGsn-1_1.dat | sc -W% - > ../DATA/OPcholMACROGsn-1_1transpose.dat
+#POPCsn1DATA=../scratch/POPCmacrog/OrderParamSN1.dat
+#POPCchol11percSN1data=../scratch/POPCmacrog/OrderParamSN1chol10.dat
+#POPCchol44percSN1data=../scratch/POPCmacrog/OrderParamSN1chol40.dat
+#POPCchol50percSN1data=../scratch/POPCmacrog/OrderParamSN1chol50.dat
+#POPCchol56percSN1data=../scratch/POPCmacrog/OrderParamSN1chol60.dat
+#paste $POPCsn1DATA $POPCchol11percSN1data $POPCchol44percSN1data $POPCchol50percSN1data $POPCchol56percSN1data | awk 'BEGIN{print "0   11   44    50    56"}{print $2"   "$5"   "$8"    "$11"   "$14}' > ../DATA/OPcholMACROGsn-1_1.dat
+#psc -r < ../DATA/OPcholMACROGsn-1_1.dat | sc -W% - > ../DATA/OPcholMACROGsn-1_1transpose.dat
 
 #CHARMM36
-POPCsn1DATA=../scratch/POPCcharmm/OrderParamSN1.dat
[…]

### 2016-10-04 MANUSCRIPT/manuscript.tex
Add authors
Updating from the original NMRlipids repo (#3)
* Add figure based on results at https://github.com/NMRLipids/NmrLipidsCholXray/issues/4

* Add results by Tom Piggot from https://github.com/NMRLipids/NmrLipidsCholXray/issues/4

@@
 
+\author{Peter Heftberger}
+%\email[]{samuli.ollila@aalto.fi}
+%\homepage[]{Your web page}
+%\thanks{}
+%\altaffiliation{}
+\affiliation{University of Graz}
+
 \author{O. H. Samuli Ollila}
@@
 \affiliation{Aalto University}
+
+\author{Georg Pabst}
+\affiliation{University of Graz}
 
@@
 \newpage
+\appendix
+\begin{center}
+{\bf SUPPLEMENTARY INFORMATION}
+\end{center}
+\section{CHARMM36 results from different simulation packages}
+The results from CHARMM36 model for lipid bilayers from different 
+simulation packages have been reported to give different results in
+the literature \cite{piggot12,lee16}. The results are mainly
+dependent on different Lennart-Jones cut--off settings, but
+all the details are not quite understood. In this work we use
+the results from Gromacs 5 with settings suggested to be optimal
+by Gromacs webpage. We also compared the results from Gromacs 5 with
+these settings to the results simulated with NAMD, OpenMM and literature
+values. The comparison is shown in Fig. \ref{programsCOMP}
+ \begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}
+
+  \caption{\label{programsCOMP}
+    Results for CHARMM36 model \cite{klauda10} from different simulation packages.
+    Discussion going on at {\it https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}.
+  }
+\end{figure}
+
 % Create the reference section using BibTe

### 2017-02-01 MANUSCRIPT/manuscript.tex
Update form factor figure
Update simulation system information in the table
Update the table
Add rough draft of intro

@@
 \section{Introduction}
-The quantitative quality of lipid-cholesterol interactions in atomistic resolution models will be determined against 
-NMR and scattering data.
+Importance of phospholipid cholesterol interactions in cellular membranes
+have been reckognized for long time \cite{??}. Various ideas about 
+lateral heterogeneities driven interactions between cholesterol, phosphatidylcholine
+other other membrane lipids has been proposed \cite{??}. 
+Especially coexistence of liquid-ordered and liquid-disordered phases
+in PC-cholesterol mixtures has been suggested to be important in
+biological lipid membranes \cite{??}.
+
+Phase diagram of PC lipid cholesterol systems, including lo-ld coexistence
+can be explained by simple models where cholesterol is assumed to 
+induce order in neighboring lipids \cite{??}. This is important
+achievement in order to understand biomembrane organization. 
+However, also more detailed interactions driving interactions 
+between cholesterol and other membrane lipid molecules are
+often relevant. Several attempts to understand these interactions
+have been made by using classical molecular dynamics simulations \cite{??}.
+These simulations give atomistic resolution information, which not
+directly accessible with any other techinique. However, 
+the simulations often suffer from inaccuracies in force field
+descriptions. Acyl chain region is usually well described in 
+molecular dynamics simulation models, which is not the case for 
+glycerol backbone and headgroup \cite{??}. 
+The same applies to the systems with cholesterol. Practically all simulation
+models qualitatively reproduce the condesing due to cholesterol, observed
+as increase in acyl chain order parameters and decrease of area per molecule.
+However, the some models significantly overestimate the changes in headgroup
+due to the addition of cholesterol \cite{??}.
+
+On the other hand, detailed analysis of lateral organization of 
+lipid-cholesterol mixtures reveal significant differences between
+different models. For example, significant angular organization
+of lipids around cholesterol was observed in simulations with Berger
+Höjtje models, while CHARMM36 simulations did not show such behaviour.
 
@@
                                &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
-CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  &  [\citenum{botan15}]  \\
-%                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & [\citenum{botan15}] \\
-%                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & [\citenum{botan15}] \\
+CHARMM36\cite{klauda10,lim12}  & POPC   & 200& 0& 0\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 22& 10\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 50& 20\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 86& 30\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 134& 40\% & 15030  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
+                               & POPC   & 200& 200& 50\% & 18000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
+Slipids\cite{jambeck12b,jambeck12,jambeck13b}  & POPC   & 200 & 0 & 0\%   & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 22 & 10\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 50 & 20\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 86 & 30\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & ? & ? & 30\% & ?  & 298 & 170 & 100 & [\citenum{slipidsCHOL30T298}]$^*$ & SI  \\
+                                               & POPC   & 200 & 134 & 40\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 200 & 50\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
+                                               & POPC   & ?& ? & 50\% & ?  & 298 & 170 & 100 & [\citenum{slipidsCHOL50T298}]$^*$ & SI  \\ 
      MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}] \\ 
@@
 
+\newpage
 \section{Comparison of form factors between experiments and simulations}

### 2017-03-19 MANUSCRIPT/manuscript.tex
Update manuscript

--- 2016-10-04 MANUSCRIPT/manuscript.tex
@@
 \section{Introduction}
-The quantitative quality of lipid-cholesterol interactions in atomistic resolution models will be determined against 
-NMR and scattering data.
+Details of intermolecular interactions determine the phase
+behaviour and details lateral organization of lipid and bilayers
+containing cholesterol \cite{??}.
+Formation lateral heterogeneities \cite{??}, lipid rafts \cite{??}
+and superlattices \cite{??} in cellular membranes have been
+suggested to be driven by interactions between cholesterol
+and lipids. While detailed experimental information of these interactions
+is relatively sparse, atomistic resolution molecular dynamics simulations
+have been widely applied to give detailed explanations of lipid bilayer
+lateral organization and lipid-cholesterol interactions \cite{??}.
+
+Simulations qualitatively reproduce the cholesterol condensation effect,
+but quantitatve comparison to robust experimental data has not been
+typically done. This is partly due to the lack of availability of
+systematic experimental data sets for the comparison and partly
+due to the lack of protocols for validating intermolecular interactions
+in lipid bilayer simulations against experiments.
+Scattering experiments are more difficult to interpret for
+mixed lipid bilayers than for single component bilayers \cite{??},
+thus area per molecule values, one of the main quantity used to
+compare simulations to experiments, have not been
+available for lipid cholesterol mixtures. Systematic experimental
+data set for lipid C-H bond order parameters with different cholesterol
+concentrations in POPC bilayer has been published only relatively recently \cite{ferreira13}.
+
+In this work we present also the experimental scattering form factor
+data for POPC-cholesterol mixtures by systematically increasing the
+cholesterol concentration. Our goal is to show that the combination
+of systematically measured C-H bond order parameter and scattering form
+factor data can be used to validate the quality of lipid-cholesterol
+intermolecular interactions MD simulations. MD simulations can be also
+potentially used to give structural interpretation for form factor
+measured from mixed systems, which is a major challenge for current methods \cite{??}. 
+The approach should be also applicable for mixed lipid bilayers with
+other than lipid-cholesterol mixtures.
+
+%Importance of phospholipid cholesterol interactions in cellular membranes
+%have been reckognized for long time \cite{??}. Various ideas about 
+%lateral heterogeneities driven interactions between cholesterol, phosphatidylcholine
+%other other membrane lipids has been proposed \cite{??}. 
+%Especially coexistence of liquid-ordered and liquid-disordered phases
+%in PC-cholesterol mixtures has been suggested to be important in
+%biological lipid membranes \cite{??}.
+
+%Phase diagram of PC lipid cholesterol systems, including lo-ld coexistence
+%can be explained by simple models where cholesterol is assumed to 
+%induce order in neighboring lipids \cite{??}. This is important
+%achievement in order to understand biomembrane organization. 
+%However, also more detailed interactions driving interactions 
+%between cholesterol and other membrane lipid molecules are
+%often relevant. Several attempts to understand these interactions
+%have been made by using classical molecular dynamics simulations \cite{??}.
+%These simulations give atomistic resolution information, which not
+%directly accessible with any other techinique. However, 
+%the simulations often suffer from inaccuracies in force field
+%descriptions. Acyl chain region is usually well described in 
+%molecular dynamics simulation models, which is not the case for 
+%glycerol backbone and headgroup \cite{??}. 
+%The same applies to the systems with cholesterol. Practically all simulation
+%models qualitatively reproduce the condesing due to cholesterol, observed
+%as increase in acyl chain order parameters and decrease of area per molecule.
+%However, the some models significantly overestimate the changes in headgroup
+%due to the addition of cholesterol \cite{??}.
+
+%On the other hand, detailed analysis of lateral organization of 
+%lipid-cholesterol mixtures reveal significant differences between
+%different models. For example, significant angular organization
+%of lipids around cholesterol was observed in simulations with Berger
+%Höjtje models, while CHARMM36 simulations did not show such behaviour.
 
@@
 
-This folder contains SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations.
+SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations
+has been measured.
 Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27°C.
@@
                                &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
-CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  &  [\citenum{botan15}]  \\
-%                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & [\citenum{botan15}] \\
-%                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-%                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & [\citenum{botan15}] \\
+CHARMM36\cite{klauda10,lim12}  & POPC   & 200& 0& 0\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 22& 10\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 50& 20\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 86& 30\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 134& 40\% & 15030  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
+                               & POPC   & 200& 200& 50\% & 18000  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
+Slipids\cite{jambeck12b,jambeck12,jambeck13b}  & POPC   & 200 & 0 & 0\%   & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 512 & 0 & 0\%   & 23943  & 310 & 170 & 100 & [\citenum{slipidsCHOL0T298}]$^*$ & SI  \\
+                                               & POPC   & 200 & 22 & 10\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 50 & 20\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 86 & 30\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 358 & 154 & 30\% & 21183  & 298 & 170 & 100 & [\citenum{slipidsCHOL30T298}]$^*$ & SI  \\
+                                               & POPC   & 200 & 134 & 40\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 200 & 50\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
+                                               & POPC   & 256 & 256 & 50\% & 20334  & 298 & 170 & 100 & [\citenum{slipidsCHOL50T298}]$^*$ & SI  \\ 
      MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}] \\ 

### 2017-03-29 scripts/areaPERmolecules.sh
Add area per molecule results

@@ -1,0 +1,39 @@
+rm ../DATA/aplCHARMM36_T310K.dat
+echo '#[chol]  apl' >> ../DATA/aplCHARMM36_T310K.dat
+awk '{print 0" "2*$1/200" "2*$1/200}' ../../MATCH/Data/Lipid_Bilayers/POPC/T310K/MODEL_CHARMM36_GROMACS5.0/area.dat >> ../DATA/aplCHARMM36_T310K.dat
+awk '{print 10" "2*$1/220" "2*$1/200}' ../../MATCH/Data/Lipid_Bilayers/POPC+10%chol/T310K/MODEL_CHARMM36_GROMACS5.0/area.dat >> ../DATA/aplCHARMM36_T310K.dat
+awk '{print 20" "2*$1/250" "2*$1/200}' ../../MATCH/Data/Lipid_Bilayers/POPC+20%chol/T310K/MODEL_CHARMM36_GROMACS5.0/area.dat >> ../DATA/aplCHARMM36_T310K.dat
+awk '{print 30" "2*$1/286" "2*$1/200}' ../../MATCH/Data/Lipid_Bilayers/POPC+30%chol/T310K/MODEL_CHARMM36_GROMACS5.0/area.dat >> ../DATA/aplCHARMM36_T310K.dat
+awk '{print 40" "2*$1/334" "2*$1/200}' ../../MATCH/Data/Lipid_Bilayers/POPC+40%chol/T310K/MODEL_CHARMM36_GROMACS5.0/area.dat >> ../DATA/aplCHARMM36_T310K.dat
+awk '{print 50" "2*$1/400" "2*$1/200}' ../../MATCH/Data/Lipid_Bilayers/POPC+50%chol/T310K/MODEL_CHARMM36_GROMACS5.0/area.dat >> ../DATA/aplCHARMM36_T310K.dat
+
+rm ../DATA/aplSLIPID_T298K.dat
+echo '#[chol]  apl' >> ../DATA/aplSLIPID_T298K.dat
+awk '{print 0" "2*$1/(512)" "2*$1/512}' /home/samuli/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T298K/MODEL_SLIPIDS/area.dat >> ../DATA/aplSLIPID_T298K.dat
[…]

### 2017-03-29 MANUSCRIPT/manuscript.tex
Update manuscript tex

@@
 
+\section{Results and Discussion}
+
+\subsection{Structural sampling of molecules in bilayers with cholesterol}
+Order parameters for C-H bond vectors in lipid bilayer systems, measured
+with 13C or 2H NMR techniques, give indirect information about structural
+sampling of indidual molecules \cite{ollila16}. The observed increase of
+acyl chain order parameters with added cholesterol in simulations and
+experiments can be explained by increased trans conformations in acyl chains \cite{??},
+which is suggested to play critical role in phase behaviour in PC lipid-cholesterol
+mixtures \cite{??}. Consequently, the correct cholesterol ordering effect is
+expected to be necessary condition for a model used to understand lipid-cholesterol
+phase behaviour.
+
+Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
+and mixture with 50 mol\% of cholesterol from different simulations and experiments
+are shown in Fig. \ref{OrderParametersCHOL}. Phase separation is not observed for this system \cite{ferreira13,??},
+thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
+Experimental acyl chain order parameters are typically well reproduced in state of the art lipid models
+for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
+presented here, except that CHARMM36 slightly overestimates the order parameters and MACROG model
+underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
+all models overestimate the order parameters. For Berger/Holtje and MACROG models the overestimation is significant,
+while less severe for CHARMM36 and Slipid models.\\
+\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
+\todo{Lipid14 results from recent article should be added in this figure} 
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
+  \caption{\label{OrderParametersCHOL}
+    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+  }
+\end{figure*}
+
+The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
+(currently only sn-1). \\
+\todo{This figure should be finished and similar for sn-2 should be done as well}
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}
+  \caption{\label{OrderParametersCHOLchanges}
+    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}
+
+\end{figure}
+
+\todo{Also experimental cholesterol order parameters are available. Maybe these should be calculated from simulations as well.}
+
+\subsection{Lipid bilayer dimensions and density profiles as a function of cholesterol}
+Cholesterol induced changes in lipid packing, thickness and density profiles are suggested
+to be relevant, for example, for membrane protein interactions and cellular membrane
+lateral organization \cite{??}. Experimental and simulation studies show that
+the cholesterol induces the so called ''condensing effect'', i.e. acyl chain ordering
+increases the membrane thickness and
+reduces the area per molecule in lipid bilayers \cite{??}. This is demonstrated
+from simulations in Fig. \ref{apls}, showing the area per PC headgroups and per
+total number of molecules as a function of cholesterol concentration from different
+models. All models show reduction of area per total amount of molecules, partly
+due to the smaller area covered by the cholesterol than lipids, but also because
+lipids ordered by cholesterol require less space. Due to the latter effect,
+the area per PC headgroup do not essentially increase up to the addition of
+$\sim$15 mol\% of cholesterol. \\
+\todo{Membrane thickness to be plotted as well.}
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/apls.eps}
+  \caption{\label{apls}
+    Area per molecules calculated from different simulation models
+    as a function of cholesterol concentration.
+  }
+\end{figure}
+
+Bilayer dimensions are indirectly experimentally accessible by scattering methods 
+and have been widely used to validate lipid bilayer simulations \cite{??}. 
+While sophisticated models are applied to extract area per molecule and thicknes for single
+component lipid bilayers, the results for multicomponent systems are more difficult to interpret \cite{??}.
+In such case the validation can be done by comparing directly measurable form factors between
+simulations and experiments \cite{??}. If simulation reproduces the form factor, it can be also
+used to interpret the experiments.  The form factors from different simulation models with
+different cholesterol concentrations are compared to experimental data Fig.~\ref{FormFactors}. \\
+\todo{Discussion to be finished when the data is solid.}
+
+ \begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
+  \caption{\label{FormFactors}
+    Form factors from simulations and experiments.
+  }
+
+  \todo{Form factor calculation method should be double checked.}
+
+  \todo{Experimental form factor amplitudes are not scaled to match with simulations, as done usually}
+
+  \todo{Not all experimental and simulation data is here.}
+
+\end{figure}
+
+\section{Conclusions}
+
 \section{Methods}
@@
 Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
+
 \subsection{Molecular dynamics simulations}
+
+Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
+details are given in references or in supplementary material.
+
 \begin{table*}[]
@@
 \end{table*}
-
-\section{comparison of acyl chain order parameters between experiments and simulations}
-Order parameters from simulations and experiments for acyl chains of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC)
-are shown in Fig. \ref{OrderParametersCHOL}.
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
-
-  \caption{\label{OrderParametersCHOL}
-    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}
-
-  \todo{Why are the order parameters of CHARMM36 too large compared to simulations even without cholesterol? 
-    Discussion in https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}
-
-  \todo{Why there is decrease in order parameters towards beginning of acyl chain in MacRog model without cholesterol?}
-
-  \todo{Do the results suggests that condensation effect is too strong in MacRog and Berger models?
-  Discussion in https://github.com/NMRLipids/NmrLipidsCholXray/issues/5}
-\end{figure*}
-
-The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
-(currently only sn-1).
-
- \begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}
-
-  \caption{\label{OrderParametersCHOLchanges}
-    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}
-
-\end{figure}
-
-\newpage
-\section{Comparison of form factors between experiments and simulations}
-The form factors calculated from different simulations with different cholesterol content are shown in 
-Fig.~\ref{FormFactors}.
-
- \begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
-
-  \caption{\label{FormFactors}
-    Form factors from simulations and experiments.
-  }
-
-  \todo{Form factor calculation method should be double checked.}
-
-  \todo{Experimental form factor amplitudes are not scaled to match with simulations, as done usually}
-
-  \todo{Not all experimental and simulation data is here.}
-
-\end{figure}
-
-\section{Conclusions}
 

### 2017-03-30 MANUSCRIPT/manuscript.tex
New version of manuscript

@@
 
-\title{Quantitative quality of lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper
+\title{NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper
 
@@
 \author{O. H. Samuli Ollila}
-\email[]{samuli.ollila@aalto.fi}
+\email[]{samuli.ollila@helsinki.fi}
 %\homepage[]{Your web page}
@@
 %\altaffiliation{}
-\affiliation{Aalto University}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
+\affiliation{Institute of Biotechnology, University of Helsinki}
 
@@
 behaviour and details lateral organization of lipid and bilayers
-containing cholesterol \cite{??}.
-Formation lateral heterogeneities \cite{??}, lipid rafts \cite{??}
-and superlattices \cite{??} in cellular membranes have been
+containing cholesterol \cite{ipsen87}.
+Formation lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
+and superlattices \cite{somerharju09} in cellular membranes have been
 suggested to be driven by interactions between cholesterol
@@
 have been widely applied to give detailed explanations of lipid bilayer
-lateral organization and lipid-cholesterol interactions \cite{??}.
+lateral organization and lipid-cholesterol interactions \cite{rog14,sodt14,??}.
+However, the simulations must reproduce the measured details, like NMR
+order parameters and scattering form factors, to be useful in intepretation
+of molecular details in lipid bilayer mixtures.
 
@@
 Scattering experiments are more difficult to interpret for
-mixed lipid bilayers than for single component bilayers \cite{??},
+mixed lipid bilayers than for single component bilayers \cite{pan12,Heftberger15,Marquardt15,??},
 thus area per molecule values, one of the main quantity used to
@@
 potentially used to give structural interpretation for form factor
-measured from mixed systems, which is a major challenge for current methods \cite{??}. 
+measured from mixed systems, which is a major challenge for current methods \cite{pan12,Heftberger15,Marquardt15,??}. 
 The approach should be also applicable for mixed lipid bilayers with
@@
 Order parameters for C-H bond vectors in lipid bilayer systems, measured
-with 13C or 2H NMR techniques, give indirect information about structural
+with $^{13}$C or $^{2}$H NMR techniques, give indirect information about structural
 sampling of indidual molecules \cite{ollila16}. The observed increase of
 acyl chain order parameters with added cholesterol in simulations and
-experiments can be explained by increased trans conformations in acyl chains \cite{??},
+experiments can be explained by increased trans conformations in acyl chains \cite{ferreira13,??},
 which is suggested to play critical role in phase behaviour in PC lipid-cholesterol
-mixtures \cite{??}. Consequently, the correct cholesterol ordering effect is
+mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is
 expected to be necessary condition for a model used to understand lipid-cholesterol
@@
 and mixture with 50 mol\% of cholesterol from different simulations and experiments
-are shown in Fig. \ref{OrderParametersCHOL}. Phase separation is not observed for this system \cite{ferreira13,??},
+are shown in Fig. \ref{OrderParametersCHOL}. Phase separation is not observed for this system \cite{ionova12,ferreira13},
 thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
@@
     Area per molecules calculated from different simulation models
-    as a function of cholesterol concentration.
+    as a function of cholesterol concentration. Solid marks are area per
+    total amount of molecules (chol+PC) and empty marks are area per PC headgroups.
+    Top figure shows absolute values and bottom figure shows changes respect
+    to pure lipid system.
   }
@@
 Bilayer dimensions are indirectly experimentally accessible by scattering methods 
-and have been widely used to validate lipid bilayer simulations \cite{??}. 
+and have been widely used to validate lipid bilayer simulations \cite{ollila16,??}. 
 While sophisticated models are applied to extract area per molecule and thicknes for single
-component lipid bilayers, the results for multicomponent systems are more difficult to interpret \cite{??}.
+component lipid bilayers, the results for multicomponent systems are more difficult to interpret \cite{pan12,Heftberger15,Marquardt15,??}.
 In such case the validation can be done by comparing directly measurable form factors between
-simulations and experiments \cite{??}. If simulation reproduces the form factor, it can be also
+simulations and experiments. If simulation reproduces the form factor, it can be also
 used to interpret the experiments.  The form factors from different simulation models with
@@
 \section{Conclusions}
+
+Cholesterol ordering effect is overestimated in Berger/Holtje and MacRog models.
+Slight overestimation is observed also in CHARMM36 and Slipid models, but more
+careful analysis is required to conclude if this is significant or not.
 
@@
 these settings to the results simulated with NAMD, OpenMM and literature
-values. The comparison is shown in Fig. \ref{programsCOMP}
+values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
+that Gromacs 5 with settings suggested in webpage gives consistent
+results with the literature and other simulation packages. Thus these
+results are used in the main body of the paper. However, order parameters
+are slightly overestimated respected to the experiments also with these
+settings. \\
+\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
  \begin{figure}[]

### 2017-03-30 DATA/changeAPLs.sh

@@ -1,0 +1,5 @@
+awk '{if(NR==1){print $0}else if(NR==2){ref1=$2; ref2=$3; print $1" "$2-ref1" "$3-ref2}else{print $1" "$2-ref1" "$3-ref2}}' aplBerger_T298K.dat > aplBerger_T298Kchange.dat
+awk '{if(NR==1){print $0}else if(NR==2){ref1=$2; ref2=$3; print $1" "$2-ref1" "$3-ref2}else{print $1" "$2-ref1" "$3-ref2}}' aplCHARMM36_T310K.dat > aplCHARMM36_T310Kchange.dat
+awk '{if(NR==1){print $0}else if(NR==2){ref1=$2; ref2=$3; print $1" "$2-ref1" "$3-ref2}else{print $1" "$2-ref1" "$3-ref2}}' aplMACROG_T310K.dat > aplMACROG_T310Kchange.dat
+awk '{if(NR==1){print $0}else if(NR==2){ref1=$2; ref2=$3; print $1" "$2-ref1" "$3-ref2}else{print $1" "$2-ref1" "$3-ref2}}' aplSLIPID_T298K.dat > aplSLIPID_T298Kchange.dat
+awk '{if(NR==1){print $0}else if(NR==2){ref1=$2; ref2=$3; print $1" "$2-ref1" "$3-ref2}else{print $1" "$2-ref1" "$3-ref2}}' aplSLIPID_T310K.dat > aplSLIPID_T310Kchange.dat

### 2017-04-12 MANUSCRIPT/manuscript.tex
Add authors
Small fixes of typos.

@@
 
+\author{Fernando Favela-Rosales}
+\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
+
 \author{Peter Heftberger}
@@
 \affiliation{University of Graz}
+
+\author{Matti Javanainen}
+\affiliation{Department of Physics, Tampere University of Technology, Tampere, Finland}
+\affiliation{University of Helsinki}
+
+\author{Josef Melcr}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
 
@@
 \affiliation{University of Graz}
+
+\author{Thomas Piggot}
+\affiliation{University of Southampton}
 
@@
 behaviour and details lateral organization of lipid and bilayers
-containing cholesterol \cite{ipsen87}.
-Formation lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
+containing cholesterol \cite{ipsen87}.~\todo{Fix Ole's name among the authors of this paper.}
+Formation of lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
 and superlattices \cite{somerharju09} in cellular membranes have been
@@
 lateral organization and lipid-cholesterol interactions \cite{rog14,sodt14,??}.
-However, the simulations must reproduce the measured details, like NMR
+However, the simulations must reproduce the measured experimental details, such as NMR
 order parameters and scattering form factors, to be useful in intepretation
@@
 Simulations qualitatively reproduce the cholesterol condensation effect,
-but quantitatve comparison to robust experimental data has not been
+but quantitative comparison to robust experimental data has not been
 typically done. This is partly due to the lack of availability of
@@
 
-In this work we present also the experimental scattering form factor
+In this work we present the experimental scattering form factor
 data for POPC-cholesterol mixtures by systematically increasing the
-cholesterol concentration. Our goal is to show that the combination
-of systematically measured C-H bond order parameter and scattering form
-factor data can be used to validate the quality of lipid-cholesterol
-intermolecular interactions MD simulations. MD simulations can be also
-potentially used to give structural interpretation for form factor
+cholesterol concentration, scanning the same concentrations that were previously studied with ssNMR~\cite{ferreira13}.
+Our goal is to show that the combination
+of systematically measured C--H bond order parameter and scattering form
+factor data can be used to validate the quality of lipid--cholesterol
+intermolecular interactions in MD simulations. MD simulations can be also
+potentially used to give structural interpretation for form factors
 measured from mixed systems, which is a major challenge for current methods \cite{pan12,Heftberger15,Marquardt15,??}. 
 The approach should be also applicable for mixed lipid bilayers with
-other than lipid-cholesterol mixtures.
+other than lipid--cholesterol mixtures.
 
@@
 Order parameters for C-H bond vectors in lipid bilayer systems, measured
-with $^{13}$C or $^{2}$H NMR techniques, give indirect information about structural
+with $^{13}$C or $^{2}$H NMR techniques, give indirect information about the structural
 sampling of indidual molecules \cite{ollila16}. The observed increase of
@@
 experiments can be explained by increased trans conformations in acyl chains \cite{ferreira13,??},
-which is suggested to play critical role in phase behaviour in PC lipid-cholesterol
+which is suggested to play critical role in the phase behaviour of PC-lipid--cholesterol
 mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is
-expected to be necessary condition for a model used to understand lipid-cholesterol
+expected to be a necessary condition for a model used to understand lipid--cholesterol
 phase behaviour.
@@
 thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
-Experimental acyl chain order parameters are typically well reproduced in state of the art lipid models
+Experimental acyl chain order parameters are typically well reproduced in the state of the art lipid models
 for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
@@
 underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
-all models overestimate the order parameters. For Berger/Holtje and MACROG models the overestimation is significant,
+all models overestimate the order parameters. For Berger/H\"oltje and MACROG models the overestimation is significant,
 while less severe for CHARMM36 and Slipid models.\\
@@
 lipids ordered by cholesterol require less space. Due to the latter effect,
-the area per PC headgroup do not essentially increase up to the addition of
+the area per PC headgroup does not essentially increase up to the addition of
 $\sim$15 mol\% of cholesterol. \\
@@
     Area per molecules calculated from different simulation models
-    as a function of cholesterol concentration. Solid marks are area per
-    total amount of molecules (chol+PC) and empty marks are area per PC headgroups.
+    as a function of cholesterol concentration. The solid symbols are area per
+    total amount of molecules (chol+PC) and the empty symbols area per PC headgroups.
     Top figure shows absolute values and bottom figure shows changes respect
@@
 used to interpret the experiments.  The form factors from different simulation models with
-different cholesterol concentrations are compared to experimental data Fig.~\ref{FormFactors}. \\
+different cholesterol concentrations are compared to experimental data in Fig.~\ref{FormFactors}. \\
 \todo{Discussion to be finished when the data is solid.}
@@
 the literature \cite{piggot12,lee16}. The results are mainly
-dependent on different Lennart-Jones cut--off settings, but
+dependent on different Lennard-Jones cut-off settings, but
 all the details are not quite understood. In this work we use
 the results from Gromacs 5 with settings suggested to be optimal
-by Gromacs webpage. We also compared the results from Gromacs 5 with
+by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
 these settings to the results simulated with NAMD, OpenMM and literature

### 2017-04-21 DATA/POPC-Cholesterol/Chol = 0 mol_/epstodat.sh
Add density results and update form factor figure

@@ -1,0 +1,8 @@
+sed 's/ /\
+/g' tst.dat > tst2.dat
+sed 's/,/ /g' tst2.dat > tst3.dat
+xaxis=$(awk '{print ($2-$1)/10}' xaxis.dat)
+xaxisPOINT=$(awk '{print $1}' xaxis.dat)
+yaxis=$(awk '{print ($2-$1)/0.1}' yaxis.dat)
+yaxisPOINT=$(awk '{print $1}' yaxis.dat)
+awk -v xaxis=$xaxis -v xaxisPOINT=$xaxisPOINT -v yaxis=$yaxis -v yaxisPOINT=$yaxisPOINT '{if(NR==1){refx=$1;refy=$2}else{refx=refx+$1; refy=refy+$2} print (refx-xaxisPOINT)/xaxis-30" "(refy-yaxisPOINT)/yaxis+0.2}' tst3.dat > tst4.dat

### 2017-04-21 DATA/POPC-Cholesterol/Chol = 50 mol_/epstodat.sh

@@ -1,0 +1,8 @@
+sed 's/ /\
+/g' densityEPS.dat > tst2.dat
+sed 's/,/ /g' tst2.dat > tst3.dat
+xaxis=$(awk '{print ($2-$1)/10}' xaxis.dat)
+xaxisPOINT=$(awk '{print $1}' xaxis.dat)
+yaxis=$(awk '{print ($2-$1)/0.1}' yaxis.dat)
+yaxisPOINT=$(awk '{print $1}' yaxis.dat)
+awk -v xaxis=$xaxis -v xaxisPOINT=$xaxisPOINT -v yaxis=$yaxis -v yaxisPOINT=$yaxisPOINT '{if(NR==1){refx=$1;refy=$2}else{refx=refx+$1; refy=refy+$2} print (refx-xaxisPOINT)/xaxis-30" "(refy-yaxisPOINT)/yaxis+0.2}' tst3.dat > tst4.dat

### 2019-05-02 MANUSCRIPT/manuscript.tex
Add author
Add density results and update form factor figure
Add fittness results

@@
 Prague 6, Czech Republic}
+
+\author{Markus Miettinen}
+\affiliation{MPI}
 
@@
 
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
+  \caption{\label{OrderParametersCHOL}
+    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+  }
+\end{figure*}
+
 The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
@@
   }
+  \todo{Form factor calculation method should be double checked.}
+  \todo{Experimental form factor amplitudes are not scaled to match with simulations, as done usually}
+  \todo{Not all experimental and simulation data is here.}
+\end{figure}
 
-  \todo{Form factor calculation method should be double checked.}
+ \begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/FFfitness.eps}
+  \caption{\label{FormFactorsFITNESS}
+    Fitness of form factors between simulations and experiments.
+  }
+ \end{figure}
 
-  \todo{Experimental form factor amplitudes are not scaled to match with simulations, as done usually}
-
-  \todo{Not all experimental and simulation data is here.}
-
+  \begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{../FIGS/densitiesEXTREMES.eps}
+  \caption{\label{densities}
+    Density profiles from simulations and experiments.
+  }
 \end{figure}

### 2019-05-02 scripts/makeFFfitnessPLOTdata.sh

@@ -5,3 +5,11 @@ grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/
 rm  ../DATA/FFfitnessCHARMM.dat
 grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T310K/MODEL_CHARMM36_GROMACS5.0/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/FFfitnessCHARMM.dat
 grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC+50%chol/T310K/MODEL_CHARMM36_GROMACS5.0/fitness.txt | awk '{print 50"   "$4}' >> ../DATA/FFfitnessCHARMM.dat
+
+rm  ../DATA/FFfitnessBERGER.dat
+grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T298K/MODEL_Berger/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/FFfitnessBERGER.dat
+grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC+50%chol/T298K/MODEL_Berger/fitness.txt | awk '{print 50"   "$4}' >> ../DATA/FFfitnessBERGER.dat
+
+rm  ../DATA/FFfitnessMACROG.dat
+grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T310K/MODEL_MACROG/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/FFfitnessMACROG.dat
+grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC+50%chol/T310K/MODEL_MACROG/fitness.txt | awk '{print 50"   "$4}' >> ../DATA/FFfitnessMACROG.dat

### 2019-05-03 scripts/calcSLOPES.py
Add order parameter slopes as a function of cholesterol

@@ -1,0 +1,59 @@
+import sys
+import numpy
+import math
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+
+def func(x, k, a):
+  return k*x+a
+CFfig=plt.figure()
+      
+
+
+
+#data1 = numpy.loadtxt(sys.argv[1], usecols=range(0,7))
+for c in range(1,15):
+  data1 = numpy.loadtxt(sys.argv[1], usecols=range(0,1)), numpy.loadtxt(sys.argv[1], usecols=range(c,c+1))
+  #data2 = numpy.loadtxt(sys.argv[1], usecols=range(0,1)), numpy.loadtxt(sys.argv[1], usecols=range(2,3))
+  #data3 = numpy.loadtxt(sys.argv[1], usecols=range(0,1)), numpy.loadtxt(sys.argv[1], usecols=range(3,4))
+  #print data1
+  
+  k1=0#[0]*3
+  k2=0#[0]*3
+  k3=0#[0]*3
+  kcov1=0#[0]*3
+  kcov2=0#[0]*3
+  kcov3=0#[0]*3
+  analL=51
+  j=0
+  for i in range(0,len(data1[0])):
+    if data1[0][i]<analL:
+      j=j+1
+      dataLOGx = [0]*j
+      dataLOGy = [0]*j
+      
+  for i in range(0,j):
+    dataLOGx[i]=data1[0][i]
+    dataLOGy[i]=data1[1][i]
+    #print dataLOGx
+  k1, kcov1 = curve_fit(func, dataLOGx, dataLOGy)
+  print c+1, k1[0],k1[1]
+    #plt.plot(dataLOGx,dataLOGy,dataLOGx,func(dataLOGx,k1))
+  if c>9:
+    color='C'+str(c-6)
+  else:
+    color='C'+str(c)
+  plt.plot(dataLOGx,dataLOGy,marker='o',linestyle='',color=str(color))
+  plt.plot(data1[0],func(data1[0],k1[0],k1[1]),color=str(color))
+
+
+CFfig.savefig('../scratch/tst.pdf')
+    
+              
+#print data1[0]
+#y=func(data1[0],k3[0],k3[1])
+#print y
+
+
+
+

### 2019-05-07 MANUSCRIPT/manuscript.tex
Update splope and fitness analysis results

@@
   \centering
-  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}
+  \includegraphics[width=8cm]{../FIGS/slopes.eps}
   \caption{\label{OrderParametersCHOLchanges}
-    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}
+    Slopes of order parameters as a function of cholesterol.
+    %Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.
+  }
 
@@
 has been measured.
-Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27°C.
+Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27$^\circ$C.
 Data were analyzed in terms of the SDP-GAP model described in Heftberger et al., J.
@@
 
+\section{Order parameter splopes as a function of cholesterol}
+ \begin{figure*}[]
+  \centering
+  \includegraphics[width=19cm]{../FIGS/slopesEXPERIMENT.pdf}
+  \caption{\label{programsCOMP}
+    Slopes of order parameters as a function of cholesterol from experiments \cite{ferreira13}.
+  }
+\end{figure*}
+
+ \begin{figure*}[]
+  \centering
+  \includegraphics[width=19cm]{../FIGS/slopesBERGER.pdf}
+  \caption{\label{programsCOMPberger}
+    Slopes of order parameters as a function of cholesterol from Berger simulations.
+  }
+\end{figure*}
+
+ \begin{figure*}[]
+  \centering
+  \includegraphics[width=19cm]{../FIGS/slopesCHARMM.pdf}
+  \caption{\label{programsCOMPcharmm}
+    Slopes of order parameters as a function of cholesterol from CHARMM36 simulations.
+  }
+\end{figure*}
+ \begin{figure*}[]
+  \centering
+  \includegraphics[width=19cm]{../FIGS/slopesMACROG.pdf}
+  \caption{\label{programsCOMPmacrog}
+    Slopes of order parameters as a function of cholesterol from MacRog simulations.
+  }
+\end{figure*}
+ \begin{figure*}[]
+  \centering
+  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDS.pdf}
+  \caption{\label{programsCOMPslipids}
+    Slopes of order parameters as a function of cholesterol from Slipids simulations.
+  }
+\end{figure*}
+ \begin{figure*}[]
+  \centering
+  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDST310K.pdf}
+  \caption{\label{programsCOMPslipids310K}
+    Slopes of order parameters as a function of cholesterol from Slipids simulations at 298 K.
+  }
+\end{figure*}
+
+% \begin{figure}[]
+%  \centering
+%  \includegraphics[width=8cm]{../FIGS/slopesBERGER.pdf}
+%  \caption{\label{programsCOMP}
+%    Slopes of order parameters as a function of cholesterol from Berger simulations.
+%  }
+%\end{figure}
+
 % Create the reference section using BibTe

### 2019-05-07 scripts/calcSLOPES.py

@@ -6,11 +6,10 @@ from scipy.optimize import curve_fit
 
 def func(x, k, a):
   return k*x+a
-CFfig=plt.figure()
-      
 
+CFfig=plt.figure(figsize=(10,10))
 
-
+analL=float(sys.argv[2])
 #data1 = numpy.loadtxt(sys.argv[1], usecols=range(0,7))
 for c in range(1,15):
   data1 = numpy.loadtxt(sys.argv[1], usecols=range(0,1)), numpy.loadtxt(sys.argv[1], usecols=range(c,c+1))
@@ -24,7 +23,6 @@ for c in range(1,15):
   kcov1=0#[0]*3
   kcov2=0#[0]*3
   kcov3=0#[0]*3
-  analL=51
   j=0
   for i in range(0,len(data1[0])):
     if data1[0][i]<analL:
@@ -43,8 +41,19 @@ for c in range(1,15):
     color='C'+str(c-6)
   else:
     color='C'+str(c)
-  plt.plot(dataLOGx,dataLOGy,marker='o',linestyle='',color=str(color))
-  plt.plot(data1[0],func(data1[0],k1[0],k1[1]),color=str(color))
+  title='C'+str(c+1)
+  plt.subplot(7,2,c)
+  plt.ylim((0.05,0.43))
+  plt.xlim((-5,65)) 
+  plt.plot(data1[0],numpy.absolute(data1[1]),marker='o',linestyle='')
+  plt.plot(data1[0],numpy.absolute(func(data1[0],k1[0],k1[1])))
+  plt.text(0,0.35,title)
+  if c<13:
+    plt.gca().axes.get_xaxis().set_ticklabels([])
+  else:
+    plt.xlabel('[cholesterol]')
+  #plt.gca().axes.get_yaxis().set_visible(False)
+  #plt.legend(bbox_to_anchor=(1, 1),loc=2, borderaxespad=0.)
 
 
 CFfig.savefig('../scratch/tst.pdf')

### 2019-05-07 scripts/calcSLOPES.sh

@@ -1,0 +1,18 @@
+python calcSLOPES.py ../DATA/EXP-Ferreira13/S_CH_POPC/OP_palmitoylTRANSPOSE.txt 51 > ../DATA/expSLOPES.dat
+mv ../scratch/tst.pdf ../FIGS/slopesEXPERIMENT.pdf
+
+
+python calcSLOPES.py ../DATA/OPcholBERGERsn-1_1transpose.dat 51 > ../DATA/bergerSLOPES.dat
+mv ../scratch/tst.pdf ../FIGS/slopesBERGER.pdf
+
+python calcSLOPES.py ../DATA/OPcholCHARMM36sn-1_1transpose.dat 41 > ../DATA/charmmSLOPES.dat
+mv ../scratch/tst.pdf ../FIGS/slopesCHARMM.pdf
+
+python calcSLOPES.py ../DATA/OPcholMACROGsn-1_1transposeFORslopes.dat 51 > ../DATA/macrogSLOPES.dat
+mv ../scratch/tst.pdf ../FIGS/slopesMACROG.pdf
+
+python calcSLOPES.py ../DATA/OPcholSLIPIDsn-1_1transpose.dat 51 > ../DATA/slipidsSLOPES.dat
+mv ../scratch/tst.pdf ../FIGS/slopesSLIPIDS.pdf
+
+python calcSLOPES.py ../DATA/OPcholSLIPIDT310Ksn-1_1transpose.dat 41 > ../DATA/slipidsT310KSLOPES.dat
+mv ../scratch/tst.pdf ../FIGS/slopesSLIPIDST310K.pdf

### 2019-05-07 scripts/makeFFfitnessPLOTdata.sh

@@ -1,10 +1,34 @@
 rm  ../DATA/FFfitnessSLIPIDS.dat
+rm  ../DATA/HGfitnessSLIPIDS.dat
+rm  ../DATA/GLYfitnessSLIPIDS.dat
+rm  ../DATA/SN1fitnessSLIPIDS.dat
+rm  ../DATA/SN2fitnessSLIPIDS.dat
 grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/FFfitnessSLIPIDS.dat
 grep 'fitness from FF' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC+50%chol/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 50"   "$4}' >> ../DATA/FFfitnessSLIPIDS.dat
+grep 'fitness from headgr ' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/HGfitnessSLIPIDS.dat
+grep 'fitness from headgr ' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC+50%chol/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 50"   "$4}' >> ../DATA/HGfitnessSLIPIDS.dat
+grep 'fitness from glys ' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/GLYfitnessSLIPIDS.dat
+grep 'fitness from glys ' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC+50%chol/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 50"   "$4}' >> ../DATA/GLYfitnessSLIPIDS.dat
+grep 'fitness from sn1 ' /home/local/osollila/NMRlipids/MATCH/Data/Lipid_Bilayers/POPC/T298K/MODEL_SLIPIDS/fitness.txt | awk '{print 0"   "$4}'  >> ../DATA/SN1fitnessSLIPIDS.dat
[…]

### 2019-05-07 scripts/makeOPchangesPLOTdata.sh

@@ -28,17 +28,19 @@
 #psc -r < ../DATA/OPcholCHARMM36sn-1_1.dat | sc -W% - > ../DATA/OPcholCHARMM36sn-1_1transpose.dat
 
 #SLIPIDS T310K
-#POPCsn1DATA=../../MATCH/Data/Lipid_Bilayers/POPC/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
-#POPCchol10percSN1data=../../MATCH/Data/Lipid_Bilayers/POPC+10%chol/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
-#POPCchol20percSN1data=../../MATCH/Data/Lipid_Bilayers/POPC+20%chol/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
-#POPCchol30percSN1data=../../MATCH/Data/Lipid_Bilayers/POPC+30%chol/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
-#paste $POPCsn1DATA $POPCchol10percSN1data $POPCchol20percSN1data $POPCchol30percSN1data | awk 'BEGIN{print "0  10  20   30"}{if(NR>1)print $2"   "$5"   "$8"   "$11}' > ../DATA/OPcholSLIPIDsn-1_1.dat
-#psc -r < ../DATA/OPcholSLIPIDsn-1_1.dat | sc -W% - > ../DATA/OPcholSLIPIDsn-1_1transpose.dat
+POPCsn1DATA=../../MATCH/Data/Lipid_Bilayers/POPC/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
+POPCchol10percSN1data=../../MATCH/Data/Lipid_Bilayers/POPC+10%chol/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
+POPCchol20percSN1data=../../MATCH/Data/Lipid_Bilayers/POPC+20%chol/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
+POPCchol30percSN1data=../../MATCH/Data/Lipid_Bilayers/POPC+30%chol/T310K/MODEL_SLIPIDS/sn-1_Order_Parameters_Simulation.dat
[…]

### 2019-05-07 MANUSCRIPT/manuscript.tex
Add separate supplementary information

--- 2019-05-02 MANUSCRIPT/manuscript.tex
@@
 \usepackage[obeyFinal]{easy-todo}
+\usepackage{xr}
+\externaldocument{manuscriptSUPPL}
 \begin{document}
@@
 
-\section{Results and Discussion}
-
-\subsection{Structural sampling of molecules in bilayers with cholesterol}
 Order parameters for C-H bond vectors in lipid bilayer systems, measured
@@
 
-Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
-and mixture with 50 mol\% of cholesterol from different simulations and experiments
-are shown in Fig. \ref{OrderParametersCHOL}. Phase separation is not observed for this system \cite{ionova12,ferreira13},
-thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
+\section{Methods}
+
+\subsection{X-ray scattering experiments}
+
+SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations
+has been measured.
+Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27$^\circ$C.
+Data were analyzed in terms of the SDP-GAP model described in Heftberger et al., J.
+Appl.
+Cryst.
+2013 and Heftberger et al.
+Biophys.
+J.
+2015.
+Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor.
+By fitting the scattered intensity data we obtain both contributions.
+Here we posted only form factors (ASCII format).
+For information on the quality of the fit we also give plots of the fitted intensity data.
+The electron density profile has been modelled in terms of the SDP model (see papers by Kucerka and coworkers), that is volume distribution functions are modelled by individual Gaussians or error functions.
+Cholesterol is also accounted for by two Gaussians.
+This model has been proposed by Jianjun Pan (USF, Tampa, FL), but is to the best of our knowledge not published (see also PhD Thesis by Peter Heftberger).
+Additional figures show the volume distribution functions and the resulting electron density profiles.
+
+Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
+
+\subsection{Molecular dynamics simulations}
+
+Molecular dynamics simulation data was collected using the open collaboration method \cite{botan15}.
+Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
+details are given in references or in supplementary material.
+
+\begin{table*}[]
+  \centering
+  \caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
+    $^a$ The number of lipid molecules
+    $^b$ The number of cholesterol molecules
+    $^c$ Cholesterol concentration (mol\%)
+    $^d$ The number of water molecules
+    $^e$ Simulation temperature
+    $^f$ The total simulation time
+    $^g$ Time frames used in the analysis
+    $^h$ Reference link for the downloadable simulation files
+    $^i$ Reference for the full simulation details
+  }\label{systemsCHOL}
+  \begin{tabular}{c c c c c c c c c c c}
+    %\hline
+    Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
+    \hline
+    Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
+    /H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
+    &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
+    &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
+    &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
+    &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
+    &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
+CHARMM36\cite{klauda10,lim12}  & POPC   & 200& 0& 0\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 22& 10\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 50& 20\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 86& 30\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
+                               & POPC   & 200& 134& 40\% & 15030  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
+                               & POPC   & 200& 200& 50\% & 18000  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
+Slipids\cite{jambeck12b,jambeck12,jambeck13b}  & POPC   & 200 & 0 & 0\%   & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 512 & 0 & 0\%   & 23943  & 310 & 170 & 100 & [\citenum{slipidsCHOL0T298}]$^*$ & SI  \\
+                                               & POPC   & 200 & 22 & 10\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 50 & 20\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 86 & 30\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
+                                               & POPC   & 358 & 154 & 30\% & 21183  & 298 & 170 & 100 & [\citenum{slipidsCHOL30T298}]$^*$ & SI  \\
+                                               & POPC   & 200 & 134 & 40\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
+                                               & POPC   & 200 & 200 & 50\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
+                                               & POPC   & 256 & 256 & 50\% & 20334  & 298 & 170 & 100 & [\citenum{slipidsCHOL50T298}]$^*$ & SI  \\ 
+     MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}] \\ 
+                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+\end{tabular}
+\end{table*}
+
+\section{Results and Discussion}
+
+\subsection{Monitoring molecular interactions between POPC and cholesterol in lipid bilayer using NMR and MD simulations}
+
+Addition of cholesterol substantially increases the acyl chain C-H order parameters in PC lipid bilayers (Fig. \ref{OrderParametersCHOL}),
+which can be explained by increased trans conformations in acyl chains \cite{ferreira13,??}.
+%Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
+%and mixture with 50 mol\% of cholesterol from different simulations and experiments
+%are shown in Fig. \ref{OrderParametersCHOL}.
+The absolute values of order parameters of sn-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
+exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
+phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
+%thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
+
 Experimental acyl chain order parameters are typically well reproduced in the state of the art lipid models
@@
 
-\section{Methods}
-
-\subsection{X-ray scattering experiments}
-
-SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations
-has been measured.
-Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27°C.
-Data were analyzed in terms of the SDP-GAP model described in Heftberger et al., J.
-Appl.
-Cryst.
-2013 and Heftberger et al.
-Biophys.
-J.
-2015.
-Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor.
-By fitting the scattered intensity data we obtain both contributions.
-Here we posted only form factors (ASCII format).
-For information on the quality of the fit we also give plots of the fitted intensity data.
-The electron density profile has been modelled in terms of the SDP model (see papers by Kucerka and coworkers), that is volume distribution functions are modelled by individual Gaussians or error functions.
-Cholesterol is also accounted for by two Gaussians.
-This model has been proposed by Jianjun Pan (USF, Tampa, FL), but is to the best of our knowledge not published (see also PhD Thesis by Peter Heftberger).
-Additional figures show the volume distribution functions and the resulting electron density profiles.
-
-Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
-
-\subsection{Molecular dynamics simulations}
-
-Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
-details are given in references or in supplementary material.
-
-\begin{table*}[]
-  \centering
-  \caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
-    $^a$ The number of lipid molecules
-    $^b$ The number of cholesterol molecules
-    $^c$ Cholesterol concentration (mol\%)
-    $^d$ The number of water molecules
-    $^e$ Simulation temperature
-    $^f$ The total simulation time
-    $^g$ Time frames used in the analysis
-    $^h$ Reference link for the downloadable simulation files
-    $^i$ Reference for the full simulation details
-  }\label{systemsCHOL}
-  \begin{tabular}{c c c c c c c c c c c}
-    %\hline
-    Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
-    \hline
-    Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
-    /H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
-    &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
-    &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
-    &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
-    &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
-    &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
-CHARMM36\cite{klauda10,lim12}  & POPC   & 200& 0& 0\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 22& 10\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 50& 20\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 86& 30\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 134& 40\% & 15030  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
-                               & POPC   & 200& 200& 50\% & 18000  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
-Slipids\cite{jambeck12b,jambeck12,jambeck13b}  & POPC   & 200 & 0 & 0\%   & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 512 & 0 & 0\%   & 23943  & 310 & 170 & 100 & [\citenum{slipidsCHOL0T298}]$^*$ & SI  \\
-                                               & POPC   & 200 & 22 & 10\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 200 & 50 & 20\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 200 & 86 & 30\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 358 & 154 & 30\% & 21183  & 298 & 170 & 100 & [\citenum{slipidsCHOL30T298}]$^*$ & SI  \\
-                                               & POPC   & 200 & 134 & 40\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
-                                               & POPC   & 200 & 200 & 50\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
-                                               & POPC   & 256 & 256 & 50\% & 20334  & 298 & 170 & 100 & [\citenum{slipidsCHOL50T298}]$^*$ & SI  \\ 
-     MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}] \\ 
-                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-\end{tabular}
-\end{table*}
-
 % Tables may be be put in the text as floats.
@@
 \end{acknowledgments}
-\newpage
-\appendix
-\begin{center}
-{\bf SUPPLEMENTARY INFORMATION}
-\end{center}
-\section{CHARMM36 results from different simulation packages}
-The results from CHARMM36 model for lipid bilayers from different 
-simulation packages have been reported to give different results in
-the literature \cite{piggot12,lee16}. The results are mainly
-dependent on different Lennard-Jones cut-off settings, but
-all the details are not quite understood. In this work we use
-the results from Gromacs 5 with settings suggested to be optimal
-by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
-these settings to the results simulated with NAMD, OpenMM and literature
-values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
-that Gromacs 5 with settings suggested in webpage gives consistent
-results with the literature and other simulation packages. Thus these
-results are used in the main body of the paper. However, order parameters
-are slightly overestimated respected to the experiments also with these
-settings. \\
-\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
- \begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}
 
-  \caption{\label{programsCOMP}
-    Results for CHARMM36 model \cite{klauda10} from different simulation packages.
-    Discussion going on at {\it https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}.
-  }
-\end{figure}
-
-% Create the reference section using BibTe
 \bibliography{refs.bib}

### 2019-05-07 MANUSCRIPT/manuscriptSUPPL.tex
Add separate supplementary information

\documentclass[journal=jpcbfk]{achemso}

\usepackage[version=3]{mhchem}
\usepackage[T1]{fontenc}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\usepackage{rotating}
\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscript}
\usepackage{soul} % for highlights with \hl{}

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}

\author{Peter Heftberger}
%\email[]{samuli.ollila@aalto.fi}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{University of Graz}

\author{Matti Javanainen}
\affiliation{Department of Physics, Tampere University of Technology, Tampere, Finland}
\affiliation{University of Helsinki}

\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}

\author{Markus Miettinen}
\affiliation{MPI}

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{Georg Pabst}
\affiliation{University of Graz}

\author{Thomas Piggot}
\affiliation{University of Southampton}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\SectionNumbersOn

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thepage}{S\arabic{page}}%

\title{SUPPLEMENTARY INFORMATION: NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper

\begin{document}

\section{CHARMM36 results from different simulation packages}
The results from CHARMM36 model for lipid bilayers from different 
simulation packages have been reported to give different results in
the literature \cite{piggot12,lee16}. The results are mainly
dependent on different Lennard-Jones cut-off settings, but
all the details are not quite understood. In this work we use
the results from Gromacs 5 with settings suggested to be optimal
by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
these settings to the results simulated with NAMD, OpenMM and literature
values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
that Gromacs 5 with settings suggested in webpage gives consistent
results with the literature and other simulation packages. Thus these
results are used in the main body of the paper. However, order parameters
are slightly overestimated respected to the experiments also with these
settings. \\
\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}

  \caption{\label{programsCOMP}
    Results for CHARMM36 model \cite{klauda10} from different simulation packages.
    Discussion going on at {\it https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}.
  }
\end{figure}

\section{Order parameter slopes as a function of cholesterol}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesEXPERIMENT.pdf}
  \caption{\label{slopes}
    Slopes of order parameters as a function of cholesterol from experiments \cite{ferreira13}.
  }
\end{figure*}

 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesBERGER.pdf}
  \caption{\label{slopesberger}
    Slopes of order parameters as a function of cholesterol from Berger simulations.
  }
\end{figure*}

 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesCHARMM.pdf}
  \caption{\label{slopescharmm}
    Slopes of order parameters as a function of cholesterol from CHARMM36 simulations.
  }
\end{figure*}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesMACROG.pdf}
  \caption{\label{slopesmacrog}
    Slopes of order parameters as a function of cholesterol from MacRog simulations.
  }
\end{figure*}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDS.pdf}
  \caption{\label{slopesslipids}
    Slopes of order parameters as a function of cholesterol from Slipids simulations.
  }
\end{figure*}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDST310K.pdf}
  \caption{\label{slopesslipids310K}
    Slopes of order parameters as a function of cholesterol from Slipids simulations at 298 K.
  }
\end{figure*}

% \begin{figure}[]
%  \centering
%  \includegraphics[width=8cm]{../FIGS/slopesBERGER.pdf}
%  \caption{\label{programsCOMP}
%    Slopes of order parameters as a function of cholesterol from Berger simulations.
%  }
%\end{figure}

% Create the reference section using BibTe
\bibliography{refs.bib}

%\newpage
%\section{APPENDIX: The NMR results reported by Tiago Ferreira}

%\listoftodos

\end{document}
%
% ****** End of file aiptemplate.tex ******

### 2019-05-07 MANUSCRIPT/manuscript.tex
Update manuscript

@@
 %thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
-
-Experimental acyl chain order parameters are typically well reproduced in the state of the art lipid models
-for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
-presented here, except that CHARMM36 slightly overestimates the order parameters and MACROG model
-underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
-all models overestimate the order parameters. For Berger/H\"oltje and MACROG models the overestimation is significant,
-while less severe for CHARMM36 and Slipid models.\\
 \todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
@@
 
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
-  \caption{\label{OrderParametersCHOL}
-    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-  }
-\end{figure*}
-
-The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
-(currently only sn-1). \\
-\todo{This figure should be finished and similar for sn-2 should be done as well}
+Experimental acyl chain order parameters of pure POPC lipid bilayers
+are well reproduced by most force fields (Fig. \ref{OrderParametersCHOL}),
+except for the C$_2$ carbon in {\it sn}-1 chain,
+which is typically the case in the state of the art lipid force fields (for review see \cite{ollila16}).
+%This is also mainly observed in the data
+%presented here, except that
+However, the order parameters in the beginning of {\it sn}-1 chain are underestimated in the
+MACROG simulation and CHARMM36 simulations ran also with other than Gromacs
+(section \ref{CHARMMprograms} in the supplementary information) slightly overestimates the order parameters.
+The absolute values of {\it sn}-1 acyl chain order parameters exhibit approximately linear increase upon addition
+of cholesterol also in simulations below equimolar mixture (Figs. \ref{slopesberger}-\ref{slopesslipids310K}).
+However, the slope of the increase (Fig. \ref{OrderParametersCHOLchanges}) and order parameters
+in equimolar mixtures (Fig. \ref{OrderParametersCHOL}) are overestimated by all force fields.
+%For Berger/H\"oltje and MACROG models the overestimation is significant,
+%while less severe for CHARMM36 and Slipid models.
+The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
+with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
 \begin{figure}[]
@@
     Slopes of order parameters as a function of cholesterol.
-    %Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.
   }
-
 \end{figure}
@@
 
+\section{Presenting the quality of lipid bilayer MD simulation with binary mixtures}
+
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
+  \caption{\label{OrderParametersCHOLfitness}
+    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+  }
+\end{figure*}
+
 \section{Conclusions}

### 2019-05-07 MANUSCRIPT/manuscriptSUPPL.tex
Update manuscript

@@
 
-\section{CHARMM36 results from different simulation packages}
+\section{CHARMM36 results from different simulation packages} \label{CHARMMprograms}
 The results from CHARMM36 model for lipid bilayers from different 
@@
 
+ \section{Effect of undulations on order parameters}
+ \todo{This should be written based on contributions in this issue:
+   https://github.com/NMRLipids/NmrLipidsCholXray/issues/16
+ }
+
 % \begin{figure}[]

### 2019-05-08 MANUSCRIPT/manuscript.tex
Update manuscript

@@
 
+\clearpage
+
 \section{Results and Discussion}
@@
 %thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
-\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
-\todo{Lipid14 results from recent article should be added in this figure} 
 \begin{figure*}[]
@@
   }
+  \todo{Lipid14 results?} 
 \end{figure*}
@@
 However, the order parameters in the beginning of {\it sn}-1 chain are underestimated in the
-MACROG simulation and CHARMM36 simulations ran also with other than Gromacs
-(section \ref{CHARMMprograms} in the supplementary information) slightly overestimates the order parameters.
+MACROG simulation and CHARMM36 simulations slightly overestimates the order parameters, also when
+used with other than Gromacs simulation packages (section \ref{CHARMMprograms} in the supplementary information).
 The absolute values of {\it sn}-1 acyl chain order parameters exhibit approximately linear increase upon addition
@@
 \subsection{Lipid bilayer dimensions and density profiles as a function of cholesterol}
-Cholesterol induced changes in lipid packing, thickness and density profiles are suggested
-to be relevant, for example, for membrane protein interactions and cellular membrane
-lateral organization \cite{??}. Experimental and simulation studies show that
-the cholesterol induces the so called ''condensing effect'', i.e. acyl chain ordering
-increases the membrane thickness and
-reduces the area per molecule in lipid bilayers \cite{??}. This is demonstrated
-from simulations in Fig. \ref{apls}, showing the area per PC headgroups and per
-total number of molecules as a function of cholesterol concentration from different
-models. All models show reduction of area per total amount of molecules, partly
+
+Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering
+form factor, which is related to the electron density along membrane normal
+via Fourier transformation \cite{pan12,Heftberger15,Marquardt15,ollila16,??}.
+%methods and have been widely used to validate lipid bilayer simulations \cite{ollila16,??}. 
+The form factor can be translated to density profiles, 
+%While sophisticated models are applied to extract
+area per lipids and bilayer thickness using SDP model or its combination with MD simulations  \cite{pan12,Heftberger15,Marquardt15,??}.
+Sophisticated approaches for single component lipid bilayers are available \cite{??},
+but multicomponent systems are more difficult to interpret \cite{pan12,Heftberger15,Marquardt15,??}.
+%In such case the validation can be done by comparing directly measurable form factors between
+%simulations and experiments. If simulation reproduces the form factor, it can be also
+%used to interpret the experiments.
+
+State of the art force fields give good agreement with experimental form factors for
+pure POPC lipid bilayers \cite{??}, which is the case also here, except for the Berger
+simulation (Fig.~\ref{FormFactors}).
+%The form factors from different simulation models with
+%different cholesterol concentrations are compared to experimental data in Fig.~\ref{FormFactors}. \\
+Upon addition of cholesterol, the third minima in form factor systemically decreases from $\sim$0.42 to $\sim$0.32
+%Cholesterol induced changes in lipid packing, thickness and density profiles are suggested
+%to be relevant, for example, for membrane protein interactions and cellular membrane
+%lateral organization \cite{??}. Experimental and simulation studies show that
+%the cholesterol induces the so called ''condensing effect'', i.e. acyl chain ordering
+%increases the membrane thickness and
+%reduces the area per molecule in lipid bilayers \cite{??}.
+This may be related to the reduced area per molecule upon addition of cholesterol (Fig. \ref{apls}).
+%, showing the area per PC headgroups and per
+%total number of molecules as a function of cholesterol concentration from different
+%models.
+The area per total amount of molecules decreas with cholesterol in all simulations, partly
 due to the smaller area covered by the cholesterol than lipids, but also because
@@
 the area per PC headgroup does not essentially increase up to the addition of
-$\sim$15 mol\% of cholesterol. \\
-\todo{Membrane thickness to be plotted as well.}
+$\sim$15 mol\% of cholesterol. 
+\begin{figure}[]
+ \centering
+ \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
+ \caption{\label{FormFactors}
+   Form factors from simulations and experiments.
+ }
+ \todo{Details about form factor calculation code is discussed in issues
+   https://github.com/NMRLipids/MATCH/issues/56 and https://github.com/NMRLipids/MATCH/issues/50.
+   Once the code is finalized, we should recalculate and check the form factors.
+   CURRENTLY, 10\%-40\% cholesterol concetrations are calculated with a code which gives possibly incorrect
+   heigths for the maxima.
+ } \\
+ \todo{The y-axis scale cannot be explicitly measured in experiments.
+   Therefore, the y-axis of the experimental form factor is typically scaled to match with simulations.
+   This currently not done, because we have several simulations which give inequal form factors, and therefore
+   it is not clear against which simulation results we should scale the experimental results.
+   I have created a issue for this discussion: https://github.com/NMRLipids/NmrLipidsCholXray/issues/18
+ } \\
+ \todo{Not all experimental and simulation data is here (maybe?).}
+\end{figure}
 \begin{figure}[]
@@
 
-Bilayer dimensions are indirectly experimentally accessible by scattering methods 
-and have been widely used to validate lipid bilayer simulations \cite{ollila16,??}. 
-While sophisticated models are applied to extract area per molecule and thicknes for single
-component lipid bilayers, the results for multicomponent systems are more difficult to interpret \cite{pan12,Heftberger15,Marquardt15,??}.
-In such case the validation can be done by comparing directly measurable form factors between
-simulations and experiments. If simulation reproduces the form factor, it can be also
-used to interpret the experiments.  The form factors from different simulation models with
-different cholesterol concentrations are compared to experimental data in Fig.~\ref{FormFactors}. \\
-\todo{Discussion to be finished when the data is solid.}
-
- \begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
-  \caption{\label{FormFactors}
-    Form factors from simulations and experiments.
-  }
-  \todo{Form factor calculation method should be double checked.}
-  \todo{Experimental form factor amplitudes are not scaled to match with simulations, as done usually}
-  \todo{Not all experimental and simulation data is here.}
-\end{figure}
-
+Visual inspection of form factors (Fig. \ref{FormFactors}) and quantitative quality estimation (Fig. \ref{FormFactorsFITNESS})
+suggests that the Berge/Holtje force field gives the best agreement with experiments with large
+cholesterol concentration. This is surprising because the same model overestimated acyl chain
+order parameter increase upon addition of cholesterol more than CHARMM36 or Slipids simulations
+(Figs. \ref{OrderParametersCHOL} and \ref{OrderParametersCHOLchanges}).
+The Berger/Holtje model also exhibits the most pronounced decrease in the area per molecule
+upon addition of cholesterol (Fig. \ref{alps}).
+On the other hand, the electron density profiles from Berger/Holtje simulations are closer
+to the result from the SDP model then other MD simulation models for both pure POPC bilayer
+and equimolar POPC/cholesterol mixture, especially in the middle of the bilayer
+where SDP model gives substantially lower density than MD simulations (Fig. \ref{densities}).
+Therefore, the good quality of form factor Berger/Holtje model with large amounts of cholesterol
+may rather indicate the importance of this low electron density in the bilayer center,
+rather than better quality of lipid-cholesterol interactions in this model.
 \begin{figure}[]
@@
   \caption{\label{densities}
-    Density profiles from simulations and experiments.
+    Electron density profiles from MD simulations and SDP model.
   }
@@
 \section{Presenting the quality of lipid bilayer MD simulation with binary mixtures}
+
+All the tested MD simulation models qualitatively reproduced experimentally observed condenced and
+ordering effects of cholesterol. However, the acyl chain ordering was overestimated and quality
+of form factors reduced with increasing amount of cholesterol. Although the biological importance
+of these details is unclear, overestimated ordering may promote formation of liquid ordered phases and
+inaccuate form factors complicate interpretation of scattering experiment with MD simulations.
+Therefore, MD simulation force fields correctly reproducing the order effects upon addition of the
+cholesterol would be invaluable.
+
+To facilitate the comparison of the quality of complex lipid bilayers,
+we introduce the quantitative quality measures which measure the quality of
+intermolecular interactions in binary lipid bilayers from different perspectives.
+The simple difference between experimental and simulated C-H bond order parameters
+give a very detailed picture about the quality of molecular conformations in all
+positions of lipid bilayer (Fig. \ref{OrderParametersCHOLfitness}).
+However, complexity of such presentation increases with increasing amount of data
+and it may be hard to present in practical format readable for human or machine.
+Using root mean square deviation across regions in the molecule simplifies the presentation,
+but compromizes the spatial resolution. Here, we divide the lipid into headgroup, glycerol
+backbone, sn-1 and sn-2 chains to present quality of binary lipid bilayer in more convenient
+format (Fig. \ref{OrderParametersCHOLfitness}).
+\todo{Discussion to be finished once we have the fitness function defined.}
 
@@
 
-Cholesterol ordering effect is overestimated in Berger/Holtje and MacRog models.
-Slight overestimation is observed also in CHARMM36 and Slipid models, but more
-careful analysis is required to conclude if this is significant or not.
+%Cholesterol ordering effect is overestimated in Berger/Holtje and MacRog models.
+%Slight overestimation is observed also in CHARMM36 and Slipid models, but more
+%careful analysis is required to conclude if this is significant or not.
 

### 2019-05-08 MANUSCRIPT/manuscript.tex
Update manusript

@@
 \section{Introduction}
+
+\todo{We should emphasize the validation aspect of intermolecular interactions in binary lipid mixtures in the introduction.}
+
 Details of intermolecular interactions determine the phase
 behaviour and details lateral organization of lipid and bilayers
-containing cholesterol \cite{ipsen87}.~\todo{Fix Ole's name among the authors of this paper.}
+containing cholesterol \cite{ipsen87}.
 Formation of lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
@@
 
+\subsection{Quantitative quality measure for lipid bilayer structure}
+
+\todo{To be written once we have agreed on the used measure.}
+
 \clearpage
@@
     Slopes of order parameters as a function of cholesterol.
+    Determined by fitting equation $S_{\rm CH}(C_{\rm chol})=k_{\rm OP}C_{\rm chol}+S_{\rm CH}(0)$
+    to the data in figures \ref{slopes}-\ref{slopesslipids310K}.
   }
@@
   }
+    \todo{Figure to be done properly when the fitness code is finished.
+    This is discussed in issues
+    https://github.com/NMRLipids/MATCH/issues/65,
+    https://github.com/NMRLipids/MATCH/issues/64,
+    https://github.com/NMRLipids/MATCH/issues/56, and
+    https://github.com/NMRLipids/MATCH/issues/50
+  }
 \end{figure}
@@
 
-\section{Presenting the quality of lipid bilayer MD simulation with binary mixtures}
+\subsection{Presenting the quality of lipid bilayer MD simulation with binary mixtures}
 
@@
     Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+  }
+  \todo{Figure to be done properly when the fitness code is finished.
+    This is discussed in issues
+    https://github.com/NMRLipids/MATCH/issues/65,
+    https://github.com/NMRLipids/MATCH/issues/64,
+    https://github.com/NMRLipids/MATCH/issues/56, and
+    https://github.com/NMRLipids/MATCH/issues/50
   }

### 2020-06-01 MANUSCRIPT/manuscript.tex
Update manuscript
Update manuscript.tex

--- 2019-05-08 MANUSCRIPT/manuscript.tex
@@
 \affiliation{University of Helsinki}
+
+\author{Jesper J. Madsen}
+\affiliation{Department of Global Health, College of Public Health}
+\affiliation{University of South Florida}
 
@@
                              & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+     Lipid14\cite{walker15}     & POPC   & 120 & 8 & 6\% &  3968 & 303 & 200 & 200 & [\citenum{lipid14sys1}]$^*$ & SI \\ 
+                          & POPC   & 108  & 20  & 16\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys2}]$^*$ & SI    \\
+                          & POPC   &  84  & 44 &  34\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys3}]$^*$ & SI    \\
+                             & POPC   & 64  & 64 & 50\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys4}]$^*$ & SI    \\
+                             & POPC   &  52  & 76 & 59\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys5}]$^*$ & SI    \\
 \end{tabular}
@@
 \begin{acknowledgments}
-% Put your acknowledgments here.
+JJM thanks the Amber community (Benjamin Madej from the Walker lab and Callum Dickson from the Gould lab) for technical assistance related to the lipid14 force field parameters.
 \end{acknowledgments}

### 2020-06-03 MANUSCRIPT/manuscript.tex
Update manuscript.tex

--- 2019-05-08 MANUSCRIPT/manuscript.tex
@@
                              & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
+     Lipid14\cite{walker14,walker15}     & POPC   & 120 & 8 & 6\% &  3968 & 303 & 200 & 200 & [\citenum{lipid14sys1}]$^*$ & SI \\ 
+                          & POPC   & 108  & 20  & 16\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys2}]$^*$ & SI    \\
+                          & POPC   &  84  & 44 &  34\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys3}]$^*$ & SI    \\
+                             & POPC   & 64  & 64 & 50\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys4}]$^*$ & SI    \\
+                             & POPC   &  52  & 76 & 59\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys5}]$^*$ & SI    \\
 \end{tabular}
@@
 \clearpage
+
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
+  \caption{\label{OrderParametersCHOL}
+    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+  }
+  \todo{Lipid14 results?} 
+\end{figure*}
 
@@
 %thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
-  \caption{\label{OrderParametersCHOL}
-    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-  }
-  \todo{Lipid14 results?} 
-\end{figure*}
 
@@
   }
+  \todo{It would be good to plot these slopes from also other experimental data in the literature.}
 \end{figure}

### 2022-12-21 MANUSCRIPT/manuscript.tex
Major update to the manuscript with new simulation data, plots, and text. Mainly by Matti Javanainen.

@@
-%% ****** Start of file aiptemplate.tex ****** %
-%%
-%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
-%%   Version 4.1 of 9 October 2009.
-%%
-%
-% This is a template for producing documents for use with 
-% the REVTEX 4.1 document class and the AIP substyles.
-% 
-% Copy this file to another name and then work on that file.
-% That way, you always have this original template file to use.
-
-%\documentclass[aip,graphicx]{revtex4-1}
-%\documentclass[aip,reprint]{revtex4-1}
-
-%\usepackage{graphicx}
-
-%\draft % marks overfull lines with a black rule on the right
-%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
-%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}
-
-\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}
-
-%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}
+\documentclass[aps,prl,superscriptaddress]{revtex4-2}
 
@@
 \usepackage[obeyFinal]{easy-todo}
+\usepackage{booktabs}
 \usepackage{xr}
-\externaldocument{manuscriptSUPPL}
+
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+
+\externaldocument{MANUSCRIPT/manuscriptSUPPL}
+
 \begin{document}
 
-% Use the \preprint command to place your local institutional report number 
-% on the title page in preprint mode.
-% Multiple \preprint commands are allowed.
-%\preprint{}
-
-\title{NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper
-
-% repeat the \author .. \affiliation  etc. as needed
-% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
-% Explanatory text should go in the []'s, 
-% actual e-mail address or url should go in the {}'s for \email and \homepage.
-% Please use the appropriate macro for the type of information
-
-% \affiliation command applies to all authors since the last \affiliation command. 
-% The \affiliation command should follow the other information.
+\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
 
@@
 \affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
-
+%
 \author{Peter Heftberger}
-%\email[]{samuli.ollila@aalto.fi}
-%\homepage[]{Your web page}
-%\thanks{}
-%\altaffiliation{}
-\affiliation{University of Graz}
-
+\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+%
 \author{Matti Javanainen}
-\affiliation{Department of Physics, Tampere University of Technology, Tampere, Finland}
-\affiliation{University of Helsinki}
-
+\email{matti.javanainen@helsinki.fi}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
+\affiliation{Institute of Biotechnology, University of Helsinki}
+%
 \author{Jesper J. Madsen}
@@
 \affiliation{University of South Florida}
-
+%
 \author{Josef Melcr}
@@
 Prague 6, Czech Republic}
-
+%
 \author{Markus Miettinen}
 \affiliation{MPI}
-
+%
 \author{O. H. Samuli Ollila}
 \email[]{samuli.ollila@helsinki.fi}
-%\homepage[]{Your web page}
-%\thanks{}
-%\altaffiliation{}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 \affiliation{Institute of Biotechnology, University of Helsinki}
-
+%
 \author{Georg Pabst}
-\affiliation{University of Graz}
-
+\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+\affiliation{BioTechMed-Graz, Graz 8010, Austria}
+%
 \author{Thomas Piggot}
-\affiliation{University of Southampton}
-
-% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
-% \noaffiliation is required (may also be used with the \author command).
-%\collaboration{}
-%\noaffiliation
+\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
+\affiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
 
@@
 \begin{abstract}
-% insert abstract here
-The quantitative quality of lipid-cholesterol interactions in atomistic resolution models will be determined against 
-NMR and scattering data.
+Cholesterol is a central building block of biomembranes, where it induces orientational order, slows down diffusion, and renders membranes generally stiffer.
+Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models provide quantitatively different behavior.
+As the field progresses towards simulations of more complex membranes, the fundamental interactions such as those between cholesterol and phospholipids---the key building blocks of biomembranes---such be modelled accurately.
+%
+Here, we perform a systematic comparison of the commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
+Comparison to deuterium order parameters and lateral diffusion coefficients from NMR as well as electron density profiles from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations.
+The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively.
+The development of force fields or the refinement of existing ones can benefit from the included X-ray scattering data set.
 \end{abstract}
 
-%\pacs{}% insert suggested PACS numbers in braces on next line
+\maketitle
 
-\maketitle %\maketitle must follow title, authors, abstract and \pacs
-
-% Body of paper goes here. Use proper sectioning commands. 
-% References should be done using the \cite, \ref, and \label commands
-
-%\label{}
 \section{Introduction}
 
-\todo{We should emphasize the validation aspect of intermolecular interactions in binary lipid mixtures in the introduction.}
+Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across the leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
+A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}.
+CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}.
+In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}.
+This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function by bound CHOL molecules \cite{gimpl2016interaction}.
 
-Details of intermolecular interactions determine the phase
-behaviour and details lateral organization of lipid and bilayers
-containing cholesterol \cite{ipsen87}.
-Formation of lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
-and superlattices \cite{somerharju09} in cellular membranes have been
-suggested to be driven by interactions between cholesterol
-and lipids. While detailed experimental information of these interactions
-is relatively sparse, atomistic resolution molecular dynamics simulations
-have been widely applied to give detailed explanations of lipid bilayer
-lateral organization and lipid-cholesterol interactions \cite{rog14,sodt14,??}.
-However, the simulations must reproduce the measured experimental details, such as NMR
-order parameters and scattering form factors, to be useful in intepretation
-of molecular details in lipid bilayer mixtures.
+While detailed experimental information of the interactions among phospholipids and CHOL is relatively sparse, atomistic resolution molecular dynamics (MD) simulations have been widely applied to provide a detailed view of the lateral organization of lipid bilayers as well as lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed}.
+However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model.
+Fortunately, atomistic \emph{force fields} have seen rapid development recently \cite{leonard2019developing}.
+The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively.
+Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included.
+Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes.
+Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}.
+Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model and even more so their combinations come with their own sets of pros and cons.
+Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature.
+This largely results from the lack of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
 
-Simulations qualitatively reproduce the cholesterol condensation effect,
-but quantitative comparison to robust experimental data has not been
-typically done. This is partly due to the lack of availability of
-systematic experimental data sets for the comparison and partly
-due to the lack of protocols for validating intermolecular interactions
-in lipid bilayer simulations against experiments.
-Scattering experiments are more difficult to interpret for
-mixed lipid bilayers than for single component bilayers \cite{pan12,Heftberger15,Marquardt15,??},
-thus area per molecule values, one of the main quantity used to
-compare simulations to experiments, have not been
-available for lipid cholesterol mixtures. Systematic experimental
-data set for lipid C-H bond order parameters with different cholesterol
-concentrations in POPC bilayer has been published only relatively recently \cite{ferreira13}.
+We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine and phosphatidylglycerol lipids \cite{bacle2021inverse}.
+Here, we extend this work to mixtures of PC and CHOL.
+We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems.
+Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.
 
-In this work we present the experimental scattering form factor
-data for POPC-cholesterol mixtures by systematically increasing the
-cholesterol concentration, scanning the same concentrations that were previously studied with ssNMR~\cite{ferreira13}.
-Our goal is to show that the combination
-of systematically measured C--H bond order parameter and scattering form
-factor data can be used to validate the quality of lipid--cholesterol
-intermolecular interactions in MD simulations. MD simulations can be also
-potentially used to give structural interpretation for form factors
-measured from mixed systems, which is a major challenge for current methods \cite{pan12,Heftberger15,Marquardt15,??}. 
-The approach should be also applicable for mixed lipid bilayers with
-other than lipid--cholesterol mixtures.
-
-%Importance of phospholipid cholesterol interactions in cellular membranes
-%have been reckognized for long time \cite{??}. Various ideas about 
-%lateral heterogeneities driven interactions between cholesterol, phosphatidylcholine
-%other other membrane lipids has been proposed \cite{??}. 
-%Especially coexistence of liquid-ordered and liquid-disordered phases
-%in PC-cholesterol mixtures has been suggested to be important in
-%biological lipid membranes \cite{??}.
-
-%Phase diagram of PC lipid cholesterol systems, including lo-ld coexistence
-%can be explained by simple models where cholesterol is assumed to 
-%induce order in neighboring lipids \cite{??}. This is important
-%achievement in order to understand biomembrane organization. 
-%However, also more detailed interactions driving interactions 
-%between cholesterol and other membrane lipid molecules are
-%often relevant. Several attempts to understand these interactions
-%have been made by using classical molecular dynamics simulations \cite{??}.
-%These simulations give atomistic resolution information, which not
-%directly accessible with any other techinique. However, 
-%the simulations often suffer from inaccuracies in force field
-%descriptions. Acyl chain region is usually well described in 
-%molecular dynamics simulation models, which is not the case for 
-%glycerol backbone and headgroup \cite{??}. 
-%The same applies to the systems with cholesterol. Practically all simulation
-%models qualitatively reproduce the condesing due to cholesterol, observed
-%as increase in acyl chain order parameters and decrease of area per molecule.
-%However, the some models significantly overestimate the changes in headgroup
-%due to the addition of cholesterol \cite{??}.
-
-%On the other hand, detailed analysis of lateral organization of 
-%lipid-cholesterol mixtures reveal significant differences between
-%different models. For example, significant angular organization
-%of lipids around cholesterol was observed in simulations with Berger
-%Höjtje models, while CHARMM36 simulations did not show such behaviour.
-
-Order parameters for C-H bond vectors in lipid bilayer systems, measured
-with $^{13}$C or $^{2}$H NMR techniques, give indirect information about the structural
-sampling of indidual molecules \cite{ollila16}. The observed increase of
-acyl chain order parameters with added cholesterol in simulations and
-experiments can be explained by increased trans conformations in acyl chains \cite{ferreira13,??},
-which is suggested to play critical role in the phase behaviour of PC-lipid--cholesterol
-mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is
-expected to be a necessary condition for a model used to understand lipid--cholesterol
-phase behaviour.
+Our earlier work has demonstrated that the structural ensembles of lipids from an MD simulation can be directly compared to those from solid state NMR experiments \cite{botan15,catte2016molecular,antila2019headgroup,bacle2021inverse}.
+This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques.
+Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying cholesterol concentrations \cite{ferreira13}.
+In addition, we extracted electron density profiles across the lipid bilayer from scattering form factor data measured in this work for a range of PC--CHOL mixtures using small angle X-ray scattering.
+These profiles can be directly compared with the same profiles calculated from the simulation data.
+Whereas NMR provides data on the conformations of individual lipids, form factor captures the details of the structure of the entire lipid bilayer \cite{ollila16}.
+One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious.
+Here, we demonstrate that data extracted from well-sampled simulations can readily be interpolated to any CHOL concentration, allowing the comparison of a single simulation set to experiments performed at different CHOL concentrations.
+We also evaluated whether the simulation models reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient NMR \cite{filippov2003effect,filippov2003influence}.
+This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}.
+With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations, identified the best models, and analyzed their molecular organization as we expect them to best reflect reality.
 
@@
 
-\subsection{X-ray scattering experiments}
+\subsection{X-ray Scattering Experiments}
+\todo{Scattering methods incomplete}
 
-SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations
-has been measured.
-Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27$^\circ$C.
-Data were analyzed in terms of the SDP-GAP model described in Heftberger et al., J.
-Appl.
-Cryst.
-2013 and Heftberger et al.
-Biophys.
-J.
-2015.
+Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract 
+thus area per molecule values, one of the main quantity used to
+compare simulations to experiments, have not been
+available for lipid cholesterol mixtures.
+
+SAXS data on POPC multilamellar vesicles (MLVs) at various CHOL concentrations has been measured.
+Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons, $T=27^\circ$C.
+Data were analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}.
 Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor.
@@
 For information on the quality of the fit we also give plots of the fitted intensity data.
-The electron density profile has been modelled in terms of the SDP model (see papers by Kucerka and coworkers), that is volume distribution functions are modelled by individual Gaussians or error functions.
-Cholesterol is also accounted for by two Gaussians.
-This model has been proposed by Jianjun Pan (USF, Tampa, FL), but is to the best of our knowledge not published (see also PhD Thesis by Peter Heftberger).
+The electron density profile has been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, that is volume distribution functions are modelled by individual Gaussians or error functions.
+CHOL is challenging to describe within the SDP model, and instead of density profiles, the form factors are often compared between simulation and experiment \cite{Kucerka08b}, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}.
+Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), but is to the best of our knowledge not published.
+%(see also PhD Thesis by Peter Heftberger). 
 Additional figures show the volume distribution functions and the resulting electron density profiles.
@@
 
-\subsection{Molecular dynamics simulations}
+\subsection{Molecular Dynamics Simulations}
 
-Molecular dynamics simulation data was collected using the open collaboration method \cite{botan15}.
-Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
-details are given in references or in supplementary material.
+We performed MD simulations of a pure POPC membrane as well as five POPC/cholesterol mixtures with the cholesterol content ranging from 11 to 47~mol-\%.
+In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes.
+The small (64 POPC) and medium (256 POPC) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC).
+The number of POPC molecules was kept constant across the different cholesterol concentrations.
+All membranes were solvated by 50 waters per lipid (POPC or cholesterol).
+The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models.
+All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}.
+The used simulation parameters are provided in Table~\ref{SItab:params}.
+The simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
+
+\begin{table*}[]
+\begin{center}
+    \caption{\label{tab:simulations}%
+    Details of the simulation systems provided for (small/medium/large) system sizes).
+The box dimensions in the membrane plane ($x/y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
+    }
+    \begin{tabular}{c|ccccc}
+    \toprule
+    Cholesterol & POPC & Cholesterol & Water & $x/y$ (nm) & $z$ (nm) \\
+    \midrule
+    0~mol-\%    & 64/256/1024 & 0/0/0       &   3200/12800/51200 & 4.4/9.0/18.1 & 8.9/8.6/8.5    \\
+    11~mol-\%   & 64/256/1024 & 8/32/128    &   3600/14400/57600 & 4.5/9.4/18.1 & 9.4/9.1/9.3    \\
+    20~mol-\%   & 64/256/1024 & 16/64/256   &   4000/16000/64000 & 4.5/9.2/18.3 & 10.2/9.9/10.0  \\
+    29~mol-\%   & 64/256/1024 & 26/104/416  &   4500/18000/72000 & 4.6/9.2/18.5 & 10.8/10.8/10.7 \\
+    38~mol-\%   & 64/256/1024 & 40/160/640  &   5200/20800/83200 & 4.8/9.5/19.2 & 11.3/11.4/11.2 \\
+    47~mol-\%   & 64/256/1024 & 56/224/896  &   6000/24000/96000 & 5.0/10.1/20.0 & 11.8/11.6/11.7 \\
+    \bottomrule
+    \end{tabular}
+\end{center}
+\end{table*}
+
+\subsection{Simulation Analyses}
+
+\paragraph{Electron Density Profiles and Bilayer Dimensions:} The electron density profiles were calculated using \texttt{gmx density} after assigning all atom types with valence electron counts.
+The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
+Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.
+
+\paragraph{Lateral diffusion coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet.
+The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
+%
+\begin{align}
+	\mathrm{MSD}=4D_\mathrm{PBC}\Delta.
+\end{align}
+
+The diffusion coefficients extracted from the three simulation box sizes were fitted with
+%
+\begin{align}
+	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
+	\label{eq:sdlatpbc}
+\end{align}
+%
+where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $mu_\mathrm{m}$ and $mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively.
+$H$ is half the thickness of the water layer and $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length \cite{vogele2018hydrodynamics}.
+The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}.
+The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).
+
+\paragraph{More and Less Ordered Populations:} The angles ($\theta$) between the C--H vectors of the last six carbons of the palmitate chain were extracted and the data across all CHOL concentrations was accumulated.
+These data were fit with two Gaussians centered at 0,
+%
+\begin{align}
+    P(\theta)=a_\mathrm{less}\exp\left(-\frac{\theta}{2\sigma_\mathrm{less}}\right)+a_\mathrm{more}\exp\left(-\frac{\theta}{2\sigma_\mathrm{more}}\right),
+    \label{eq:twogauss}
+\end{align}
+%
+where the two standard deviations ($\sigma$) correspond to the more and less ordered components.
+After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations was independently fitted by Eq.~\eqref{eq:twogauss} with fixed $\sigma$ values.
+These fits provided the pre-factors, whose ratio $a_\mathrm{more}/(a_\mathrm{more}+a_\mathrm{less})$ corresponds to the fraction of the more ordered component in the system.
+
+This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not the nature.
+In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would vary systematically as a function of CHOL concentration.
+
+\paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields.
+The density profiles (as a function of distance from membrane core along the membrane normal), oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains), and diffusion coefficients were interpolated to create continuous 2D matrix for density profiles and order parameters, and a continuous 1D vector for the diffusion coefficients.
+For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over one dimension which was either the distance from membrane core or the carbon atom number, respectively.
+This resulted in a 1D deviation vector as a function of CHOL concentration.
+For diffusion coefficients, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
+All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.
+
+\section{Results and Discussion}
+
+\todo{Results: volmaps of chains around cholesterol?}\\
+\todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
+\todo{Results: Add missing 30\% data to the scattering plots}\\
+\todo{FF minima analysis}
+
+\subsection{Cholesterol Condensing Effect is Manifested Differently in Different Force Fields}
+
+Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}.
+The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}.
+Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.
+
+Thus, we compare here the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species.
+For reference, all form factor profiles from experiment and simulation are shown in Fig.~\ref{SIfig:scattering}.
+As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment.
+Thus, we have also traced the minima in these profiles, and they are plotted in Fig.~\ref{SIfig:ffminima}.
+As simulations and experiments were performed at somewhat different CHOL concentrations, we generated two-dimensional maps for the density profiles using interpolation (see Methods).
+These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column).
+The original electron density profiles used in the generation of these maps are shown in Fig.~\ref{SIfig:densprofs}.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\linewidth]{../FIGS/densitymaps.pdf}
+  \caption{\label{fig:densmaps}%
+  \textbf{Electron density profiles, thickness and area per lipid as a function of CHOL concentration.}
+  %
+  Top: Electron density maps for the simulations using four different force fields.
+  %
+  Bottom left: Electron density map for the experiment.
+The scaling is uniform between all simulations and experiment.
+The original electron density profiles are shown in Fig.~\ref{SIfig:densprofs}.
+The scattering experiments are interpolated to 298~K from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$).
+The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR.
+The effect of system size on the density profiles in simulations is demonstrated in Fig.~\ref{SIfig:densprofssize} in the SI.
+  %
+  Bottom center: Bilayer thickness.
+Thickness is defined as twice the distance from the peak in electron density to the membrane core.
+Experimental data are extracted in a similar manner from electron density profiles obtained with X-ray scattering.
+  %
+  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids. 
+  %
+  The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
+  }
+\end{figure}
+
+Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region rich in electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
+However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields.
+CHOL renders the membrane thicker, resulting in the deviation of the high-density bands with increasing amount of CHOL.
+In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL.
+In Lipid17, however, the membrane gets thicker until the maximum studied CHOL concentration of 47~mol-\%.
+In experiment, the effect of CHOL is generally weaker, and some saturation is evident at high CHOL concentrations.
+These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness.
+Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating, and the same is true for MacRog at higher cholesterol concentrations.
+This uniform membrane structures could indicate small undulations, whereas the less sharp Lipid17 and especially Slipids profiles could indicate a more flexible membrane.
+
+Comparison against experiments (bottom left panel in Fig.~\ref{fig:densmaps}) indicates that the membrane structure is more uniform in the experiment, as the bands are sharper.
+This is also evident in the original electron density profiles in Fig.~\ref{SIfig:densprofs}.
+System size plays a role here, as demonstrated by the density maps calculated for the CHARMM36 simulations in three sizes, shown in Fig.~\ref{SIfig:densprofssize}.
+The larger the system, the larger the bilayer undulations are, which can smear the features of the density profiles.
+Still, even in the smallest system, the band intensities are not as localized as in the experiment.
+
+Comparison of the thickness values (bottom middle panel in Fig.~\ref{fig:densmaps}) demonstrates that the Lipid17 bilayers are too thin at low CHOL concentrations, whereas other models agree well with experiment in this regime.
+Still, all simulation models overshoot the experiment at high CHOL concentrations.
+Experiment and all simulation models except Lipid17 demonstrate saturation after 30~mol-\%.
+
+The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}.
+The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
+However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, \textit{i.e.} a certain amount of CHOL could be added to a phospholipid bilayer without increasing or decreasing its total area \cite{javanainen2017two}.
+We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}.
+The dependence of APL on CHOL concentration follows the trends in thickness inversely in general, yet provide curious differences between the force fields.
+Lipid17 has the largest APL across the entire CHOL contration range, and the addition of CHOL always leads to an increase in the membrane area.
+MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect.
+The profiles for Slipids and CHARMM36 are very similar, and the partial area of CHOL is small or even zero until a concentration of 20~mol-\%.
+Importantly, the observed differences in CHOL condensing effect take place at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
+Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained by \citeauthor{pabst2000structural}\cite{pabst2000structural} and \citeauthor{kuvcerka2006structure}\cite{kuvcerka2006structure} for POPC at 303~K.
+
+\subsection{Acyl Chain Ordering Varies Greatly Between the Force Fields}
+
+Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of indidual molecules \cite{ollila16}.
+Moreover, these parameters are readily extracted from the simulation.
+An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}.
+Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.
+
+Here, as the NMR experiment \cite{ferreira13} was again performed at different CHOL concentrations than simulations (and X-ray scattering measurements), we generated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments, and subtracted these maps to obtain the deviation of simulation from experiment.
+These maps are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate}.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\linewidth]{../FIGS/OP_chains.pdf}
+  \caption{\label{fig:OPmaps}%
+  POPC acyl chain order parameter deviation from experiments.
+Data are shown for palmitate (top row) and oleate (bottom row) and for our force fields (columns).
+Negative values indicate that order is too high ($-S_\mathrm{CD}$ values too negative) in the simulations.
+The values that are within the estimated error range of 0.02 from experiment are considered to be correct ``correct'' and coloured in white.
+The order parameters calculated for the two or three identical carbons were very similar and thus averaged over in simulations, as they have identical values in experiments.
+For the C2 carbon of the oleate chain, the order parameters are forked in simulations and experiments.
+We thus calculated the differences for both the larger and smaller values, and averaged this deviation for the C2 column of oleate.
+  }
+\end{figure}
+
+These original maps in Fig.~\ref{SIfig:orderparameters} demonstrate that the addition of cholesterol substantially increases the acyl chain C--H order parameters for all simulation models and experiments, which can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}.
+The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
+exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
+phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
+
+The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive way for a quantitative comparison of the behaviors of the force fields.
+Here, deviations in the range [$0.02$,0.02] are considered to fall within the error estimate and are thus shown in white.
+Blue indicates that the order parameters are too low in the simulation (their absolute value is larger, \textit{i.e.} the chains are too ordered).
+Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
+
+In CHARMM36, the both the palmitate and oleate chains get too ordered upon increasing CHOL concentration, yet the oleate C9 carbon next to the double bond is too disordered in the simulation.
+The oleate chain shows best agreement with experiment in Slipids with only slight excess ordering at higher CHOL concentrations.
+The excess ordering is more severe for the palmitate chain.
+Still, the major discrepancy between Slipids and experiment is the drastically too disordered C2 and C3 carbons of the palmitate chain.
+This effect was not present in the original Slipids model \cite{botan15}, indicating that it was introduced in the recent reparametrization that improved the head group and glycerol backbone structures of Slipids \cite{grote2020optimization}.
+Lipid17 provides the best overall agreement with experiment, as no segments deviate significantly from experiment at any CHOL concentrations.
+At low CHOL concentrations, the middle of the oleate chain is too disordered, which agrees with the large APL observed in Fig.~\ref{fig:densmaps}.
+At large CHOL concentrations, the oleate chain agrees well with experiment, whereas the palmitate chain has a slightly too ordered C9--C12 segment, yet the deviation is significantly smaller than in the other force fields.
+MacRog behaves reasonably well at low CHOL concentrations, yet the C2--C5 carbons in palmitate are slightly too disordered, whereas the C2 of oleate is somewhat too ordered.
+However, at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.
+
+The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}.
+The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL\cite{botan15} largely persist across all studied CHOL concentrations.
+Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.
+
+\todo{Also experimental cholesterol order parameters are available. Maybe these should be calculated from simulations as well.}
+
+\subsection{The Force Fields Predict Very Different Lateral Mobilities}
+
+Apart from the ordering effect on the bilayer structure, CHOL is also known to make them stiffer and less dynamic \cite{rog2009ordering,filippov2003effect,filippov2003influence}.
+The comparison between lateral diffusion coefficients extracted from simulation and experiment has been limited due to a box-size dependence observed in simulations performed using periodic boundary conditions \cite{vogele2016divergent,vogele2018hydrodynamics,camley2015strong}.
+Here, we tackle this issue by performing simulations with three system sizes, which allows the extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent}  \cite{vogele2018hydrodynamics,vogele2018hydrodynamics}.
+Importantly, reference values are available for POPC/CHOL mixtures from $^1$H pulsed field gradient NMR diffusion measurements on label-free macroscopically aligned bilayers.
+The PBC-corrected lateral diffusion coefficients are shown in the top panel of Fig.~\ref{fig:dynamics} together with experimental values.
+The size-dependence of these values extracted from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=8.7cm]{../FIGS/dynamics.pdf}
+  \caption{\label{fig:dynamics}%
+  \textbf{Dynamic properties of the POPC/cholesterol mixtures.}
+  Top: Lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:dpbc}.
+Experimental data are taken from NMR measurements on well-hydrated samples \cite{filippov2003influence,filippov2003effect}.
+  %
+  Bottom: Shear viscosities obtained from the finite-size correction.
+  %
+  The size-dependence of lateral diffusion and the fits used to obtain the PBC-corrected diffusion coefficients and shear viscosities are shown in Figs.~\ref{SIfig:dvssize} and \ref{SIfig:dvschol}.
+  }
+\end{figure}
+
+The lipid models again show significantly different behavior.
+CHARMM36 and Lipid17 provide drastically too large diffusion coefficients at low CHOL concentrations, yet they converge towards experimental estimates.
+This similar behavior is somewhat surprising, as they behave very differently in terms of both area per POPC (Fig.~\ref{fig:densmaps}) and acyl chain ordering (Fig.~\ref{fig:OPmaps}).
+For CHARMM36, the excessive ordering observed at high CHOL concentrations counterbalances the too fast dynamics observed at low CHOL concentrations.
+However, the behavior of Lipid17 cannot be explained by the order parameters, as they agree with experiment across CHOL concentrations.
+The area per POPC is exaggerated at all CHOL concentrations, yet the deviation from experiment decreases towards larger CHOL concentrations.
+Thus, it seems that CHARMM36 dynamics correlate with chain ordering, whereas those of Lipid17 go hand in hand with area per POPC.
+With MacRog, the dynamics of POPC are already slightly too slow compared to experiment, and the excessive ordering at large CHOL concentrations leads to a larger underestimation of the experimental values.
+Curiously, Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms other force fields in terms of lateral dynamics.
+
+As the results in Fig.~\ref{SIfig:dvschol} demonstrate, the PBC correction can qualitatively change the trends observed for diffusion coefficients.
+This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration.
+For example, Lipid17 and CHARMM36 underestimate the experimental values in all system sizes used here and the effect of cholesterol seems to be well reproduced.
+However, the corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations so that the simulations and experiment almost agree at 47~mol-\% CHOL.
+With MacRog, even the corrected values fall below the experimental ones, yet the slowdown effect of CHOL is again stronger than in experiment.
+Slipids is a clear outlier, as the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent.
+These results highlight that little can be said about CHOL-dependence of lateral diffusion coefficients without performing multiple simulations required for the PBC correction at multiple CHOL concentrations.
+When fine-tuning interaction parameters, this can lead to a massive number of simulations.
+
+The PBC correction, Eq.~\eqref{eq:sdlatpbc} also provides the shear viscosity of the membrane as a fitting parameter.
+Instead of shear viscosity, the surface viscosity is sometimes reported, yet this is simply the product of shear viscosity and membrane thickness (available in Fig.~\ref{fig:densmaps}).
+The shear viscosity values are in line with the lateral diffusion coefficients.
+
+\subsection{Cholesterol Induces Order in a Non-collective Manner}
+
+CHOL could induce order in a lipid membrane in two ways; it could either increase the average order of all phospholipids (``collective transition''), or increase the prevalence of the ordered lipids (``two state model'').
+To clarify this effect, we extracted the C--H angle distributions with respect to the $xy$ plane for the last six carbons of the palmitate chain from all force fields.
+We first accumulated data at all CHOL concentrations, and observed that these data could be fitted with two Gaussian functions.
+These Gaussians were centered at 0$^\circ$, and with the different standard deviations they corresponding to the more and less ordered acyl chains.
+The obtained standard deviations for the more and less ordered chains only showed minor differences between the force fields (rightmost panel in Fig.~\ref{SIfig:chdist}).
+We then fitted the C--H angle distributions for all CHOL concentrations with the two Gaussians with the obtained standard deviations and with their means set to 0.
+These Gaussians provided provided excellent fits to the data with $R^2=0.991\pm 0.005$ (see the leftmost panel of Fig.~\ref{SIfig:chdist} for the fits to the Slipids data).
+The prefactors of the Gaussian functions were considered to represent the fractions of the more and less ordered components in the system.
+In pure POPC, the ordered component was absent, yet its prevalence increased linearly as a function of CHOL, after saturating at high CHOL concentrations (rightmost panel in Fig.~\ref{SIfig:chdist}).
+The success of the two-Gaussian fits in describing the data indicates that with all force fields the bilayer is composed of lipids ordered by CHOL and lipids corresponding to their state in pure POPC membrane, \textit{i.e.} corresponding to the ``two state model'', and CHOL simply changes the fraction of these two states present.
+Indeed, single Gaussian fits with the standard deviation as a free parameter fail when CHOL is present, indicating that the ordering does not occur \textit{via} a ``collective transition''.
+
+\subsection{Quality of Force Fields at Various Cholesterol Concentrations}
+
+The previous sections highlight that the agreement of the studied force fields with experiment depends on both the parameter in question or the CHOL concentration.
+To capture these features in an easily digestible form, we have calculated the relative deviations of the density profiles, the order parameters of the two acyl chains, and the diffusion coefficients between the experiments and all the force fields.
+The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter.
+These data are shown in the two leftmost columns in Fig.~\ref{fig:deviation} as stacked plots.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\linewidth]{../FIGS/deviation.pdf}
+  \caption{\label{fig:deviation}%
+  \textbf{Total deviation of the force fields from experimental data.} 
+%
+  Two leftmost columns:
+  The relative deviations from density profiles, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
+  %
+  The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
+The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
+The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
+%
+  Top right panel:
+  The zoomed-in view to the deviation in density profiles.
+%
+  Bottom right panel:
+  Effect of system size on the deviation of density profiles for the CHARMM36 simulations.
+  }
+\end{figure}
+
+It is evident that the diffusion coefficients and order parameters (``OPs'') contribute more to the total deviation than the density profiles, and the latter are barely visible in the two leftmost columns of Fig.~\ref{fig:deviation}.
+Therefore, the density profile deviations for the different force fields are shown also in the top right panel of Fig.~\ref{fig:deviation}, which reveals fairly similar deviations of 3--5\% for all force fields.
+The lower left panel shows the effect of system size on the deviation of CHARMM36 density profiles from experiment.
+Fig.~\ref{SIfig:densprofssize} demonstrated that smaller systems have more sharp density peaks due to smaller fluctuations, and the small and medium systems indeed agree slightly better with experiment at low CHOL concentrations.
+However, at larger CHOL concentrations the membranes become stiffer, the smearing caused by the bilayer fluctuations is eliminated, and the deviation from experiment is similar with all system sizes.
+
+Coming back to the comparison of force fields in the two leftmost columns of Fig.~\ref{fig:deviation}, we see that CHARMM36 and MacRog never provide the best overall agreement with experiment.
+The latter performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
+At low CHOL concentration, Slipids provides the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
+Still, Slipids is the second best force field even in the high CHOL regime.
+Lipid17 provides the worst agreement with experiment at low CHOL concentrations, but improves significantly towards larger concentrations so that at above $\sim$30~mol-\% of CHOL, Lipid17 provides the best overall agreement with experiment.
+
+\subsection{Molecular Level Structure of PC--CHOL Mixtures}
+
+Molecular dynamics simulations provide a description of the molecular organization of POPC and CHOL in the bilayer, yet the reliability of this description depends on the quality of the used force field.
+In the previous section we established that Slipids and Lipid17 provide the best agreement with experimental properties at low and high CHOL concentrations, respectively.
+Thus, it is reasonable to assume that the molecular pictures provided by these two force fields at the corresponding CHOL concentration regimes are the most realistic.
+
+Eearlier studies have analyzed the organization of CHOL and DSPC chains and observed five and three density peaks for DSPC chains and other CHOL molecules around CHOL molecules \cite{martinez2010cholesterol}.
+
+\textcolor{red}{2D DENSITY MAPS HERE}
+
+Apart from the lateral organization of the molecules, their orientations might also differ between the force fields.
+However, we observed only small differences in the tilt of CHOL molecules, defined by the angle between a vector spanning the rigid ring structure as well as the $z$ axis (membrane lies in the $x,y$ plane).
+The mean angles and their standard deviations are shown in Fig.~\ref{SIfig:choltilt}.
+Lipid17 and Slipids show very consistent values.
+CHARMM36 has slightly smaller tilt angles than them across all CHOL concentrations.
+MacRog has the most tilted CHOL molecules at low CHOL concentrations, but their tilt angle is the smallest at large CHOL concentrations.
+Still, the differences are overall minor.
+
+\section{Conclusions}
+
+To conclude, the recent all-atom force fields all capture the overall trends of CHOL--PC interaction reasonably well; The addition of CHOL leads to acyl chain ordering, which goes hand in hand with the increase in bilayer thickness.
+Area per lipid increases less than what would be expected if the areas of POPC and CHOL were additive, manifesting a condensation effect of CHOL.
+However, beyond these qualitative features, the force fields behave very differently.
+At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, CHOL has a positive partial area in Slipids and Lipid17, a zero partial area in CHARMM36, and a negative partial area in MacRog.
+In addition to these trends, the absolute values of area per lipid and bilayer thickness vary substantially.
+At larger CHOL concentrations all force fields predict a too thick membrane, indicating an excessive ordering effect of CHOL.
+This is indeed verified by the order parameter maps, which suggest that the palmitate chain is overly ordered in all force fields to a varying degree.
+The best agreement with experiment is achieved with Slipids at low CHOL concentration and with Lipid17 at high CHOL concentration.
+In terms of dynamics, the lateral diffusion in pure POPC is too fast in Lipid17 and CHARMM36, and the addition of CHOL slows them both down more than in experiments.
+MacRog has overall too slow dynamics, whereas Slipids agrees quantitatively with the experiment, indicating that it provides the best description of lipid diffusion.
+
+None of the force fields is clearly the best.
+Slipids shows great description of lateral dynamics, and reasonable behavior in terms of membrane thickness and acyl chain ordering.
+Lipid17 provides the best description for the acyl chain ordering, but is otherwise in a relatively poor agreement with experiment.
+Thus, the choice of force field depends not only on the choice of membrane properties (strutural vs.
+dynamic) relevant to the problem, but also on the CHOL concentration range of interest.
+At low  concentrations, Slipids is a safe choice, but at higher concentrations Lipid17 might provide the most realistic description.
+Since many research questions will involve other lipids, proteins, sugars, or drugs, so the quality of the force fields describing these additional molecules might be the decisive factor, together with force field compatibility.
+Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behavior, yet we believe that evaluating the magnitude of these effects is well beyond the scope of this work.
+
+\section{Acknowledgements}
+We acknowledge CSC -- IT Center for Science for computational resources.
+%
+JJM thanks the Amber community (Benjamin Madej from the Walker lab and Callum Dickson from the Gould lab) for technical assistance related to the lipid14 force field parameters. 
+%
+MJ thanks the Academy of Finland (Postdoctoral researcher grant no. 338160) and the Emil Aaltonen foundation for funding.
+
+\bibliography{refs.bib}
+
+%\newpage
+%\section{APPENDIX: The NMR results reported by Tiago Ferreira}
+
+\listoftodos
+
+\end{document}
 
@@
 
-\subsection{Quantitative quality measure for lipid bilayer structure}
-
-\todo{To be written once we have agreed on the used measure.}
-
-\clearpage
-
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
-  \caption{\label{OrderParametersCHOL}
-    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-  }
-  \todo{Lipid14 results?} 
-\end{figure*}
-
-\section{Results and Discussion}
-
-\subsection{Monitoring molecular interactions between POPC and cholesterol in lipid bilayer using NMR and MD simulations}
-
-Addition of cholesterol substantially increases the acyl chain C-H order parameters in PC lipid bilayers (Fig. \ref{OrderParametersCHOL}),
-which can be explained by increased trans conformations in acyl chains \cite{ferreira13,??}.
-%Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
-%and mixture with 50 mol\% of cholesterol from different simulations and experiments
-%are shown in Fig. \ref{OrderParametersCHOL}.
-The absolute values of order parameters of sn-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
-exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
-phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
-%thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
-
-Experimental acyl chain order parameters of pure POPC lipid bilayers
-are well reproduced by most force fields (Fig. \ref{OrderParametersCHOL}),
-except for the C$_2$ carbon in {\it sn}-1 chain,
-which is typically the case in the state of the art lipid force fields (for review see \cite{ollila16}).
-%This is also mainly observed in the data
-%presented here, except that
-However, the order parameters in the beginning of {\it sn}-1 chain are underestimated in the
-MACROG simulation and CHARMM36 simulations slightly overestimates the order parameters, also when
-used with other than Gromacs simulation packages (section \ref{CHARMMprograms} in the supplementary information).
-The absolute values of {\it sn}-1 acyl chain order parameters exhibit approximately linear increase upon addition
-of cholesterol also in simulations below equimolar mixture (Figs. \ref{slopesberger}-\ref{slopesslipids310K}).
-However, the slope of the increase (Fig. \ref{OrderParametersCHOLchanges}) and order parameters
-in equimolar mixtures (Fig. \ref{OrderParametersCHOL}) are overestimated by all force fields.
-%For Berger/H\"oltje and MACROG models the overestimation is significant,
-%while less severe for CHARMM36 and Slipid models.
-The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
-with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/slopes.eps}
-  \caption{\label{OrderParametersCHOLchanges}
-    Slopes of order parameters as a function of cholesterol.
-    Determined by fitting equation $S_{\rm CH}(C_{\rm chol})=k_{\rm OP}C_{\rm chol}+S_{\rm CH}(0)$
-    to the data in figures \ref{slopes}-\ref{slopesslipids310K}.
-  }
-  \todo{It would be good to plot these slopes from also other experimental data in the literature.}
-\end{figure}
-
-\todo{Also experimental cholesterol order parameters are available. Maybe these should be calculated from simulations as well.}
-
-\subsection{Lipid bilayer dimensions and density profiles as a function of cholesterol}
-
-Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering
-form factor, which is related to the electron density along membrane normal
-via Fourier transformation \cite{pan12,Heftberger15,Marquardt15,ollila16,??}.
-%methods and have been widely used to validate lipid bilayer simulations \cite{ollila16,??}. 
-The form factor can be translated to density profiles, 
-%While sophisticated models are applied to extract
-area per lipids and bilayer thickness using SDP model or its combination with MD simulations  \cite{pan12,Heftberger15,Marquardt15,??}.
-Sophisticated approaches for single component lipid bilayers are available \cite{??},
-but multicomponent systems are more difficult to interpret \cite{pan12,Heftberger15,Marquardt15,??}.
-%In such case the validation can be done by comparing directly measurable form factors between
-%simulations and experiments. If simulation reproduces the form factor, it can be also
-%used to interpret the experiments.
-
-State of the art force fields give good agreement with experimental form factors for
-pure POPC lipid bilayers \cite{??}, which is the case also here, except for the Berger
-simulation (Fig.~\ref{FormFactors}).
-%The form factors from different simulation models with
-%different cholesterol concentrations are compared to experimental data in Fig.~\ref{FormFactors}. \\
-Upon addition of cholesterol, the third minima in form factor systemically decreases from $\sim$0.42 to $\sim$0.32
-%Cholesterol induced changes in lipid packing, thickness and density profiles are suggested
-%to be relevant, for example, for membrane protein interactions and cellular membrane
-%lateral organization \cite{??}. Experimental and simulation studies show that
-%the cholesterol induces the so called ''condensing effect'', i.e. acyl chain ordering
-%increases the membrane thickness and
-%reduces the area per molecule in lipid bilayers \cite{??}.
-This may be related to the reduced area per molecule upon addition of cholesterol (Fig. \ref{apls}).
-%, showing the area per PC headgroups and per
-%total number of molecules as a function of cholesterol concentration from different
-%models.
-The area per total amount of molecules decreas with cholesterol in all simulations, partly
-due to the smaller area covered by the cholesterol than lipids, but also because
-lipids ordered by cholesterol require less space. Due to the latter effect,
-the area per PC headgroup does not essentially increase up to the addition of
-$\sim$15 mol\% of cholesterol. 
 \begin{figure}[]
@@
 
+The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
+with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
 \begin{figure}[]
   \centering
-  \includegraphics[width=8cm]{../FIGS/densitiesEXTREMES.eps}
-  \caption{\label{densities}
-    Electron density profiles from MD simulations and SDP model.
+  \includegraphics[width=8cm]{../FIGS/slopes.eps}
+  \caption{\label{OrderParametersCHOLchanges}
+    Slopes of order parameters as a function of cholesterol.
+    Determined by fitting equation $S_{\rm CH}(C_{\rm chol})=k_{\rm OP}C_{\rm chol}+S_{\rm CH}(0)$
+    to the data in figures \ref{slopes}-\ref{slopesslipids310K}.
   }
+  \todo{It would be good to plot these slopes from also other experimental data in the literature.}
 \end{figure}
+
+% \begin{figure}[]
+%   \centering
+%   \includegraphics[width=8cm]{../FIGS/densitiesEXTREMES.eps}
+%   \caption{\label{densities}
+%     Electron density profiles from MD simulations and SDP model.
+%   }
+% \end{figure}
+
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
+  \caption{\label{OrderParametersCHOLfitness}
+    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+  }
+  \todo{Figure to be done properly when the fitness code is finished.
+    This is discussed in issues
+    https://github.com/NMRLipids/MATCH/issues/65,
+    https://github.com/NMRLipids/MATCH/issues/64,
+    https://github.com/NMRLipids/MATCH/issues/56, and
+    https://github.com/NMRLipids/MATCH/issues/50
+  }
+\end{figure*}
+
+% \begin{figure*}[]
+%   \centering
+%   \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
+%   \caption{\label{OrderParametersCHOL}
+%     Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+%   }
+%   \todo{Lipid14 results?} 
+% \end{figure*}
 
@@
 \todo{Discussion to be finished once we have the fitness function defined.}
-
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
-  \caption{\label{OrderParametersCHOLfitness}
-    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-  }
-  \todo{Figure to be done properly when the fitness code is finished.
-    This is discussed in issues
-    https://github.com/NMRLipids/MATCH/issues/65,
-    https://github.com/NMRLipids/MATCH/issues/64,
-    https://github.com/NMRLipids/MATCH/issues/56, and
-    https://github.com/NMRLipids/MATCH/issues/50
-  }
-\end{figure*}
-
-\section{Conclusions}
-
-%Cholesterol ordering effect is overestimated in Berger/Holtje and MacRog models.
-%Slight overestimation is observed also in CHARMM36 and Slipid models, but more
-%careful analysis is required to conclude if this is significant or not.
-
-% Tables may be be put in the text as floats.
-% Here is an example of the general form of a table:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
-% \begin{tabular}{} command.
-%
-% \begin{table}
-% \caption{\label{} }
-% \begin{tabular}{}
-% \end{tabular}
-% \end{table}
-
-% If you have acknowledgments, this puts in the proper section head.
-\begin{acknowledgments}
-JJM thanks the Amber community (Benjamin Madej from the Walker lab and Callum Dickson from the Gould lab) for technical assistance related to the lipid14 force field parameters.
-\end{acknowledgments}
-
-\bibliography{refs.bib}
-
-%\newpage
-%\section{APPENDIX: The NMR results reported by Tiago Ferreira}
-
-\listoftodos
-
-\end{document}
-%
-% ****** End of file aiptemplate.tex ******

### 2022-12-21 MANUSCRIPT/manuscriptSUPPL.tex
Update manuscript
Update manuscriptSUPPL.tex

@@
 \externaldocument{manuscript}
-\usepackage{soul} % for highlights with \hl{}
+\usepackage{soul}
 
@@
 \author{Peter Heftberger}
-%\email[]{samuli.ollila@aalto.fi}
-%\homepage[]{Your web page}
-%\thanks{}
-%\altaffiliation{}
 \affiliation{University of Graz}
@@
 \email{samuli.ollila@helsinki.fi}
-%\homepage[]{Your web page}
-%\thanks{}
-%\altaffiliation{}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 
-% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
-% \noaffiliation is required (may also be used with the \author command).
-%\collaboration{}
-%\noaffiliation
-
 \SectionNumbersOn
@@
 \begin{document}
+
+\section{Simulation Parameters}
+
+The force field specific simulation parameters used for each force field are listed in Table~\ref{tab:params}.
+The parameter input files (\texttt{.mdp}) are provided in the Zenodo portal (links in the main text).
+
+For all simulations, we used an integration time step of 2~fs and the leap-frog integrator of GROMACS.
+The isothermal--isobarit (NPT) ensemble was used with a temperature of 298~K and a pressure of 1~bar.
+All simulations were 1~\textmu{}s long, and trajectories were written every 1~ns.
+The P-LINCS constraint algorithm \cite{hess97,hess07} was used for the bonds noted in Table~\ref{tab:params}.
+
+\begin{sidewaystable}[]
+\begin{center}
+    \caption{\label{SItab:params}%
+    Simulation parameters used for different force fields.
+    }
+    \begin{tabular}{lcccc}
+    \toprule
+    \textbf{Parameter} & \textbf{CHARMM36} & \textbf{Slipids} & \textbf{Lipid17} & \textbf{MacRog} \\
+    \midrule % NB interactions
+    Neighbour list type & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} \\
+    Long-range electrostatics & PME & PME & PME & PME \\
+    LJ cut-off & 1.2~nm & 1.4~nm & 0.9~nm & 1.0~nm \\
+    LJ modifier & Force switch 1.0--1.2~nm & -- & -- & -- \\
+    Dispersion correction & -- & Energy \& pressure \cite{shirts2007accurate} & Energy \& pressure \cite{shirts2007accurate} & Energy \& pressure \cite{shirts2007accurate} \\
+    \midrule % P/T control
+    Thermostat & Nos\'{e}--Hoover \cite{hoover85,nose84} & Stochastic rescaling \cite{bussi07} & Nos\'{e}--Hoover \cite{hoover85,nose84} & Stochastic rescaling \cite{bussi07} \\
+    Time constant ($T$) & 1~ps & 0.5~ps & 1~ps & 0.1~ps \\
+    Coupling groups & Lipids \& water & Lipids \& water & Lipids \& water & Lipids \& water \\
+    Barostat & Parrinello--Rahman \cite{parrinello1981polymorphic} & Berendsen \cite{berendsen1984molecular} & Parrinello--Rahman \cite{parrinello1981polymorphic} & Parrinello--Rahman \cite{parrinello1981polymorphic} \\
+    Coupling type ($P$) & semi-isotropic & semi-isotropic & semi-isotropic & semi-isotropic \\
+    Time constant ($P$) & 5~ps & 10~ps & 5~ps & 4~ps\\
+    Compressibility & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar \\
+    \midrule % Constraints
+    Constraints & Bonds with H & All bonds & Bonds with H & All bonds \\
+    \midrule 
+    Water model & TIPS3P \cite{durell1994solvent}  & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} \\
+    FF Source & CHARMM-GUI & CHARMM-GUI & CHARMM-GUI & Refs.~\citenum{kulig2016cis} \& \citenum{macrogpopc} \\
+    Parameter source & CHARMM-GUI & Slipids website & CHARMM-GUI & Refs.~\citenum{kulig2016cis} \& \citenum{macrogpopc} \\
+    \bottomrule
+    \end{tabular}
+\end{center}
+\end{sidewaystable}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.9\linewidth]{../FIGS/d_vs_size.pdf}
+    \caption{\label{SIfig:dvssize}%
+    \textbf{Dependence of lateral diffusion coefficients on simulation box size.} The values calculated for the lipid centres of mass with \texttt{gmx msd} after eliminating leaflet drift.
+The values for the three system sizes are shown as markers together with fits of Eq.~\eqref{eq:pbc}.
+The values extrapolated to infinite system sizes are also shown in the separate column.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.9\linewidth]{../FIGS/d_vs_chol.pdf}
+    \caption{\label{SIfig:dvschol}%
+     \textbf{Dependence of lateral diffusion coefficients on cholesterol concentration.} Data are shown for all system sizes; small (S), medium (M), and large (L).
+The values extrapolated to infinity are shown as well ($\infty$).
+Experimental measured at two hydration levels, 30 m-\% and 55-\% of water \cite{filippov2003effect,filippov2003influence}.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=8.7cm]{../FIGS/apl_vs_size.pdf}
+    \caption{\label{SIfig:aplvssize}%
+     \textbf{Dependence of area per lipid on simulation box size.} Area per lipid is calculated by dividing the membrane area by the number of lipids in one leaflet.
+Error bars show standard error extracted using block averaging in \texttt{gmx analyze}.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/scattering.pdf}
+    \caption{\label{SIfig:scattering}%
+     \textbf{Scattering intensities as a function of scattering vector from experiments and simulations.} Each of the profiles is shifted vertically with respect to the previous one, by 1 for the experimental profiles and by 50 for the computational ones.
+The minima are marked by filled circles to guide the eye.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/densityprofiles.pdf}
+    \caption{\label{SIfig:densprofs}%
+    \textbf{Electron density profiles.}
+    Each of the profiles is shifted vertically with respect to the previous one by 100 units.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/palmitate.pdf}
+    \caption{\label{SIfig:palmitate}%
+     \textbf{Effect of cholesterol on the acyl chain order parameters of the POPC $sn$-1 (palmitate) chain.}
+     The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/oleate.pdf}
+    \caption{\label{SIfig:oleate}%
+     \textbf{Effect of cholesterol on the acyl chain order parameters of the POPC $sn$-2 (oleate) chain.}
+     The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
+Since the order parameters measured for the two hydrogens bound to the C2 carbon differ, they are both shown in the plots.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/OP_headgroup.pdf}
+    \caption{\label{SIfig:headgroups}%
+     \textbf{The deviation of POPC head group parameters from experimental values as a function of CHOL concentration.}
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=8.7cm]{../FIGS/choltilt.pdf}
+    \caption{\label{SIfig:choltilt}%
+     \textbf{Effect of cholesterol concentration on the tilt of the cholesterol molecules.}
+     The tilt is defined as the angle between the vector connecting the C3 and C17 carbons at the two ends of the rigid ring structure of cholesterol.
+The hydroxyl group is bound to C3 and the hydrocarbon chain to C17.
+The tilt distributions were gathered from the two leaflets, fitted with a Gamma distribution, and the mean values and standard deviations were extracted and shown here as the value and error estimate, respectively.
+The cholesterol concentration was identical between the simulations, but the points are horizontally slightly shifted for clarity.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/densityprofiles_size.pdf}
+    \caption{\label{SIfig:densprofssize}%
+    \textbf{Effect of system size on the density profiles.}
+    Membrane undulations are larger in the larger systems, which leads to the smearing of the electron density profiles.
+Here, data are shown for CHARMM36 in the large (1024 POPC in total), medium (256 POPC in total), or small (64 POPC in total) systems.
+The experimental electron density profile is shown for comparison.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/CH.pdf}
+    \caption{\label{SIfig:chdist}%
+    \textbf{Distributions of C--H bond orientations.}
+    Leftmost panel: The probability distributions for the angle of C--H vectors of the last 6 carbons in the palmitate chain of POPC with respect to the $xy$-plane.
+All distributions are centered around 0, but have different shapes.
+The markers show data extracted from Slipids simulations at various CHOL concentrations, and the solid lines show two-Gaussian fits with fixed standard deviations to the data.
+These Gaussians correspond to the more and less ordered palmitate chains.
+The fixed standard deviations were extracted from a fit of two Gaussians to the data accumulated across all CHOL concentrations.
+The fit with the fixed standard deviations to all CHOL concentrations are excellent with $R^2=0.991\pm 0.005$ (mean$\pm$standard deviation).
+    %
+    Middle panel: The standard deviations of the two Gaussians fitted to the C--H distribution (such as the one in the leftmost panel) for the four force fields. 
+    %
+    Rightmost panel: The fraction of the more ordered component as a function of CHOL concentration extracted simply by dividing the prefactor in the Gaussian distribution for the more ordered component by the sum of the two prefactors.
+Notably, in the CHOL-free system, the ordered component is not present.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=8.7cm]{../FIGS/firstmin.pdf}
+    \caption{\label{SIfig:firstmin}%
+    \textbf{Effecf of CHOL on the location of the first minima in the form factor.}
+    The minima are extracted from the form factors (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab and interpolated to all CHOL concentrations.
+    }
+\end{figure}
+
+\clearpage
+
+\bibliography{refs.bib}
+
+\end{document}
 
@@
 
+\clearpage
 \section{Order parameter slopes as a function of cholesterol}
@@
 
- \section{Effect of undulations on order parameters}
+\clearpage
+ \section{Effect of undulations on order parameters} \label{undulations}
  \todo{This should be written based on contributions in this issue:
@@
 %\end{figure}
-
-% Create the reference section using BibTe
-\bibliography{refs.bib}
-
-%\newpage
-%\section{APPENDIX: The NMR results reported by Tiago Ferreira}
-
-%\listoftodos
-
-\end{document}
-%
-% ****** End of file aiptemplate.tex ******

### 2023-01-03 MANUSCRIPT/manuscript.tex

@@
 
+\todo{Update Author list \& Affiliations}
+
 \date{\today}
@@
 \begin{abstract}
-Cholesterol is a central building block of biomembranes, where it induces orientational order, slows down diffusion, and renders membranes generally stiffer.
-Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models provide quantitatively different behavior.
-As the field progresses towards simulations of more complex membranes, the fundamental interactions such as those between cholesterol and phospholipids---the key building blocks of biomembranes---such be modelled accurately.
+Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, and thereby renders the membrane stiffer.
+Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models predict quantitatively different behavior.
+Such limitations are easily neglected, as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions.
+Still, for their validity, it is crucial that the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol, are modelled accurately.
 %
-Here, we perform a systematic comparison of the commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
-Comparison to deuterium order parameters and lateral diffusion coefficients from NMR as well as electron density profiles from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations.
+Here, we perform a systematic comparison of the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
+Comparison of lipid bilayer simulations with deuterium order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations.
 The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively.
-The development of force fields or the refinement of existing ones can benefit from the included X-ray scattering data set.
+%The performance of the commonly used CHARMM36 model is subpar, as it overestimates diffusion at low cholesterol concentrations, and gets excessively ordered upon the addition of cholesterol. 
+The accompanied X-ray scattering data set measured for this work will foster future force field development and refinement.
 \end{abstract}
@@
 
-Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across the leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
+\todo{Check that the uses of ``model'' and ``force field'' are consistent}
+
+Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
 A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}.
@@
 In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}.
-This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function by bound CHOL molecules \cite{gimpl2016interaction}.
+This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.
 
@@
 However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model.
-Fortunately, atomistic \emph{force fields} have seen rapid development recently \cite{leonard2019developing}.
+Fortunately, in addition to constant refinement of existing parameters, atomistic \emph{force fields} have also seen the expansion to other classes of biomolecules, including lipids, sugars, and nucleic acids \cite{leonard2019developing,marrink2019computational}.
 The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively.
 Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included.
+
 Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes.
 Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}.
-Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model and even more so their combinations come with their own sets of pros and cons.
+Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons.
 Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature.
-This largely results from the lack of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
+This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
 
-We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine and phosphatidylglycerol lipids \cite{bacle2021inverse}.
+We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine (PS) lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}.
 Here, we extend this work to mixtures of PC and CHOL.
@@
 This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques.
-Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying cholesterol concentrations \cite{ferreira13}.
-In addition, we extracted electron density profiles across the lipid bilayer from scattering form factor data measured in this work for a range of PC--CHOL mixtures using small angle X-ray scattering.
-These profiles can be directly compared with the same profiles calculated from the simulation data.
-Whereas NMR provides data on the conformations of individual lipids, form factor captures the details of the structure of the entire lipid bilayer \cite{ollila16}.
+Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}.
+In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer.
+The form factor and electron density profiles can be compared with the same profiles calculated from the simulation data.
+Whereas NMR provides data on the conformations of individual lipids, form factors from scattering captures the details of the structure of the entire lipid bilayer \cite{ollila16}.
+
 One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious.
-Here, we demonstrate that data extracted from well-sampled simulations can readily be interpolated to any CHOL concentration, allowing the comparison of a single simulation set to experiments performed at different CHOL concentrations.
-We also evaluated whether the simulation models reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient NMR \cite{filippov2003effect,filippov2003influence}.
+Here, we demonstrate that if the simulations and experiments sample multiple CHOL concentrations, the order parameters and electron density profiles can be readily interpolated to intermediate CHOL concentrations.
+This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison.
+
+We also evaluated whether the simulation force fields reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient (PFG) NMR \cite{filippov2003effect,filippov2003influence}.
 This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}.
-With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations, identified the best models, and analyzed their molecular organization as we expect them to best reflect reality.
+With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations.
+We then analyzed the molecular organization in the best-performing force field at low and high CHOL range, as we expect them to best reflect this organization in reality.
 
@@
 \subsection{X-ray Scattering Experiments}
-\todo{Scattering methods incomplete}
 
-Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract 
-thus area per molecule values, one of the main quantity used to
-compare simulations to experiments, have not been
-available for lipid cholesterol mixtures.
+\todo{Complete scattering methods}
 
-SAXS data on POPC multilamellar vesicles (MLVs) at various CHOL concentrations has been measured.
-Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons, $T=27^\circ$C.
-Data were analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}.
-Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor.
-By fitting the scattered intensity data we obtain both contributions.
-Here we posted only form factors (ASCII format).
-For information on the quality of the fit we also give plots of the fitted intensity data.
-The electron density profile has been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, that is volume distribution functions are modelled by individual Gaussians or error functions.
-CHOL is challenging to describe within the SDP model, and instead of density profiles, the form factors are often compared between simulation and experiment \cite{Kucerka08b}, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}.
-Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), but is to the best of our knowledge not published.
-%(see also PhD Thesis by Peter Heftberger). 
-Additional figures show the volume distribution functions and the resulting electron density profiles.
+%Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract  thus area per molecule values, one of the main quantity used to compare simulations to experiments, have not been available for lipid cholesterol mixtures.
 
-Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
+We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments.
+The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}.
+The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions.
+In this work, only the form factors are discussed.
+The electron density profiles have been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, \textit{i.e.} volume distribution functions are modelled by individual Gaussians or error functions.
+CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}.
+Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}.
+%Additional figures show the volume distribution functions and the resulting electron density profiles.
+%Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
 
@@
 
-We performed MD simulations of a pure POPC membrane as well as five POPC/cholesterol mixtures with the cholesterol content ranging from 11 to 47~mol-\%.
+We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%.
 In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes.
-The small (64 POPC) and medium (256 POPC) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC).
-The number of POPC molecules was kept constant across the different cholesterol concentrations.
-All membranes were solvated by 50 waters per lipid (POPC or cholesterol).
+The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules).
+The number of POPC molecules was kept constant across the different CHOL concentrations.
+All membranes were solvated by 50 waters per lipid (POPC or CHOL).
 The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models.
 All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}.
-The used simulation parameters are provided in Table~\ref{SItab:params}.
-The simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
+The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
+
+\todo{Run Lipid21 for revision?}
 
@@
     \toprule
-    Cholesterol & POPC & Cholesterol & Water & $x/y$ (nm) & $z$ (nm) \\
+    $[$CHOL$]$ & POPC & CHOL & Water & $x/y$ (nm) & $z$ (nm) \\
     \midrule
@@
 
-\paragraph{Electron Density Profiles and Bilayer Dimensions:} The electron density profiles were calculated using \texttt{gmx density} after assigning all atom types with valence electron counts.
+\paragraph{Electron Density Profiles and Bilayer Dimensions:} The form factors and electron density profiles were calculated using the \texttt{form\_factor.py} script in the NMRlipids Databank (\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts.
 The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
@@
 
-\paragraph{Lateral diffusion coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet.
+\paragraph{Order Parameters:} The C--H order parameters were calculated as
+%
+\begin{align}
+    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
+\end{align}
+%
+where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane.
+In the POPC acyl chains, the values for the 2 (3) hydrogens of CH$_2$ (CH$_3$) groups were averaged, since these groups rotate freely and thus the order parameters are identical for both (all) hydrogens in experiments.
+In simulations, the values are also very similar, justifying this approach.
+An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses.
+The C--H order parameters for the POPC head group were calculated similarly, but values for both hydrogens are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.
+
+\paragraph{Lateral Diffusion Coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet with the \texttt{gmx msd} tool.
 The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
@@
 The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).
+The final simulation box dimensions were used for $L$ and $H$.
+For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.
 
@@
 \todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
-\todo{Results: Add missing 30\% data to the scattering plots}\\
 \todo{FF minima analysis}
@@
 Experimental data are taken from NMR measurements on well-hydrated samples \cite{filippov2003influence,filippov2003effect}.
+The hollow circles show data extracted for Slipids using a shorter Lennard-Jones cutoff (see text).
   %
@@
 
+Slipids has the largest Lennard-Jones (LJ) cutoff of the used force fields of 1.4~nm (see Table~\ref{SItab:params}), and this cutoff is known to affect dynamic properties \cite{leonard2018comparison}.
+Therefore, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids but with a cutoff of 0.9~nm corresponding to that of Lipid17, and the PBC-corrected diffusion coefficient values are shown in the top panel of Fig.~\ref{fig:dynamics} as hollow green circles.
+The decrease in cutoff indeed leads to a minor increase in the lateral diffusion coefficient.
+However, the LJ cutoff alone does not explain the differences between force fields, indicating that they truly spark from differences in the force field interaction parameters.
+
 As the results in Fig.~\ref{SIfig:dvschol} demonstrate, the PBC correction can qualitatively change the trends observed for diffusion coefficients.
 This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration.
-For example, Lipid17 and CHARMM36 underestimate the experimental values in all system sizes used here and the effect of cholesterol seems to be well reproduced.
+For example, Lipid17 and CHARMM36 underestimate the experimental values in all system sizes used here and the effect of CHOL seems to be well reproduced.
 However, the corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations so that the simulations and experiment almost agree at 47~mol-\% CHOL.
@@
   Two leftmost columns:
-  The relative deviations from density profiles, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
+  The relative deviations from the first two minima in the form factor), palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
   %
   The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
-The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
+The differences in the form factor minima in Fig.~\ref{fig:SIfig:ffminima} between experiment and simulation were calculated, normalized against experimental data, and summed together.
 The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
@@
   Top right panel: 
-  The zoomed-in view to the deviation in density profiles.
+  The zoomed-in view to the deviation in the location of the first minimum in the form factor.
 %
   Bottom right panel:
-  Effect of system size on the deviation of density profiles for the CHARMM36 simulations.
+  Deviation of the density profiles calculated from the form factor using the SDS model.
+The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
   }
 \end{figure}
+
+\todo{adjust text to highlight changes in Fig.~\ref{fig:deviation}}
+\todo{include description of Fig.~\ref{SIfig:ffminima}}
 

### 2023-01-03 MANUSCRIPT/manuscriptSUPPL.tex

@@
     \centering
-    \includegraphics[width=8.7cm]{../FIGS/firstmin.pdf}
-    \caption{\label{SIfig:firstmin}%
-    \textbf{Effecf of CHOL on the location of the first minima in the form factor.}
-    The minima are extracted from the form factors (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab and interpolated to all CHOL concentrations.
+    \includegraphics[width=8.7cm]{../FIGS/ffminima.pdf}
+    \caption{\label{SIfig:ffminima}%
+    \textbf{Effecf of CHOL on the location of the first two minima in the form factor.}
+    The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
     }

### 2023-03-24 MANUSCRIPT/manuscript.tex
Update text

@@
-\documentclass[aps,prl,superscriptaddress]{revtex4-2}
-
-\usepackage{rotating} 
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{setspace}
-\usepackage{amsmath}
-\usepackage{epstopdf}
-\usepackage[obeyFinal]{easy-todo}
+\documentclass[journal=jctcce]{achemso}
+\usepackage[version=3]{mhchem} 
+\usepackage[T1]{fontenc}       
+\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+\usepackage{xcolor}
 \usepackage{booktabs}
 \usepackage{xr}
+\usepackage{multirow}
+% \usepackage{rotating} 
+% \usepackage{times}
+% \usepackage{graphicx}
+% \usepackage{setspace}
+% \usepackage{amsmath}
+% \usepackage{epstopdf}
+\usepackage[obeyFinal]{easy-todo}
 
@@
 }
-
-\externaldocument{MANUSCRIPT/manuscriptSUPPL}
-
-\begin{document}
-
-\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
+\myexternaldocument{manuscriptSUPPL}
 
@@
 \author{Matti Javanainen}
-\email{matti.javanainen@helsinki.fi}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 Prague 6, Czech Republic}
-\affiliation{Institute of Biotechnology, University of Helsinki}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\email{matti.javanainen@helsinki.fi}
 %
@@
 \affiliation{Department of Global Health, College of Public Health}
-\affiliation{University of South Florida}
+\alsoaffiliation{University of South Florida}
 %
@@
 \author{O. H. Samuli Ollila}
-\email[]{samuli.ollila@helsinki.fi}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 Prague 6, Czech Republic}
-\affiliation{Institute of Biotechnology, University of Helsinki}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\email{samuli.ollila@helsinki.fi}
 %
@@
 \affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
-\affiliation{BioTechMed-Graz, Graz 8010, Austria}
+\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
 %
@@
 \affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
-\affiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
+\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
+
+\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
+
+\begin{document}
 
 \todo{Update Author list \& Affiliations}
-
-\date{\today}
 
@@
 
-\todo{Complete scattering methods}
-
-%Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract  thus area per molecule values, one of the main quantity used to compare simulations to experiments, have not been available for lipid cholesterol mixtures.
-
 We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments.
 The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}.
+All lipids were purchased from Avanti Polar Lipids  (Alabaster, AL, USA) as dry powders and were used without any further purification.
+All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria).
+Fully hydrated MLVs were prepared using rapid solvent exchange as detailed  previously \cite{rieder2015optimizing,belivcka2017high}.
+This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content.
+
 The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions.
@@
 where the two standard deviations ($\sigma$) correspond to the more and less ordered components.
-After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations was independently fitted by Eq.~\eqref{eq:twogauss} with fixed $\sigma$ values.
+After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations were independently fitted by Eq.~\eqref{eq:twogauss} with the fixed $\sigma$ values.
 These fits provided the pre-factors, whose ratio $a_\mathrm{more}/(a_\mathrm{more}+a_\mathrm{less})$ corresponds to the fraction of the more ordered component in the system.
 
-This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not the nature.
-In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would vary systematically as a function of CHOL concentration.
+This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not their natures.
+In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would decrease as a function of CHOL concentration as the membrane gets more ordered.
 
 \paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields.
-The density profiles (as a function of distance from membrane core along the membrane normal), oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains), and diffusion coefficients were interpolated to create continuous 2D matrix for density profiles and order parameters, and a continuous 1D vector for the diffusion coefficients.
-For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over one dimension which was either the distance from membrane core or the carbon atom number, respectively.
-This resulted in a 1D deviation vector as a function of CHOL concentration.
-For diffusion coefficients, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
+%The density profiles (as a function of distance from membrane core along the membrane normal), 
+The 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, the electron density profiles (as a function of normal distance from bilayer center), the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and diffusion coefficients were interpolated to create continuous 2D matrices for the order parameters and density profiles, and a continuous 1D vector for the diffusion coefficients and form factor minima.
+For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles).
+This resulted in 1D deviation vectors as a function of CHOL concentration.
+For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
 All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.
@@
 \todo{Results: volmaps of chains around cholesterol?}\\
-\todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
-\todo{FF minima analysis}
+\todo{Results: SI: RDFs of POPC chains and CHOL around CHOL}\\
+\todo{Unify colours in figures}\\
 
@@
 Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}.
-The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}.
+The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}.
 Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.
 
-Thus, we compare here the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species.
-For reference, all form factor profiles from experiment and simulation are shown in Fig.~\ref{SIfig:scattering}.
+Thus, we compare here the form factors and the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species or their segments.
+For reference, all scattering intensitities from experiment are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, the form factor profiles from experiment and simulation in Fig.~\ref{SIfig:scattering} and the density profiles from experiment and simulation in Fig.~\ref{SIfig:densprofs}.
+
 As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment.
-Thus, we have also traced the minima in these profiles, and they are plotted in Fig.~\ref{SIfig:ffminima}.
-As simulations and experiments were performed at somewhat different CHOL concentrations, we generated two-dimensional maps for the density profiles using interpolation (see Methods).
+Thus, we first traced the first two minima in the form factors, and they are plotted in Fig.~\ref{SIfig:ffminima}.
+As simulations and experiments were performed at somewhat different CHOL concentrations, we  interpolated the minima to the entire studied range of CHOL concentrations.
+The plots in Fig.~\ref{SIfig:ffminima} highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields.
+CHARMM36 seems to be off more than the other three force fields.
+Above $\sim$25~mol-\% of CHOL, the location of the minimum shifts to larger wave vector values in the experiment, which is curiously not captured by any of the force fields.
+The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields.
+Slipids and Lipid17 are generally in best agreement with the experiment, whereas MacRog and CHARMM36 show poorer agreement.
+
+Moving on to the electron density profiles extracted from the form factors, we generated two-dimensional maps for the density profiles using interpolation of the density profiles in Fig.~\ref{SIfig:densprofs} (see Methods).
 These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column).
-The original electron density profiles used in the generation of these maps are shown in Fig.~\ref{SIfig:densprofs}.
 
@@
 The original electron density profiles are shown in Fig.~\ref{SIfig:densprofs}.
-The scattering experiments are interpolated to 298~K from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$).
-The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR.
 The effect of system size on the density profiles in simulations is demonstrated in Fig.~\ref{SIfig:densprofssize} in the SI.
@@
   Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids.
+The experimental estimates, interpolated to 298~K, are taken from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$).
+The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR.
   %
@@
 
-Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region rich in electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
+Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region containing electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
 However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields.
-CHOL renders the membrane thicker, resulting in the deviation of the high-density bands with increasing amount of CHOL.
+CHOL renders the membrane thicker, resulting in the spreading of the high-density bands with increasing amount of CHOL.
 In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL.
@@
 In experiment, the effect of CHOL is generally weaker, and some saturation is evident at high CHOL concentrations.
-These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness.
-Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating, and the same is true for MacRog at higher cholesterol concentrations.
-This uniform membrane structures could indicate small undulations, whereas the less sharp Lipid17 and especially Slipids profiles could indicate a more flexible membrane.
+These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness extracted from the density peak positions.
+Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating smaller membrane fluctuations that would smear the bands, and the same is true for MacRog at higher cholesterol concentrations.
+The less sharp bands for Lipid17 and especially Slipids profiles indicate that the membranes are more flexible with these force fields.
 
@@
 Still, even in the smallest system, the band intensities are not as localized as in the experiment.
+This difference likely partially stems from the features of the used SDP model, but it could also indicate different elastic properties between simulation and experiment.
 
@@
 The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
-However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, \textit{i.e.} a certain amount of CHOL could be added to a phospholipid bilayer without increasing or decreasing its total area \cite{javanainen2017two}.
+However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without affecting its total area \cite{javanainen2017two}.
 We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}.
@@
 Importantly, the observed differences in CHOL condensing effect take place at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
-Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained by \citeauthor{pabst2000structural}\cite{pabst2000structural} and \citeauthor{kuvcerka2006structure}\cite{kuvcerka2006structure} for POPC at 303~K.
+Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained for POPC by scattering and NMR experiments performed by different teams (See Fig.~\ref{fig:densmaps} and its caption).
+Still, all the simulation force fields provide areas per lipid that fall within the experimental spread for pure POPC.
 
@@
 
-Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of indidual molecules \cite{ollila16}.
+Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of individual molecules \cite{ollila16}.
 Moreover, these parameters are readily extracted from the simulation.
-An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}.
+An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by am increased fraction of \textit{anti} conformations in the acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}.
 Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.
 
-Here, as the NMR experiment \cite{ferreira13} was again performed at different CHOL concentrations than simulations (and X-ray scattering measurements), we generated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments, and subtracted these maps to obtain the deviation of simulation from experiment.
+Here, as the NMR experiments \cite{ferreira13} were again performed at different CHOL concentrations than simulations (and X-ray scattering measurements), we generated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments, and subtracted these maps to obtain the deviation of simulation from experiment.
 These maps are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate}.
@@
   POPC acyl chain order parameter deviation from experiments.
-Data are shown for palmitate (top row) and oleate (bottom row) and for our force fields (columns).
+Data are shown for palmitate (top row) and oleate (bottom row) and for the four force fields (columns).
 Negative values indicate that order is too high ($-S_\mathrm{CD}$ values too negative) in the simulations.
-The values that are within the estimated error range of 0.02 from experiment are considered to be correct ``correct'' and coloured in white.
+The values that are within the estimated error range of 0.02 from experiment are considered to be ``correct'' and coloured in white.
 The order parameters calculated for the two or three identical carbons were very similar and thus averaged over in simulations, as they have identical values in experiments.
@@
 
-These original maps in Fig.~\ref{SIfig:orderparameters} demonstrate that the addition of cholesterol substantially increases the acyl chain C--H order parameters for all simulation models and experiments, which can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}.
-The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
-exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
-phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
-
+These original maps in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate} demonstrate that the addition of cholesterol substantially increases the acyl chain C--H order parameters for all simulation models and experiments, which can be explained by increased \textit{anti} conformations in acyl chains \cite{ferreira13,??}.
+%The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13} exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig.
+\ref{slopes}), which is expected because phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
 The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive way for a quantitative comparison of the behaviors of the force fields.
-Here, deviations in the range [$0.02$,0.02] are considered to fall within the error estimate and are thus shown in white.
-Blue indicates that the order parameters are too low in the simulation (their absolute value is larger, \textit{i.e.} the chains are too ordered).
+Here, deviations in the range [--0.02,0.02] are considered to fall within the error estimate and are thus shown in white.
+Blue indicates that the order parameters are too low in the simulation (their absolute value is larger and the chains are too ordered).
 Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
@@
 The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}.
-The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL\cite{botan15} largely persist across all studied CHOL concentrations.
+The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL \cite{botan15} largely persist across all studied CHOL concentrations.
 Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.
 
-\todo{Also experimental cholesterol order parameters are available. Maybe these should be calculated from simulations as well.}
+\todo{Cholesterol order parameters? Into the deviation metric?}
 
@@
   %
-  Bottom: Shear viscosities obtained from the finite-size correction.
+  Bottom: Shear viscosities obtained from the finite-size correction. Eq.~\eqref{eq:dpbc}.
   %
@@
 The lipid models again show significantly different behavior.
-CHARMM36 and Lipid17 provide drastically too large diffusion coefficients at low CHOL concentrations, yet they converge towards experimental estimates.
+CHARMM36 and Lipid17 provide drastically too large diffusion coefficients at low CHOL concentrations, yet they converge towards experimental estimates upon the addition of CHOL.
 This similar behavior is somewhat surprising, as they behave very differently in terms of both area per POPC (Fig.~\ref{fig:densmaps}) and acyl chain ordering (Fig.~\ref{fig:OPmaps}).
 For CHARMM36, the excessive ordering observed at high CHOL concentrations counterbalances the too fast dynamics observed at low CHOL concentrations.
-However, the behavior of Lipid17 cannot be explained by the order parameters, as they agree with experiment across CHOL concentrations.
+However, the behavior of Lipid17 cannot be explained by the order parameters, as they agree with experiment relatively well across all studied CHOL concentrations.
 The area per POPC is exaggerated at all CHOL concentrations, yet the deviation from experiment decreases towards larger CHOL concentrations.
@@
 However, the corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations so that the simulations and experiment almost agree at 47~mol-\% CHOL.
-With MacRog, even the corrected values fall below the experimental ones, yet the slowdown effect of CHOL is again stronger than in experiment.
+With MacRog, even the corrected values fall below the experimental ones, yet the slowdown effect of CHOL again seems stronger than in experiment.
 Slipids is a clear outlier, as the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent.
@@
 These Gaussians were centered at 0$^\circ$, and with the different standard deviations they corresponding to the more and less ordered acyl chains.
-The obtained standard deviations for the more and less ordered chains only showed minor differences between the force fields (rightmost panel in Fig.~\ref{SIfig:chdist}).
+The obtained standard deviations for the more and less ordered chains only showed minor differences between the force fields (rightmost panel in Fig.~\ref{fig:chdist}).
 We then fitted the C--H angle distributions for all CHOL concentrations with the two Gaussians with the obtained standard deviations and with their means set to 0.
-These Gaussians provided provided excellent fits to the data with $R^2=0.991\pm 0.005$ (see the leftmost panel of Fig.~\ref{SIfig:chdist} for the fits to the Slipids data).
+These Gaussians provided provided excellent fits to the data with $R^2=0.991\pm 0.005$ (see the leftmost panel of Fig.~\ref{fig:chdist} for the fits to the Slipids data).
 The prefactors of the Gaussian functions were considered to represent the fractions of the more and less ordered components in the system.
-In pure POPC, the ordered component was absent, yet its prevalence increased linearly as a function of CHOL, after saturating at high CHOL concentrations (rightmost panel in Fig.~\ref{SIfig:chdist}).
+In pure POPC, the ordered component was absent, yet its prevalence increased linearly as a function of CHOL, after saturating at high CHOL concentrations (rightmost panel in Fig.~\ref{fig:chdist}).
 The success of the two-Gaussian fits in describing the data indicates that with all force fields the bilayer is composed of lipids ordered by CHOL and lipids corresponding to their state in pure POPC membrane, \textit{i.e.} corresponding to the ``two state model'', and CHOL simply changes the fraction of these two states present.
 Indeed, single Gaussian fits with the standard deviation as a free parameter fail when CHOL is present, indicating that the ordering does not occur \textit{via} a ``collective transition''.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/CH.pdf}
+    \caption{\label{fig:chdist}%
+    \textbf{Distributions of C--H bond orientations.}
+    Leftmost panel: The probability distributions for the angle of C--H vectors of the last 6 carbons in the palmitate chain of POPC with respect to the $xy$-plane.
+All distributions are centered around 0, but have different shapes.
+The markers show data extracted from Slipids simulations at various CHOL concentrations, and the solid lines show two-Gaussian fits with fixed standard deviations to the data.
+These Gaussians correspond to the more and less ordered palmitate chains.
+The fixed standard deviations were extracted from a fit of two Gaussians to the data accumulated across all CHOL concentrations.
+The fit with the fixed standard deviations to all CHOL concentrations are excellent with $R^2=0.991\pm 0.005$ (mean $\pm$ standard deviation).
+    %
+    Middle panel: The standard deviations of the two Gaussians fitted to the C--H distribution (such as the one in the leftmost panel) for the four force fields. 
+    %
+    Rightmost panel: The fraction of the more ordered component as a function of CHOL concentration extracted simply by dividing the prefactor in the Gaussian distribution for the more ordered component by the sum of the two prefactors.
+Notably, in the CHOL-free system, the ordered component is not present.
+    }
+\end{figure}
 
@@
 The previous sections highlight that the agreement of the studied force fields with experiment depends on both the parameter in question or the CHOL concentration.
-To capture these features in an easily digestible form, we have calculated the relative deviations of the density profiles, the order parameters of the two acyl chains, and the diffusion coefficients between the experiments and all the force fields.
+To capture these features in an easily digestible form, we have calculated the relative deviations of the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients between the experiments and all the force fields.
 The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter.
@@
   Two leftmost columns:
-  The relative deviations from the first two minima in the form factor), palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
+  The relative deviations from the first two minima in the form factor, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
   %
@@
 
-\todo{adjust text to highlight changes in Fig.~\ref{fig:deviation}}
-\todo{include description of Fig.~\ref{SIfig:ffminima}}
-
-It is evident that the diffusion coefficients and order parameters (``OPs'') contribute more to the total deviation than the density profiles, and the latter are barely visible in the two leftmost columns of Fig.~\ref{fig:deviation}.
-Therefore, the density profile deviations for the different force fields are shown also in the top right panel of Fig.~\ref{fig:deviation}, which reveals fairly similar deviations of 3--5\% for all force fields.
-The lower left panel shows the effect of system size on the deviation of CHARMM36 density profiles from experiment.
-Fig.~\ref{SIfig:densprofssize} demonstrated that smaller systems have more sharp density peaks due to smaller fluctuations, and the small and medium systems indeed agree slightly better with experiment at low CHOL concentrations.
-However, at larger CHOL concentrations the membranes become stiffer, the smearing caused by the bilayer fluctuations is eliminated, and the deviation from experiment is similar with all system sizes.
-
-Coming back to the comparison of force fields in the two leftmost columns of Fig.~\ref{fig:deviation}, we see that CHARMM36 and MacRog never provide the best overall agreement with experiment.
-The latter performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
+The comparison of force fields in the two leftmost columns of Fig.~\ref{fig:deviation} reveals that CHARMM36 and MacRog never provide the best overall agreement with experiment.
+In fact, for CHARMM36 all contributions to the deviation metric are always significant, and it shows the poorest agreement with the form factor minima.
+MacRog performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
 At low CHOL concentration, Slipids provides the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
@@
 Lipid17 provides the worst agreement with experiment at low CHOL concentrations, but improves significantly towards larger concentrations so that at above $\sim$30~mol-\% of CHOL, Lipid17 provides the best overall agreement with experiment.
+Since the contributions from the form factor minima deviations are barely visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
+This panel demonstrates that apart from CHARMM36, all force fields capture the location of the 1\textsuperscript{st} minimum relatively well at low CHOL concentrations, but deviate significantly at larger CHOL concentrations.
+Finally, we also analyzed the deviations in the density profiles in the bottom right panel of Fig.~\ref{fig:deviation}.
+This analysis demonstrates that even the significant deviations as well as the trend as a function of CHOL concentration in the form factor minima are lost as form factors are converted to density profiles.
+This indicates that the form factor minima are a better target value for force field development and refinement.
 
@@
 
-Eearlier studies have analyzed the organization of CHOL and DSPC chains and observed five and three density peaks for DSPC chains and other CHOL molecules around CHOL molecules \cite{martinez2010cholesterol}.
+Earlier studies have analyzed the organization of CHOL and DSPC chains and observed five and three density peaks for DSPC chains and other CHOL molecules around CHOL molecules \cite{martinez2010cholesterol}.
 
-\textcolor{red}{2D DENSITY MAPS HERE}
+\textcolor{red}{2D density maps}
+\textcolor{red}{RDFs to SI?}
 
@@
 
-None of the force fields is clearly the best.
-Slipids shows great description of lateral dynamics, and reasonable behavior in terms of membrane thickness and acyl chain ordering.
-Lipid17 provides the best description for the acyl chain ordering, but is otherwise in a relatively poor agreement with experiment.
+None of the force fields unambigously provides the best agreement with experiment.
+Slipids shows great description of lateral dynamics, and reasonable behavior in terms of membrane thickness and acyl chain ordering, especially at lower and intermediate CHOL concentrations.
+On the other hand, Lipid17 provides a good description for the palmitate chain ordering, but is otherwise in a relatively poor agreement with experiment except for higher CHOL concentrations.
 Thus, the choice of force field depends not only on the choice of membrane properties (strutural vs.
@@
 
-\section{Acknowledgements}
+\begin{suppinfo}
+Simulation parameters used for all the studied force fields.
+%
+Form factor plots from simulations and experiments with the traced minima at various cholesterol concentrations.
+%
+Interpolated 1\textsuperscript{st} and 2\textsuperscript{nd} minima of the form factors as a function of cholesterol concentration from simulation and experiment.
+%
+Electron density profiles from simulation and experiment at various cholesterol concentrations.
+%
+Deuterium order parameters of both acyl chains of POPC from simulation and experiment at various cholesterol concentrations.
+%
+Deviation of POPC head group order parameters from experiment as a function of cholesterol concentration.
+%
+Lateral diffusion coefficients with varying cholesterol concentrations as a function of simulation box size from simulations together with the extrapolation of these values to an infinite box size.
+%
+Diffusion coefficients as a function of cholesterol concentration from simulations with varying box sizes together with comparison to experiment.
+%
+Effect of simulation box size on the area per lipid across the studied cholesterol concentrations.
+%
+Effect of simulation box size on the density profiles across the studied cholesterol concentrations.
+%
+Cholesterol tilt as a function of cholesterol concentration from simulations.
+\end{suppinfo}
+
+\begin{acknowledgement}
 We acknowledge CSC -- IT Center for Science for computational resources.
@@
 MJ thanks the Academy of Finland (Postdoctoral researcher grant no. 338160) and the Emil Aaltonen foundation for funding.
+\end{acknowledgement}
 

### 2023-03-24 MANUSCRIPT/manuscriptSUPPL.tex

@@
 \usepackage{microtype}
+\usepackage{soul}
 \usepackage{xr}
+
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+    \addFileDependency{MANUSCRIPT/#1.tex}%
+    \addFileDependency{MANUSCRIPT/#1.aux}%
+}
+
 \externaldocument{manuscript}
-\usepackage{soul}
 
@@
 \affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
-
+%
 \author{Peter Heftberger}
-\affiliation{University of Graz}
-
+\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+%
 \author{Matti Javanainen}
-\affiliation{Department of Physics, Tampere University of Technology, Tampere, Finland}
-\affiliation{University of Helsinki}
-
+\email{matti.javanainen@helsinki.fi}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+%
 \author{Jesper J. Madsen}
 \affiliation{Department of Global Health, College of Public Health}
-\affiliation{University of South Florida}
-
+\alsoaffiliation{University of South Florida}
+%
 \author{Josef Melcr}
@@
 Prague 6, Czech Republic}
-
+%
 \author{Markus Miettinen}
 \affiliation{MPI}
-
+%
 \author{O. H. Samuli Ollila}
@@
 Prague 6, Czech Republic}
-\affiliation{Institute of Biotechnology, University of Helsinki}
-
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+%
 \author{Georg Pabst}
-\affiliation{University of Graz}
-
+\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
+%
 \author{Thomas Piggot}
-\affiliation{University of Southampton}
+\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
+\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
 
@@
 
-\title{SUPPLEMENTARY INFORMATION: NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper
+\title{SUPPLEMENTARY INFORMATION FOR\\ 
+    NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Resolution Molecular Dynamics Simulations}
 
@@
 
+\tableofcontents
+
 \section{Simulation Parameters}
 
-The force field specific simulation parameters used for each force field are listed in Table~\ref{tab:params}.
+The force field specific simulation parameters used for each force field are listed in Table~\ref{SItab:params}.
 The parameter input files (\texttt{.mdp}) are provided in the Zenodo portal (links in the main text).
-
 For all simulations, we used an integration time step of 2~fs and the leap-frog integrator of GROMACS.
-The isothermal--isobarit (NPT) ensemble was used with a temperature of 298~K and a pressure of 1~bar.
+The isothermal--isobaric (NPT) ensemble was used with a temperature of 298~K and a pressure of 1~bar.
 All simulations were 1~\textmu{}s long, and trajectories were written every 1~ns.
-The P-LINCS constraint algorithm \cite{hess97,hess07} was used for the bonds noted in Table~\ref{tab:params}.
+The P-LINCS constraint algorithm \cite{hess97,hess07} was used for the bonds noted in Table~\ref{SItab:params}.
 
@@
 
+\clearpage
+
+\section{Additional Results}
+
+\subsection{Scattering Intensities}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt0.pdf}
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt5.pdf}\\
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt10.pdf}
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt15.pdf}\\
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt20.pdf}
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt25.pdf}
+    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the next figure.}
+    \label{SIfig:scattering1}
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt30.pdf}
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt35.pdf}\\
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt40.pdf}
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt45.pdf}\\
+    \includegraphics[width=0.49\linewidth]{FIGS/scatt50.pdf}
+    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the previous figure.}
+    \label{SIfig:scattering2}
+\end{figure}
+
+\subsection{Form Factors \& Electron Density Profiles}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.86\linewidth]{../FIGS/scattering.pdf}
+    \caption{\label{SIfig:scattering}%
+     \textbf{Scattering intensities as a function of scattering vector from experiments and simulations.} Each of the profiles is shifted vertically with respect to the previous one, by 1 for the experimental profiles and by 50 for the computational ones.
+The minima are marked by filled circles to guide the eye.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=8.7cm]{../FIGS/ffminima.pdf}
+    \caption{\label{SIfig:ffminima}%
+    \textbf{Effecf of CHOL on the location of the first two minima in the form factor.}
+    The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/densityprofiles.pdf}
+    \caption{\label{SIfig:densprofs}%
+    \textbf{Electron density profiles.}
+    Each of the profiles is shifted vertically with respect to the previous one by 100 units.
+    }
+\end{figure}
+
+\clearpage
+\subsection{Order Parameters}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.9\linewidth]{../FIGS/palmitate.pdf}
+    \caption{\label{SIfig:palmitate}%
+     \textbf{Effect of cholesterol on the acyl chain order parameters of the POPC $sn$-1 (palmitate) chain.}
+     The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.9\linewidth]{../FIGS/oleate.pdf}
+    \caption{\label{SIfig:oleate}%
+     \textbf{Effect of cholesterol on the acyl chain order parameters of the POPC $sn$-2 (oleate) chain.}
+     The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
+Since the order parameters measured for the two hydrogens bound to the C2 carbon differ, they are both shown in the plots.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{../FIGS/OP_headgroup.pdf}
+    \caption{\label{SIfig:headgroups}%
+     \textbf{The deviation of POPC head group parameters from experimental values as a function of CHOL concentration.}
+    }
+\end{figure}
+
+\clearpage
+\subsection{Dynamic Properties}
+
 \begin{figure}[htb!]
@@
 
+\clearpage
+\subsection{Finite-Size Effects}
+
 \begin{figure}[htb!]
@@
 Error bars show standard error extracted using block averaging in \texttt{gmx analyze}.
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/scattering.pdf}
-    \caption{\label{SIfig:scattering}%
-     \textbf{Scattering intensities as a function of scattering vector from experiments and simulations.} Each of the profiles is shifted vertically with respect to the previous one, by 1 for the experimental profiles and by 50 for the computational ones.
-The minima are marked by filled circles to guide the eye.
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/densityprofiles.pdf}
-    \caption{\label{SIfig:densprofs}%
-    \textbf{Electron density profiles.}
-    Each of the profiles is shifted vertically with respect to the previous one by 100 units.
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/palmitate.pdf}
-    \caption{\label{SIfig:palmitate}%
-     \textbf{Effect of cholesterol on the acyl chain order parameters of the POPC $sn$-1 (palmitate) chain.}
-     The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/oleate.pdf}
-    \caption{\label{SIfig:oleate}%
-     \textbf{Effect of cholesterol on the acyl chain order parameters of the POPC $sn$-2 (oleate) chain.}
-     The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
-Since the order parameters measured for the two hydrogens bound to the C2 carbon differ, they are both shown in the plots.
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/OP_headgroup.pdf}
-    \caption{\label{SIfig:headgroups}%
-     \textbf{The deviation of POPC head group parameters from experimental values as a function of CHOL concentration.}
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=8.7cm]{../FIGS/choltilt.pdf}
-    \caption{\label{SIfig:choltilt}%
-     \textbf{Effect of cholesterol concentration on the tilt of the cholesterol molecules.}
-     The tilt is defined as the angle between the vector connecting the C3 and C17 carbons at the two ends of the rigid ring structure of cholesterol.
-The hydroxyl group is bound to C3 and the hydrocarbon chain to C17.
-The tilt distributions were gathered from the two leaflets, fitted with a Gamma distribution, and the mean values and standard deviations were extracted and shown here as the value and error estimate, respectively.
-The cholesterol concentration was identical between the simulations, but the points are horizontally slightly shifted for clarity.
     }
@@
 
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/CH.pdf}
-    \caption{\label{SIfig:chdist}%
-    \textbf{Distributions of C--H bond orientations.}
-    Leftmost panel: The probability distributions for the angle of C--H vectors of the last 6 carbons in the palmitate chain of POPC with respect to the $xy$-plane.
-All distributions are centered around 0, but have different shapes.
-The markers show data extracted from Slipids simulations at various CHOL concentrations, and the solid lines show two-Gaussian fits with fixed standard deviations to the data.
-These Gaussians correspond to the more and less ordered palmitate chains.
-The fixed standard deviations were extracted from a fit of two Gaussians to the data accumulated across all CHOL concentrations.
-The fit with the fixed standard deviations to all CHOL concentrations are excellent with $R^2=0.991\pm 0.005$ (mean$\pm$standard deviation).
-    %
-    Middle panel: The standard deviations of the two Gaussians fitted to the C--H distribution (such as the one in the leftmost panel) for the four force fields. 
-    %
-    Rightmost panel: The fraction of the more ordered component as a function of CHOL concentration extracted simply by dividing the prefactor in the Gaussian distribution for the more ordered component by the sum of the two prefactors.
-Notably, in the CHOL-free system, the ordered component is not present.
-    }
-\end{figure}
+\clearpage
+\subsection{Cholesterol Tilt}
 
@@
     \centering
-    \includegraphics[width=8.7cm]{../FIGS/ffminima.pdf}
-    \caption{\label{SIfig:ffminima}%
-    \textbf{Effecf of CHOL on the location of the first two minima in the form factor.}
-    The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
+    \includegraphics[width=8.7cm]{../FIGS/choltilt.pdf}
+    \caption{\label{SIfig:choltilt}%
+     \textbf{Effect of cholesterol concentration on the tilt of the cholesterol molecules.}
+     The tilt is defined as the angle between the vector connecting the C3 and C17 carbons at the two ends of the rigid ring structure of cholesterol.
+The hydroxyl group is bound to C3 and the hydrocarbon chain to C17.
+The tilt distributions were gathered from the two leaflets, fitted with a Gamma distribution, and the mean values and standard deviations were extracted and shown here as the value and error estimate, respectively.
+The cholesterol concentration was identical between the simulations, but the points are horizontally slightly shifted for clarity.
     }

### 2023-04-13 MANUSCRIPT/manuscriptSUPPL.tex
Update pfs
Updates from Overleaf

@@
 
-\section{Simulation Parameters}
+\section{Simulation Details}
 
@@
 The P-LINCS constraint algorithm \cite{hess97,hess07} was used for the bonds noted in Table~\ref{SItab:params}.
+NMRlipids databank~\cite{NMRlipidsDatabank} ID numbers of simulations are given in table~\ref{IDtable}.
 
@@
 \end{sidewaystable}
+
+\clearpage
+
+\begin{table}[]
+\begin{tabular}{ccc}
+\textbf{CHARMM36} \\
+\toprule
+\textbf{System} & \textbf{ID} \\
+\midrule
+1024POPC\_128CHOL\_57600SOL\_298K  &  426 \\ 
+64POPC\_3200SOL\_298K  &  546 \\ 
+1024POPC\_896CHOL\_96000SOL\_298K  &  543 \\ 
+64POPC\_40CHOL\_5200SOL\_298K  &  620 \\ 
+64POPC\_56CHOL\_6000SOL\_298K  &  91 \\ 
+1024POPC\_416CHOL\_72000SOL\_298K  &  412 \\ 
+1024POPC\_256CHOL\_64000SOL\_298K  &  88 \\ 
+256POPC\_64CHOL\_16000SOL\_298K  &  72 \\ 
+64POPC\_8CHOL\_3600SOL\_298K  &  109 \\ 
+256POPC\_104CHOL\_18000SOL\_298K  &  298 \\ 
+64POPC\_16CHOL\_4000SOL\_298K  &  525 \\ 
+64POPC\_26CHOL\_4500SOL\_298K  &  393 \\ 
+1024POPC\_640CHOL\_83200SOL\_298K  &  550 \\ 
+256POPC\_160CHOL\_20800SOL\_298K  &  275 \\ 
+256POPC\_224CHOL\_24000SOL\_298K  &  166 \\ 
+256POPC\_32CHOL\_14400SOL\_298K  &  119 \\ 
+\end{tabular}
+\quad
+\begin{tabular}{cc}
+\textbf{Slipids} \\
+\toprule
+\textbf{System} & \textbf{ID} \\
+\midrule
+Slipids\_POPC\_CHOL20\_L  &  691 \\ 
+Slipids\_POPC\_CHOL47\_M  &  712 \\ 
+Slipids\_POPC\_CHOL29\_S  &  693 \\ 
+Slipids\_POPC\_CHOL20\_S  &  692 \\ 
+Slipids\_POPC\_CHOL20\_M  &  687 \\ 
+Slipids\_POPC\_CHOL11\_S  &  672 \\ 
+Slipids\_POPC\_CHOL11\_M  &  681 \\ 
+Slipids\_POPC\_L  &  696 \\ 
+Slipids\_POPC\_CHOL38\_S  &  661 \\ 
+Slipids\_POPC\_M  &  708 \\ 
+Slipids\_POPC\_CHOL38\_L  &  670 \\ 
+Slipids\_POPC\_S  &  664 \\ 
+Slipids\_POPC\_CHOL29\_L  &  709 \\ 
+Slipids\_POPC\_CHOL29\_M  &  659 \\ 
+Slipids\_POPC\_CHOL11\_L  &  697 \\ 
+Slipids\_POPC\_CHOL38\_M  &  668 \\ 
+Slipids\_POPC\_CHOL47\_S  &  703 \\ 
+Slipids\_POPC\_CHOL47\_L  &  682 \\
+\end{tabular}
+\quad
+\begin{tabular}{cc}
+\textbf{Lipid17} \\
+\toprule
+\textbf{System} & \textbf{ID} \\
+\midrule
+Lipid17\_POPC\_CHOL29\_M  &  700 \\ 
+Lipid17\_POPC\_CHOL47\_L  &  717 \\ 
+Lipid17\_POPC\_CHOL11\_L  &  663 \\ 
+Lipid17\_POPC\_CHOL29\_L  &  683 \\ 
+Lipid17\_POPC\_CHOL20\_S  &  662 \\ 
+Lipid17\_POPC\_CHOL47\_M  &  714 \\ 
+Lipid17\_POPC\_CHOL38\_S  &  688 \\ 
+Lipid17\_POPC\_L  &  684 \\ 
+Lipid17\_POPC\_S  &  715 \\ 
+Lipid17\_POPC\_CHOL20\_M  &  666 \\ 
+Lipid17\_POPC\_M  &  657 \\ 
+Lipid17\_POPC\_CHOL29\_S  &  686 \\ 
+Lipid17\_POPC\_CHOL38\_L  &  673 \\ 
+Lipid17\_POPC\_CHOL47\_S  &  707 \\ 
+Lipid17\_POPC\_CHOL11\_S  &  680 \\ 
+Lipid17\_POPC\_CHOL20\_L  &  667 \\ 
+Lipid17\_POPC\_CHOL38\_M  &  694 \\ 
+Lipid17\_POPC\_CHOL11\_M  &  689 \\
+\end{tabular}
+\quad
+\begin{tabular}{cc}
+\textbf{MacRog} \\
+\toprule
+\textbf{System} & \textbf{ID} \\
+\midrule
+MacRog\_POPC\_CHOL38\_S  &  716 \\ 
+MacRog\_POPC\_CHOL29\_L  &  705 \\ 
+MacRog\_POPC\_CHOL47\_M  &  699 \\ 
+MacRog\_POPC\_CHOL47\_L  &  671 \\ 
+MacRog\_POPC\_CHOL20\_L  &  677 \\ 
+MacRog\_POPC\_S  &  674 \\ 
+MacRog\_POPC\_CHOL20\_M  &  706 \\ 
+MacRog\_POPC\_CHOL38\_L  &  698 \\ 
+MacRog\_POPC\_CHOL38\_M  &  679 \\ 
+MacRog\_POPC\_M  &  675 \\ 
+MacRog\_POPC\_L  &  658 \\ 
+MacRog\_POPC\_CHOL29\_M  &  655 \\ 
+MacRog\_POPC\_CHOL11\_L  &  660 \\ 
+MacRog\_POPC\_CHOL11\_S  &  665 \\ 
+MacRog\_POPC\_CHOL29\_S  &  690 \\ 
+MacRog\_POPC\_CHOL11\_M  &  695 \\ 
+MacRog\_POPC\_CHOL47\_S  &  704 \\ 
+MacRog\_POPC\_CHOL20\_S  &  702 \\
+\end{tabular}
+\caption{NMRlipids databank IDs of performed simulations.}\label{IDtable}
+\end{table}
 

### 2023-04-13 MANUSCRIPT/manuscript.tex

@@
 In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}.
-This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.
+This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation}, %\textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic},
+in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.
 
-While detailed experimental information of the interactions among phospholipids and CHOL is relatively sparse, atomistic resolution molecular dynamics (MD) simulations have been widely applied to provide a detailed view of the lateral organization of lipid bilayers as well as lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed}.
-However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model.
-Fortunately, in addition to constant refinement of existing parameters, atomistic \emph{force fields} have also seen the expansion to other classes of biomolecules, including lipids, sugars, and nucleic acids \cite{leonard2019developing,marrink2019computational}.
+While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations are widely used to provide a detailed view of complex membranes and the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}.
+This is further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable simulations of complex membranes approaching realistic membrane mixtures.
+
 The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively.
-Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included.
+Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines that enables rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.
 
-Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes.
-Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}.
-Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons.
+While simulating of complex membranes with cholesterol have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge.
+%relies on the quality of the used model. 
+%Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons.
 Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature.
 This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
+%
+Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C-H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}.
+%This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. 
+This approach has been useful for finding the best models for headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE) \cite{bacle2021inverse} and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}.
+Furthermore, quantitative quality measures based on C-H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}.
+However, such automatic quality evaluation is limited to simulations that have experimental data available with corresponding composition and temperature.
+%One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and 
+Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automized in the NMRlipids databank.
 
-We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine (PS) lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}.
-Here, we extend this work to mixtures of PC and CHOL.
-We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems.
-Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.
-
-Our earlier work has demonstrated that the structural ensembles of lipids from an MD simulation can be directly compared to those from solid state NMR experiments \cite{botan15,catte2016molecular,antila2019headgroup,bacle2021inverse}.
-This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques.
-Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}.
+Here, we demonstrate how simulations of binary POPC-cholesterol mixtures can be evaluated against experimental NMR and X-ray scattering data
+%order parameters and electron density profiles can be readily 
+by interpolating through multiple CHOL concentrations.
+Because effect of cholesterol on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously~\cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in cholesterol induced lateral membrane heterogeneity.
+%to intermediate CHOL concentrations.
+This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison.
+We also evaluated 
+%whether the simulation force fields reproduce 
+the dependence of lateral diffusion coefficients on cholesterol against pulsed field gradient (PFG) NMR experiments~\cite{filippov2003effect,filippov2003influence} using the recent theoretical work that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}.
+With the structural and dynamic comparisons established, we estimated the quality of popular force fields at different CHOL concentrations.
+We then used the best-performing force fields to analyze molecular organization at different cholesterol concentrations, as we expect them to best reflect this organization in reality.
+%
+%Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby 
+While we focus here on POPC-cholesterol mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. 
+%Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.
+%Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}.
 In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer.
@@
 
-One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious.
-Here, we demonstrate that if the simulations and experiments sample multiple CHOL concentrations, the order parameters and electron density profiles can be readily interpolated to intermediate CHOL concentrations.
-This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison.
-
-We also evaluated whether the simulation force fields reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient (PFG) NMR \cite{filippov2003effect,filippov2003influence}.
-This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}.
-With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations.
-We then analyzed the molecular organization in the best-performing force field at low and high CHOL range, as we expect them to best reflect this organization in reality.
-
 \section{Methods}
@@
 \subsection{X-ray Scattering Experiments}
-
-We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments.
+Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high} .
+This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content.
+Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification.
+All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria).
 The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}.
-All lipids were purchased from Avanti Polar Lipids  (Alabaster, AL, USA) as dry powders and were used without any further purification.
-All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria).
-Fully hydrated MLVs were prepared using rapid solvent exchange as detailed  previously \cite{rieder2015optimizing,belivcka2017high}.
-This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content.
-
-The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions.
-In this work, only the form factors are discussed.
-The electron density profiles have been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, \textit{i.e.} volume distribution functions are modelled by individual Gaussians or error functions.
-CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}.
-Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}.
+The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtained both contributions.
+The electron density profiles were modelled from form factors using the SDP model where volume distribution functions are modelled by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}.
+%CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}.
+Cholesterol was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
 %Additional figures show the volume distribution functions and the resulting electron density profiles.
-%Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)
+Membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
+The electron density maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
 
@@
 We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%.
-In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes.
-The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules).
+Systems were simulated using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models.
+In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256 or 1024 POPC molecules).
+%The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules).
 The number of POPC molecules was kept constant across the different CHOL concentrations.
 All membranes were solvated by 50 waters per lipid (POPC or CHOL).
-The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models.
 All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}.
 The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
+All simulations were stored to the NMRlipids databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}.
 
@@
 
-\paragraph{Electron Density Profiles and Bilayer Dimensions:} The form factors and electron density profiles were calculated using the \texttt{form\_factor.py} script in the NMRlipids Databank (\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts.
-The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
+\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles 
+%were calculated using the \texttt{form\_factor.py} script in 
+automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank} 
+%(\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts. 
+were used.
+Similarly to the experimental X-ray scattering data, membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
+Locations of maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
 Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.
-
-\paragraph{Order Parameters:} The C--H order parameters were calculated as
 %
-\begin{align}
-    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
-\end{align}
-%
-where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane.
-In the POPC acyl chains, the values for the 2 (3) hydrogens of CH$_2$ (CH$_3$) groups were averaged, since these groups rotate freely and thus the order parameters are identical for both (all) hydrogens in experiments.
-In simulations, the values are also very similar, justifying this approach.
+%\paragraph{Order Parameters:} The C--H order parameters were calculated as
+%%
+%\begin{align}
+%    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
+%\end{align}
+%%
+%where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane. 
+To simplify the interpolation, C-H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged.
+These groups rotate freely and thus the order parameters are essentially identical for both (all) hydrogens in experiments and simulations.
+%, the values are also very similar, justifying this approach. 
 An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses.
-The C--H order parameters for the POPC head group were calculated similarly, but values for both hydrogens are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.
+The C--H order parameters for the POPC head group
+%were calculated similarly, but values for both hydrogens 
+are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.
 
@@
 %
-where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $mu_\mathrm{m}$ and $mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively.
-$H$ is half the thickness of the water layer and $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length \cite{vogele2018hydrodynamics}.
+where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $H$ is half the thickness of the water layer, $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length, and $\mu_\mathrm{m}$ and $\mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively~\cite{vogele2018hydrodynamics}.
 The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}.
@@
 For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.
+
+\paragraph{Quantitative Quality evaluation of cholesterol effect on membrane properties:} To ease the evaluation of simulations against experimental data with non-matching concentrations, we interpolated the effect of cholesterol in simulations and experiments for cholesterol concentrations ranging from 0\% to 50\%.
+%The density profiles (as a function of distance from membrane core along the membrane normal), 
+Continuous 2D matrices were created by interpolation for the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and the electron density profiles (as a function of normal distance from bilayer center), while 1D interpolation was used for the cholesterol dependence of the 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, and diffusion coefficients.
+These interpolations were then used to calculate deviations (in \%) from experimental values across CHOL concentrations to quantify the quality of the lipid force fields.
+For the 2D matrices, the absolute values of the differences between matrices from simulations and experiments were first calculated.
+The averages of differences over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles) were then calculated.
+This resulted in 1D deviation vectors as a function of CHOL concentration.
+For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
+All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of cholesterol concentration.
+For C-H bond order parameters, also the deviation matrices between simulations and experiments were used to illustrate quality of simulations.
 
@@
 
-\paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields.
-%The density profiles (as a function of distance from membrane core along the membrane normal), 
-The 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, the electron density profiles (as a function of normal distance from bilayer center), the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and diffusion coefficients were interpolated to create continuous 2D matrices for the order parameters and density profiles, and a continuous 1D vector for the diffusion coefficients and form factor minima.
-For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles).
-This resulted in 1D deviation vectors as a function of CHOL concentration.
-For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
-All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.
-
 \section{Results and Discussion}
@@
 
-\subsection{Cholesterol Condensing Effect is Manifested Differently in Different Force Fields}
+\subsection{Acyl Chain Ordering Varies Greatly Between the Force Fields}
 
-Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}.
-The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}.
-Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.
+Cholesterol is known to induce order in cell membranes by increasing the fraction of \textit{anti} conformations in the acyl chains of phospholipids \cite{ferreira13}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}.
+Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.
+Cholesterol ordering effect can be experimentally quantified by measuring the C-H bond order parameters using $^{13}$C or $^{2}$H NMR techniques that can be also directly compared with MD simulations~\cite{ollila16}.
+Simulations generally reproduce the cholesterol ordering effect, but order parameters often deviate from experiments at high cholesterol concentrations~\cite{ferreira13,madej15}.
+Furthermore, it has not been clear how accurately different force field parameters capture the details of lipid-cholesterol interactions and which parameters would give most realistic results for simulations of complex mixtures where such interactions play critical roles.
 
-Thus, we compare here the form factors and the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species or their segments.
-For reference, all scattering intensitities from experiment are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, the form factor profiles from experiment and simulation in Fig.~\ref{SIfig:scattering} and the density profiles from experiment and simulation in Fig.~\ref{SIfig:densprofs}.
+Here, we evaluate the cholesterol ordering effect in state of the art force field parameters against C-H bond order parameter data from $^{13}$C NMR experiments measured from POPC-cholesterol mixture with cholesterol concentratons ranging between 0-60\%~\cite{ferreira13}.
+To this end, we first interpolated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments.
+These maps were then subtracted to obtain the deviation maps between simulations and experiments.
+The deviation maps of different force fields are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate}.
 
-As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment.
-Thus, we first traced the first two minima in the form factors, and they are plotted in Fig.~\ref{SIfig:ffminima}.
-As simulations and experiments were performed at somewhat different CHOL concentrations, we  interpolated the minima to the entire studied range of CHOL concentrations.
+The cholesterol ordering effect is manifested in original profiles in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} as substantial increase in absolute values of acyl chain C--H bond order parameters upon addition of cholesterol in all simulation models and experiments.
+%, which can be explained by increased \textit{anti} conformations in acyl chains \cite{ferreira13,??}. 
+%The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13} exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig.
+\ref{slopes}), which is expected because phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
+The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive view for a quantitative comparison of different force fields with experiments.
+In white regions, the simulation results are considered to fall within the experimental error as the deviations are in the range of [--0.02,0.02].
+Blue indicates that the order parameters are too negative, i.e., the acyl chains are too ordered in simulations, and \textit{vice versa} for red.
+Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\linewidth]{../FIGS/OP_chains.pdf}
+  \caption{\label{fig:OPmaps}%
+  POPC acyl chain order parameter deviation from experiments.
+Data are shown for palmitate (top row) and oleate (bottom row) and for the four force fields (columns).
+Negative values indicate that order is too high ($-S_\mathrm{CH}$ values too negative) in the simulations.
+The values that are within the estimated experimental error range of 0.02~\cite{ollila16} are
+  %considered to be ``correct'' and 
+  coloured in white.
+Order parameters of identical hydrogens attached to the same carbons were averaged with the exception of C2 carbon of the oleate chain with the forked order parameters for which differences for both the larger and smaller values were calculated, and the average of this deviation is shown in the C2 column of oleate.
+  }
+\end{figure}
+
+In CHARMM36, the both the palmitate and oleate chains get too ordered upon increasing CHOL concentration, yet the oleate C9 carbon next to the double bond is too disordered in the simulation.
+The oleate chain shows best agreement with experiment in Slipids with only slight excess ordering at higher CHOL concentrations.
+The excess ordering is more severe for the palmitate chain.
+Still, the major discrepancy between Slipids and experiment is the drastically too disordered C2 and C3 carbons of the palmitate chain.
+This effect was not present in the original Slipids model \cite{NMRlipidsDatabank}, indicating that it was introduced in the recent reparametrization that improved the head group and glycerol backbone structures of Slipids \cite{grote2020optimization}.
+Lipid17 provides the best overall agreement with experiment, as no segments deviate significantly from experiment at any CHOL concentrations.
+At low CHOL concentrations, the middle of the oleate chain is too disordered.
+%, which agrees with the large APL observed in Fig.~\ref{fig:densmaps}. 
+At large CHOL concentrations, the oleate chain agrees well with experiment, whereas the palmitate chain has a slightly too ordered C9--C12 segment, yet the deviation is significantly smaller than in the other force fields.
+MacRog behaves reasonably well at low CHOL concentrations, yet the C2--C5 carbons in palmitate are slightly too disordered, whereas the C2 of oleate is somewhat too ordered.
+However, at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.
+
+%The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}.
+The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL \cite{botan15} largely persist across all studied CHOL concentrations.
+Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.
+
+\todo{Cholesterol order parameters? Into the deviation metric?}
+
+\subsection{Cholesterol Effect on membrane properties is Manifested Differently in Different Force Fields}
+
+Cholesterol induced ordering straigthens the acyl chain conformations that leads to thickening of membranes.
+While acyl chain order and membrane thickness are well correlated~\cite{NMRlipidsDatabank}, lipid bilayer dimensions can be accessed more directly by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16}.
+Electron density profile, area per lipid and bilayer thickness can be extracted from the form factor using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,doktorova2020molecular}.
+%Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.
+To complement the evaluation of cholesterol ordering effect against NMR order parameters, we measured also X-ray scattering form factors from POPC-cholesterol mixtures with systematically increasing cholesterol concentrations.
+%
+%Thus, we compare here the form factors and the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species or their segments.
+For reference,
+Scattering intensitities from experiments are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, form factors from experiments and simulations in Fig.~\ref{SIfig:scattering}, and density profiles in Fig.~\ref{SIfig:densprofs}.
+
+While the absolute and relative lobe amplitudes depend on simulation box size, minima in form factors correlate with membrane dimensions~\cite{NMRlipidsDatabank}.
+Thus, we interpolated the locations of first two minima in the form factors to the entire studied range of CHOL concentrations in Fig.~\ref{SIfig:ffminima}.
 The plots in Fig.~\ref{SIfig:ffminima} highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields.
@@
 The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields.
-Slipids and Lipid17 are generally in best agreement with the experiment, whereas MacRog and CHARMM36 show poorer agreement.
+Slipids and Lipid17 are generally in better agreement with the experiment than MacRog and CHARMM36.
 
-Moving on to the electron density profiles extracted from the form factors, we generated two-dimensional maps for the density profiles using interpolation of the density profiles in Fig.~\ref{SIfig:densprofs} (see Methods).
-These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column).
+Effect of cholesterol on structural properties of bilayers is compared between the SDP model (based on experimental form factors) and MD simulation results in Fig.~\ref{fig:densmaps}.
+All models demonstrate increasing thickness upon addition of cholesterol that saturates after approximately 30~mol-\% (bottom middle panel in Fig.~\ref{fig:densmaps}).
+MD simulations agree well with the SDP model below 30~mol-\% but overshoot the SDP results at high cholesterol concentrations.
+Lipid17 simulations are exception as they predict thinner membranes than SDP model at low cholesterol concentrations, and clear saturation of thickness increase is not observed.
+%
+%The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}.
+The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
+However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without affecting its total area \cite{javanainen2017two}.
+We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}.
+%
+The dependence of APL on CHOL concentration follows the trends in thickness inversely in general (bottom right panel of Fig.~\ref{fig:densmaps}), yet provide curious differences between the force fields at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
+Lipid17 has the largest APL across the entire CHOL contration range.
+%, and the addition of CHOL always leads to an increase in the membrane area. 
+MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect.
+The profiles for Slipids and CHARMM36 are very similar with the small or zero cholesterol partial area until a concentration of 20~mol-\%.
+%Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained for POPC by scattering and NMR experiments performed by different teams (See Fig.~\ref{fig:densmaps} and its caption).
+Still, all the simulation force fields provide areas per lipid that fall within the experimental spread for pure POPC.
 
@@
 
+For more detailed comparison of membrane structure, 
+%Moving on to the electron density profiles extracted from the form factors, 
+we interpolated the changes in electron density profiles along membrane normal (Fig.~\ref{SIfig:densprofs}) as a function of cholesterol concentration to create two-dimensional electron density maps shown in Fig.~\ref{fig:densmaps}.
 Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region containing electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
 However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields.
-CHOL renders the membrane thicker, resulting in the spreading of the high-density bands with increasing amount of CHOL.
+%CHOL renders the membrane thicker, resulting in the spreading of the high-density bands with increasing amount of CHOL.
 In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL.
@@
 
-Comparison of the thickness values (bottom middle panel in Fig.~\ref{fig:densmaps}) demonstrates that the Lipid17 bilayers are too thin at low CHOL concentrations, whereas other models agree well with experiment in this regime.
-Still, all simulation models overshoot the experiment at high CHOL concentrations.
-Experiment and all simulation models except Lipid17 demonstrate saturation after 30~mol-\%.
-
-The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}.
-The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
-However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without affecting its total area \cite{javanainen2017two}.
-We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}.
-The dependence of APL on CHOL concentration follows the trends in thickness inversely in general, yet provide curious differences between the force fields.
-Lipid17 has the largest APL across the entire CHOL contration range, and the addition of CHOL always leads to an increase in the membrane area.
-MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect.
-The profiles for Slipids and CHARMM36 are very similar, and the partial area of CHOL is small or even zero until a concentration of 20~mol-\%.
-Importantly, the observed differences in CHOL condensing effect take place at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
-Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained for POPC by scattering and NMR experiments performed by different teams (See Fig.~\ref{fig:densmaps} and its caption).
-Still, all the simulation force fields provide areas per lipid that fall within the experimental spread for pure POPC.
-
-\subsection{Acyl Chain Ordering Varies Greatly Between the Force Fields}
-
-Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of individual molecules \cite{ollila16}.
-Moreover, these parameters are readily extracted from the simulation.
-An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by am increased fraction of \textit{anti} conformations in the acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}.
-Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.
-
-Here, as the NMR experiments \cite{ferreira13} were again performed at different CHOL concentrations than simulations (and X-ray scattering measurements), we generated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments, and subtracted these maps to obtain the deviation of simulation from experiment.
-These maps are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate}.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\linewidth]{../FIGS/OP_chains.pdf}
-  \caption{\label{fig:OPmaps}%
-  POPC acyl chain order parameter deviation from experiments.
-Data are shown for palmitate (top row) and oleate (bottom row) and for the four force fields (columns).
-Negative values indicate that order is too high ($-S_\mathrm{CD}$ values too negative) in the simulations.
-The values that are within the estimated error range of 0.02 from experiment are considered to be ``correct'' and coloured in white.
-The order parameters calculated for the two or three identical carbons were very similar and thus averaged over in simulations, as they have identical values in experiments.
-For the C2 carbon of the oleate chain, the order parameters are forked in simulations and experiments.
-We thus calculated the differences for both the larger and smaller values, and averaged this deviation for the C2 column of oleate.
-  }
-\end{figure}
-
-These original maps in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate} demonstrate that the addition of cholesterol substantially increases the acyl chain C--H order parameters for all simulation models and experiments, which can be explained by increased \textit{anti} conformations in acyl chains \cite{ferreira13,??}.
-%The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13} exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig.
-\ref{slopes}), which is expected because phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
-The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive way for a quantitative comparison of the behaviors of the force fields.
-Here, deviations in the range [--0.02,0.02] are considered to fall within the error estimate and are thus shown in white.
-Blue indicates that the order parameters are too low in the simulation (their absolute value is larger and the chains are too ordered).
-Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
-
-In CHARMM36, the both the palmitate and oleate chains get too ordered upon increasing CHOL concentration, yet the oleate C9 carbon next to the double bond is too disordered in the simulation.
-The oleate chain shows best agreement with experiment in Slipids with only slight excess ordering at higher CHOL concentrations.
-The excess ordering is more severe for the palmitate chain.
-Still, the major discrepancy between Slipids and experiment is the drastically too disordered C2 and C3 carbons of the palmitate chain.
-This effect was not present in the original Slipids model \cite{botan15}, indicating that it was introduced in the recent reparametrization that improved the head group and glycerol backbone structures of Slipids \cite{grote2020optimization}.
-Lipid17 provides the best overall agreement with experiment, as no segments deviate significantly from experiment at any CHOL concentrations.
-At low CHOL concentrations, the middle of the oleate chain is too disordered, which agrees with the large APL observed in Fig.~\ref{fig:densmaps}.
-At large CHOL concentrations, the oleate chain agrees well with experiment, whereas the palmitate chain has a slightly too ordered C9--C12 segment, yet the deviation is significantly smaller than in the other force fields.
-MacRog behaves reasonably well at low CHOL concentrations, yet the C2--C5 carbons in palmitate are slightly too disordered, whereas the C2 of oleate is somewhat too ordered.
-However, at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.
-
-The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}.
-The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL \cite{botan15} largely persist across all studied CHOL concentrations.
-Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.
-
-\todo{Cholesterol order parameters? Into the deviation metric?}
-
 \subsection{The Force Fields Predict Very Different Lateral Mobilities}
@@
 Here, we tackle this issue by performing simulations with three system sizes, which allows the extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent} \cite{vogele2018hydrodynamics,vogele2018hydrodynamics}.
-Importantly, reference values are available for POPC/CHOL mixtures from $^1$H pulsed field gradient NMR diffusion measurements on label-free macroscopically aligned bilayers.
-The PBC-corrected lateral diffusion coefficients are shown in the top panel of Fig.~\ref{fig:dynamics} together with experimental values.
-The size-dependence of these values extracted from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
+The size-dependence of lipid lateral diffusion from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
+The PBC-corrected lateral diffusion coefficients are shown in the top panel of Fig.~\ref{fig:dynamics} together with experimental values from $^1$H pulsed field gradient NMR diffusion measurements on label-free macroscopically aligned bilayers~\cite{filippov2003influence,filippov2003effect}.
 
@@
 
+\subsection{Quality of Force Fields at Various Cholesterol Concentrations}
+
+The previous sections reveal differences in predictions of 
+%highlight that the agreement of the studied 
+force fields for the effect of cholesterol on membrane properties. 
+%with experiment depends on both the parameter in question or the CHOL concentration. 
+%To capture these features in an easily digestible form, 
+To streamline the selection of force fields that best capture the effect of cholesterol on different membrane properties, we used the interpolated data to calculate the relative deviations from experiments (difference of simulated and experimental values divided by the experimental value) for the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients from all the force fields.
+%The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\linewidth]{../FIGS/deviation.pdf}
+  \caption{\label{fig:deviation}%
+  \textbf{Total deviation of the force fields from experimental data.} 
+%
+  Two leftmost columns:
+  The relative deviations from the first two minima in the form factor, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
+  %
+  The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
+The differences in the form factor minima in Fig.~\ref{fig:SIfig:ffminima} between experiment and simulation were calculated, normalized against experimental data, and summed together.
+The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
+%
+  Top right panel: 
+  The zoomed-in view to the deviation in the location of the first minimum in the form factor.
+%
+  Bottom right panel:
+  Deviation of the density profiles calculated from the form factor using the SDS model.
+The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
+  }
+\end{figure}
+
+The deviations in Fig.~\ref{fig:deviation} 
+%The comparison of force fields in the two leftmost columns of Fig.~\ref{fig:deviation} 
+reveals that Slipids provides overall the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
+%Still, Slipids is the second best force field even in the high CHOL regime.
+Lipid17 provides the worst agreement with experiment at low CHOL concentrations, but improves significantly towards larger concentrations so that at
+Lipid17 provides slightly better agreement with experiment above $\sim$30~mol-\% of CHOL than Slipids, but exhibits major deviation from experiments with low cholesterol concentrations. 
+%CHARMM36 and MacRog never provide the best overall agreement with experiment. 
+%In fact,  all contributions to , and it shows the poorest agreement with the form factor minima. 
+MacRog performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
+The deviation metric for CHARMM36 are significant with all cholesterol concentrations.
+%
+Since the contributions from the form factor minima deviations are barely visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
+This panel demonstrates that apart from CHARMM36, all force fields capture the location of the 1\textsuperscript{st} minimum relatively well at low CHOL concentrations, but deviate significantly at larger CHOL concentrations.
+
+Finally, we also analyzed the deviations in the density profiles in the bottom right panel of Fig.~\ref{fig:deviation}.
+This analysis demonstrates that even the significant deviations as well as the trend as a function of CHOL concentration in the form factor minima are lost as form factors are converted to density profiles.
+This indicates that the form factor minima are a better target value for force field development and refinement.
+
 \subsection{Cholesterol Induces Order in a Non-collective Manner}
@@
 
-\subsection{Quality of Force Fields at Various Cholesterol Concentrations}
-
-The previous sections highlight that the agreement of the studied force fields with experiment depends on both the parameter in question or the CHOL concentration.
-To capture these features in an easily digestible form, we have calculated the relative deviations of the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients between the experiments and all the force fields.
-The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter.
-These data are shown in the two leftmost columns in Fig.~\ref{fig:deviation} as stacked plots.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\linewidth]{../FIGS/deviation.pdf}
-  \caption{\label{fig:deviation}%
-  \textbf{Total deviation of the force fields from experimental data.} 
-%
-  Two leftmost columns:
-  The relative deviations from the first two minima in the form factor, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
-  %
-  The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
-The differences in the form factor minima in Fig.~\ref{fig:SIfig:ffminima} between experiment and simulation were calculated, normalized against experimental data, and summed together.
-The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
-%
-  Top right panel: 
-  The zoomed-in view to the deviation in the location of the first minimum in the form factor.
-%
-  Bottom right panel:
-  Deviation of the density profiles calculated from the form factor using the SDS model.
-The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
-  }
-\end{figure}
-
-The comparison of force fields in the two leftmost columns of Fig.~\ref{fig:deviation} reveals that CHARMM36 and MacRog never provide the best overall agreement with experiment.
-In fact, for CHARMM36 all contributions to the deviation metric are always significant, and it shows the poorest agreement with the form factor minima.
-MacRog performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
-At low CHOL concentration, Slipids provides the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
-Still, Slipids is the second best force field even in the high CHOL regime.
-Lipid17 provides the worst agreement with experiment at low CHOL concentrations, but improves significantly towards larger concentrations so that at above $\sim$30~mol-\% of CHOL, Lipid17 provides the best overall agreement with experiment.
-Since the contributions from the form factor minima deviations are barely visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
-This panel demonstrates that apart from CHARMM36, all force fields capture the location of the 1\textsuperscript{st} minimum relatively well at low CHOL concentrations, but deviate significantly at larger CHOL concentrations.
-Finally, we also analyzed the deviations in the density profiles in the bottom right panel of Fig.~\ref{fig:deviation}.
-This analysis demonstrates that even the significant deviations as well as the trend as a function of CHOL concentration in the form factor minima are lost as form factors are converted to density profiles.
-This indicates that the form factor minima are a better target value for force field development and refinement.
-
 \subsection{Molecular Level Structure of PC--CHOL Mixtures}
@@
 
-To conclude, the recent all-atom force fields all capture the overall trends of CHOL--PC interaction reasonably well; The addition of CHOL leads to acyl chain ordering, which goes hand in hand with the increase in bilayer thickness.
-Area per lipid increases less than what would be expected if the areas of POPC and CHOL were additive, manifesting a condensation effect of CHOL.
+The tested all-atom force fields captured the most important general effects of cholesterol on membrane properties: increased acyl chain ordering, concomitant increase in bilayer thickness, and reduced lateral diffusion rate of lipids.
+However, quantitative comparison reveals differences between force fields and their qualities evaluated against NMR and X-ray scattering data.
+Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results up to 20~mol-\% of cholesterol, but overestimate acyl chain ordering and membrane thickness after further addition of cholesterol.
+Apparent exception to this are Lipid17 simulations, yet their seemingly better agreement with experiments can be explained by underestimated acyl chain order with low cholesterol concentrations that approaches experiments upon addition of cholesterol and finally overshoots the experimental results.
+In conclusion, unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the cholesterol ordering effect, particularly above 20~mol-\% of cholesterol.
+Previously published comparison~\cite{ferreira13} suggests the same conclusion for historically relevant Berger/H{\"o}ltje force field parameters that were not included in this work.
+
+Besides the ordering effect, effect of cholesterol on membrane properties has been discussed also in terms of the cholesterol condensing effect which 
+%The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}.
+The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
+However, condensation could also
+refers to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol (negative partial area) \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
+%Area per lipid increases less than what would be expected if the areas of POPC and CHOL were additive, manifesting a condensation effect of CHOL.
 However, beyond these qualitative features, the force fields behave very differently.
-At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, CHOL has a positive partial area in Slipids and Lipid17, a zero partial area in CHARMM36, and a negative partial area in MacRog.
-In addition to these trends, the absolute values of area per lipid and bilayer thickness vary substantially.
+At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, only MacRog predicts negative partial area for cholesterol, while CHARMM36 predicts zero partial area, and Slipids and Lipid17 predict small positive partial area.
+Considering that Slipids and Lipid17 perform best in our quality evaluation against experiments, yet still slightly overestimating the cholesterol condensing effect, our results suggests that cholesterol has positive but small partial area.
+%In addition to these trends, the absolute values of area per lipid and bilayer thickness vary substantially.
 At larger CHOL concentrations all force fields predict a too thick membrane, indicating an excessive ordering effect of CHOL.
 This is indeed verified by the order parameter maps, which suggest that the palmitate chain is overly ordered in all force fields to a varying degree.
-The best agreement with experiment is achieved with Slipids at low CHOL concentration and with Lipid17 at high CHOL concentration.
+
+%The best agreement with experiment is achieved with Slipids at low CHOL concentration and with Lipid17 at high CHOL concentration.
 In terms of dynamics, the lateral diffusion in pure POPC is too fast in Lipid17 and CHARMM36, and the addition of CHOL slows them both down more than in experiments.
@@
 
-None of the force fields unambigously provides the best agreement with experiment.
-Slipids shows great description of lateral dynamics, and reasonable behavior in terms of membrane thickness and acyl chain ordering, especially at lower and intermediate CHOL concentrations.
+%None of the force fields unambigously provides the best agreement with experiment. 
+Considering also lateral dynamics, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid-cholesterol interactions play major role.
+%and reasonable behavior in terms of membrane thickness and acyl chain ordering, especially at lower and intermediate CHOL concentrations.
 On the other hand, Lipid17 provides a good description for the palmitate chain ordering, but is otherwise in a relatively poor agreement with experiment except for higher CHOL concentrations.
@@
 At low concentrations, Slipids is a safe choice, but at higher concentrations Lipid17 might provide the most realistic description.
-Since many research questions will involve other lipids, proteins, sugars, or drugs, so the quality of the force fields describing these additional molecules might be the decisive factor, together with force field compatibility.
+Nevertheless, all tested parameters capture qualitative effects of cholesterol on membrane properties relatively well and differences between force fields are clearly smaller than, for example, in the case of PC-PS lipid mixtures~\cite{antila2022emerging}.
+Therefore, quality of selected force field for other molecules, such as,
+%Since many research questions will involve 
+other lipids, proteins, sugars, or drugs, 
+%so the quality of the force fields describing these additional molecules 
+together with force field compatibility might be more relevant decisive factor for simulations of complex systems.
 Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behavior, yet we believe that evaluating the magnitude of these effects is well beyond the scope of this work.
+
+Our results demonstrate that quality evaluation of lipid mixture simulations against experimental NMR and X-ray scattering data give consistent results for how accurately force field parameters capture intermolecular interactions.
+The interpolation approach introduced here extends the NMRlipids databank quality metrics~\cite{NMRlipidsDatabank} beyond individual systems enabling automatic ranking of not only lipid mixtures but also membrane mixed with other molecules such as ions.
+Such tools will strongly support the emerging endeavours for automatic improvements of force field parameters~\cite{antila2022emerging}.
 

### 2023-04-13 MANUSCRIPT/manuscript.tex
Update text

@@
 While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations are widely used to provide a detailed view of complex membranes and the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}.
-This is further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable simulations of complex membranes approaching realistic membrane mixtures.
+This is further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable the simulations of ever more complex and realistic membranes .
 

### 2023-05-02 MANUSCRIPT/manuscriptSUPPL.tex
Update manuscript

@@
 
-\author{Fernando Favela-Rosales}
-\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
+% \author{Fernando Favela-Rosales}
+% \affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
 %
@@
 \alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\email{matti.javanainen@helsinki.fi}
 %
-\author{Jesper J. Madsen}
-\affiliation{Department of Global Health, College of Public Health}
-\alsoaffiliation{University of South Florida}
+% \author{Jesper J. Madsen}
+% \affiliation{Department of Global Health, College of Public Health}
+% \alsoaffiliation{University of South Florida}
 %
-\author{Josef Melcr}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
+% \author{Josef Melcr}
+% \affiliation{Institute of Organic Chemistry and Biochemistry,
+% Academy of Sciences of the Czech Republic, 
+% Prague 6, Czech Republic}
 %
 \author{Markus Miettinen}
-\affiliation{MPI}
+\affiliation{Department of Chemistry, University of Bergen, Norway}
+\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
+
 %
 \author{O. H. Samuli Ollila}
-\email{samuli.ollila@helsinki.fi}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 \alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\email{samuli.ollila@helsinki.fi}
 %
@@
 %
-\author{Thomas Piggot}
-\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
-\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
+% \author{Thomas Piggot}
+% \affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
+% \alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
 
@@
     Neighbour list type & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} \\
-    Long-range electrostatics & PME & PME & PME & PME \\
+    Long-range electrostatics & PME \cite{essman95,darden93} & PME \cite{essman95,darden93} & PME \cite{essman95,darden93} & PME \cite{essman95,darden93} \\
     LJ cut-off & 1.2~nm & 1.4~nm & 0.9~nm & 1.0~nm \\
@@
     Water model & TIPS3P \cite{durell1994solvent}  & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} \\
-    FF Source & CHARMM-GUI & CHARMM-GUI & CHARMM-GUI & Refs.~\citenum{kulig2016cis} \& \citenum{macrogpopc} \\
-    Parameter source & CHARMM-GUI & Slipids website & CHARMM-GUI & Refs.~\citenum{kulig2016cis} \& \citenum{macrogpopc} \\
+    FF Source & CHARMM-GUI & CHARMM-GUI & CHARMM-GUI & Refs.~\citenum{Kulig16} \& \citenum{macrogpopc} \\
+    Parameter source & CHARMM-GUI & Slipids website & CHARMM-GUI & Refs.~\citenum{Kulig16} \& \citenum{macrogpopc} \\
     \bottomrule
@@
     \centering
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt0.pdf}
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt5.pdf}\\
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt10.pdf}
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt15.pdf}\\
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt20.pdf}
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt25.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt0.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt5.pdf}\\
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt10.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt15.pdf}\\
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt20.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt25.pdf}
     \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the next figure.}
@@
 
+\clearpage
+
 \begin{figure}[htb!]
     \centering
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt30.pdf}
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt35.pdf}\\
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt40.pdf}
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt45.pdf}\\
-    \includegraphics[width=0.49\linewidth]{FIGS/scatt50.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt30.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt35.pdf}\\
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt40.pdf}
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt45.pdf}\\
+    \includegraphics[width=0.45\linewidth]{../FIGS/scatt50.pdf}
     \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the previous figure.}
@@
 \end{figure}
+
+\clearpage
 
@@
 The minima are marked by filled circles to guide the eye.
-    }
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=8.7cm]{../FIGS/ffminima.pdf}
-    \caption{\label{SIfig:ffminima}%
-    \textbf{Effecf of CHOL on the location of the first two minima in the form factor.}
-    The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
     }
@@
     \textbf{Dependence of lateral diffusion coefficients on simulation box size.} The values calculated for the lipid centres of mass with \texttt{gmx msd} after eliminating leaflet drift.
-The values for the three system sizes are shown as markers together with fits of Eq.~\eqref{eq:pbc}.
+The values for the three system sizes are shown as markers together with fits of Eq.~\eqref{eq:sdlatpbc}.
 The values extrapolated to infinite system sizes are also shown in the separate column.
@@
 \clearpage
-\subsection{Cholesterol Tilt}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=8.7cm]{../FIGS/choltilt.pdf}
-    \caption{\label{SIfig:choltilt}%
-     \textbf{Effect of cholesterol concentration on the tilt of the cholesterol molecules.}
-     The tilt is defined as the angle between the vector connecting the C3 and C17 carbons at the two ends of the rigid ring structure of cholesterol.
-The hydroxyl group is bound to C3 and the hydrocarbon chain to C17.
-The tilt distributions were gathered from the two leaflets, fitted with a Gamma distribution, and the mean values and standard deviations were extracted and shown here as the value and error estimate, respectively.
-The cholesterol concentration was identical between the simulations, but the points are horizontally slightly shifted for clarity.
-    }
-\end{figure}
-
-\clearpage
 
@@
 \end{document}
-
-\section{CHARMM36 results from different simulation packages} \label{CHARMMprograms}
-The results from CHARMM36 model for lipid bilayers from different 
-simulation packages have been reported to give different results in
-the literature \cite{piggot12,lee16}. The results are mainly
-dependent on different Lennard-Jones cut-off settings, but
-all the details are not quite understood. In this work we use
-the results from Gromacs 5 with settings suggested to be optimal
-by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
-these settings to the results simulated with NAMD, OpenMM and literature
-values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
-that Gromacs 5 with settings suggested in webpage gives consistent
-results with the literature and other simulation packages. Thus these
-results are used in the main body of the paper. However, order parameters
-are slightly overestimated respected to the experiments also with these
-settings. \\
-\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
- \begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}
-
-  \caption{\label{programsCOMP}
-    Results for CHARMM36 model \cite{klauda10} from different simulation packages.
-    Discussion going on at {\it https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}.
-  }
-\end{figure}
-
-\clearpage
-\section{Order parameter slopes as a function of cholesterol}
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=19cm]{../FIGS/slopesEXPERIMENT.pdf}
-  \caption{\label{slopes}
-    Slopes of order parameters as a function of cholesterol from experiments \cite{ferreira13}.
-  }
-\end{figure*}
-
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=19cm]{../FIGS/slopesBERGER.pdf}
-  \caption{\label{slopesberger}
-    Slopes of order parameters as a function of cholesterol from Berger simulations.
-  }
-\end{figure*}
-
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=19cm]{../FIGS/slopesCHARMM.pdf}
-  \caption{\label{slopescharmm}
-    Slopes of order parameters as a function of cholesterol from CHARMM36 simulations.
-  }
-\end{figure*}
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=19cm]{../FIGS/slopesMACROG.pdf}
-  \caption{\label{slopesmacrog}
-    Slopes of order parameters as a function of cholesterol from MacRog simulations.
-  }
-\end{figure*}
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDS.pdf}
-  \caption{\label{slopesslipids}
-    Slopes of order parameters as a function of cholesterol from Slipids simulations.
-  }
-\end{figure*}
- \begin{figure*}[]
-  \centering
-  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDST310K.pdf}
-  \caption{\label{slopesslipids310K}
-    Slopes of order parameters as a function of cholesterol from Slipids simulations at 298 K.
-  }
-\end{figure*}
-
-\clearpage
- \section{Effect of undulations on order parameters} \label{undulations}
- \todo{This should be written based on contributions in this issue:
-   https://github.com/NMRLipids/NmrLipidsCholXray/issues/16
- }
-
-% \begin{figure}[]
-%  \centering
-%  \includegraphics[width=8cm]{../FIGS/slopesBERGER.pdf}
-%  \caption{\label{programsCOMP}
-%    Slopes of order parameters as a function of cholesterol from Berger simulations.
-%  }
-%\end{figure}

### 2023-05-02 MANUSCRIPT/manuscript.tex

@@
 \usepackage{multirow}
-% \usepackage{rotating} 
-% \usepackage{times}
-% \usepackage{graphicx}
-% \usepackage{setspace}
-% \usepackage{amsmath}
-% \usepackage{epstopdf}
-\usepackage[obeyFinal]{easy-todo}
 
@@
 \author{Markus Miettinen}
-\affiliation{MPI}
+\affiliation{Department of Chemistry, University of Bergen, Norway}
+\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
+
 %
@@
 
-\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
+%\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
+\title{Quantitative Comparison With Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions}
 
@@
 
-\todo{Update Author list \& Affiliations}
+\begin{abstract}
+Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, renders the membrane stiffer, and drives domain formation.
+Molecular dynamics simulations have played a crucial role in resolving these effects at molecular level, yet it has recently become evident that different simulation force fields predict quantitatively different behavior.
+Although easily neglected, identifying such limitations are becoming increasingly important as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions where it is crucial to accurately model the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol.
+%
+Here, we define quantitative quality measures for binary lipid mixtures in membranes against C--H bond order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering.
+We then perform a systematic evaluation on how well the commonly used force fields describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
+None of the tested force fields clearly outperforms the others across the tested properties and conditions, yet Slipids parameters provide the best overall performance.
 
-\begin{abstract}
-Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, and thereby renders the membrane stiffer.
-Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models predict quantitatively different behavior.
-Such limitations are easily neglected, as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions.
-Still, for their validity, it is crucial that the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol, are modelled accurately.
-%
-Here, we perform a systematic comparison of the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
-Comparison of lipid bilayer simulations with deuterium order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations.
-The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively.
-%The performance of the commonly used CHARMM36 model is subpar, as it overestimates diffusion at low cholesterol concentrations, and gets excessively ordered upon the addition of cholesterol. 
-The accompanied X-ray scattering data set measured for this work will foster future force field development and refinement.
+The quality evaluation metrics introduced in this work will particularly foster the future force field development and refinement for multi-component membranes with automated approaches.
 \end{abstract}
@@
 
-\todo{Check that the uses of ``model'' and ``force field'' are consistent}
-
 Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
@@
 CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}.
-In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}.
-This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation}, %\textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic},
-in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.
+In the cellular setting, the interaction between other lipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}.
+This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation}, in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane,manna2016mechanism}.
 
-While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations are widely used to provide a detailed view of complex membranes and the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}.
-This is further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable the simulations of ever more complex and realistic membranes .
+While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations have been employed to obtain a detailed view on the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}.
+Such efforts are further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable the simulations of ever more complex and thus realistic membranes \cite{lorent2020plasma}.
 
-The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively.
-Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines that enables rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.
+The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the force field by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig16}, respectively.
+Notably, simulations using CHARMM36, Lipid17, and Slipids can now be set up using CHARMM-GUI for multiple simulation engines, which enables the rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.
 
-While simulating of complex membranes with cholesterol have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge.
-%relies on the quality of the used model. 
-%Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons.
-Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature.
-This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
-%
-Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C-H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}.
-%This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. 
-This approach has been useful for finding the best models for headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE) \cite{bacle2021inverse} and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}.
-Furthermore, quantitative quality measures based on C-H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}.
-However, such automatic quality evaluation is limited to simulations that have experimental data available with corresponding composition and temperature.
-%One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and 
-Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automized in the NMRlipids databank.
+While simulations of complex membranes with CHOL have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge.
+Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C--H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}.
+This approach has been useful for finding the best force fields for the description of the headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE)~\cite{bacle2021inverse} and phosphatidylglycerol (PG)~lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}.
+Furthermore, quantitative quality measures based on C--H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}.
+However, such automatic quality evaluation is limited to simulations for which experimental data at a corresponding composition and temperature are available.
+Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automatized in the NMRlipids databank.
 
-Here, we demonstrate how simulations of binary POPC-cholesterol mixtures can be evaluated against experimental NMR and X-ray scattering data
-%order parameters and electron density profiles can be readily 
-by interpolating through multiple CHOL concentrations.
-Because effect of cholesterol on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously~\cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in cholesterol induced lateral membrane heterogeneity.
-%to intermediate CHOL concentrations.
-This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison.
-We also evaluated 
-%whether the simulation force fields reproduce 
-the dependence of lateral diffusion coefficients on cholesterol against pulsed field gradient (PFG) NMR experiments~\cite{filippov2003effect,filippov2003influence} using the recent theoretical work that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}.
+Here, we demonstrate how simulations of binary POPC--CHOL mixtures can be evaluated against experimental NMR and X-ray scattering data by interpolating through multiple CHOL concentrations.
+Because effect of CHOL on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously \cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in CHOL induced lateral membrane heterogeneity.
+We also evaluated the dependence of lateral diffusion coefficients on CHOL against pulsed field gradient (PFG) NM experiments \cite{filippov2003effect,filippov2003influence} using the recent theoretical framework that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}.
 With the structural and dynamic comparisons established, we estimated the quality of popular force fields at different CHOL concentrations.
-We then used the best-performing force fields to analyze molecular organization at different cholesterol concentrations, as we expect them to best reflect this organization in reality.
-%
-%Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby 
-While we focus here on POPC-cholesterol mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. 
-%Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.
-%Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}.
-In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer.
-The form factor and electron density profiles can be compared with the same profiles calculated from the simulation data.
-Whereas NMR provides data on the conformations of individual lipids, form factors from scattering captures the details of the structure of the entire lipid bilayer \cite{ollila16}.
+While we focus here on POPC-CHOL mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems.
 
@@
 \subsection{X-ray Scattering Experiments}
-Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high} .
-This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content.
+
+Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high}.
+This avoids the precipitation of CHOL crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% CHOL content.
 Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification.
@@
 The electron density profiles were modelled from form factors using the SDP model where volume distribution functions are modelled by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}.
-%CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}.
-Cholesterol was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
-%Additional figures show the volume distribution functions and the resulting electron density profiles.
+CHOL was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}.
 Membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
@@
 We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%.
-Systems were simulated using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models.
+Systems were simulated using four commonly used force fields, namely CHARMM36 (often abbreviated ``C36'' in figure legends in this work) \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig16} force fields.
 In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256 or 1024 POPC molecules).
-%The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules).
 The number of POPC molecules was kept constant across the different CHOL concentrations.
 All membranes were solvated by 50 waters per lipid (POPC or CHOL).
+The small membranes (with 64 lipids) were first generated using CHARMM-GUI and equilibrated using the standard protocols for CHARMM36, Slipids, and Lipid17 for which inputs are readily available from CHARMM-GUI \cite{lee16,lee2020charmm}.
+Then, the atomic coordinates were replicated in the membrane plane to create the 4- and 16-fold larger simulation systems.
+Since CHARMM-GUI does not support MacRog, the production simulations were initiated from equilibrated CHARMM36 structures, since they share the atom ordering.
 All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}.
+The first 10~ns of each simulation were omitted from analyses.
 The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
 All simulations were stored to the NMRlipids databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}.
-
-\todo{Run Lipid21 for revision?}
+The CHARMM36 simulations have been previously analyzed for their dynamic properties in Ref.~\citenum{fabian2023protein}.
 
@@
 
-\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles 
-%were calculated using the \texttt{form\_factor.py} script in 
-automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank} 
-%(\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts. 
-were used.
+\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles, automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank}, were used.
 Similarly to the experimental X-ray scattering data, membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
@@
 Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.
-%
-%\paragraph{Order Parameters:} The C--H order parameters were calculated as
-%%
-%\begin{align}
-%    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
-%\end{align}
-%%
-%where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane. 
-To simplify the interpolation, C-H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged.
+
+To simplify the interpolation of order parameter data to intermediate CHOL concentrations (see \textit{``Quantitative quality evaluation of the effect of CHOL on membrane properties''} below), C--H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged.
 These groups rotate freely and thus the order parameters are essentially identical for both (all) hydrogens in experiments and simulations.
-%, the values are also very similar, justifying this approach. 
-An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses.
-The C--H order parameters for the POPC head group
-%were calculated similarly, but values for both hydrogens 
-are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.
+An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both simulations and experiments, and were thus treated separately in our analyses.
+The C--H order parameters for the POPC head group are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.
 
-\paragraph{Lateral Diffusion Coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet with the \texttt{gmx msd} tool.
+\paragraph{Lateral Diffusion Coefficients:} Lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet with the \texttt{gmx msd} tool.
 The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
@@
 
-The diffusion coefficients extracted from the three simulation box sizes were fitted with
+The diffusion coefficients extracted from the three simulation box sizes (and thus membrane edge lengths $L$) were fitted with
 %
@@
 %
-where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $H$ is half the thickness of the water layer, $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length, and $\mu_\mathrm{m}$ and $\mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively~\cite{vogele2018hydrodynamics}.
+where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}$ is the Boltzmann constant, $T$ is the temperature, $H$ is half the thickness of the water layer, $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffman--Delbr\"{u}ck length, and $\mu_\mathrm{m}$ and $\mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively~\cite{vogele2018hydrodynamics}.
+Thus, the fit of Eq.~\eqref{eq:sdlatpbc} to the $D_\mathrm{PBC}$ values calculated from the simulation as a function of simulation box size $L$ has two free parameters, namely $D_\infty$ and $\mu_\mathrm{m}$, both of which are free of finite size effects and can thus be compared to experiment.
 The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}.
-The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).
-The final simulation box dimensions were used for $L$ and $H$.
-For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.
+The water viscosity value of $\mu_\mathrm{f}= 0.3228$~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).
+The simulation box dimensions, membrane edge length $L$ and the dimension normal to the membrane plane ($z$) $L_z$, were taken from the final configuration of each simulation.
+$L_z$ was only needed in the calculation of $H$ (see below).
+Membrane thickness $h$ was obtained as described in \textit{``Structural properties''} above.
+As $h$ and $H=(L_z-h)/2$ are constants in Eq.~\eqref{eq:sdlatpbc}, the average values from the three systems sizes were used in the fits.
 
-\paragraph{Quantitative Quality evaluation of cholesterol effect on membrane properties:} To ease the evaluation of simulations against experimental data with non-matching concentrations, we interpolated the effect of cholesterol in simulations and experiments for cholesterol concentrations ranging from 0\% to 50\%.
-%The density profiles (as a function of distance from membrane core along the membrane normal), 
-Continuous 2D matrices were created by interpolation for the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and the electron density profiles (as a function of normal distance from bilayer center), while 1D interpolation was used for the cholesterol dependence of the 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, and diffusion coefficients.
+\paragraph{Quantitative quality evaluation of the effect of CHOL on membrane properties:} To ease the evaluation of simulations against experimental data with non-matching CHOL concentrations, we interpolated the effect of CHOL in simulations and experiments for CHOL concentrations ranging from 0\% to 46\%.
+2D matrices were created by interpolation for the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains and CHOL concentration) using the \texttt{interp2} function in Matlab based on linear interpolation.
+Similar matrices were also generated for the electron density profiles (as a function of normal distance from bilayer center and CHOL concentration) but with the \texttt{scatteredInterpolant} function in Matlab based on Delanuay triangulation \cite{amidror2002scattered}.
+Linear 1D interpolation with the \texttt{interp1} function in Matlab was used for the CHOL-dependence of the 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations and diffusion coefficients.
 These interpolations were then used to calculate deviations (in \%) from experimental values across CHOL concentrations to quantify the quality of the lipid force fields.
 For the 2D matrices, the absolute values of the differences between matrices from simulations and experiments were first calculated.
-The averages of differences over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles) were then calculated.
+The averages of differences over the carbon atoms in the acyl chain (order parameters) or across the entire $z$ coordinate in the simulation (density profiles) were then calculated.
 This resulted in 1D deviation vectors as a function of CHOL concentration.
 For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
-All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of cholesterol concentration.
-For C-H bond order parameters, also the deviation matrices between simulations and experiments were used to illustrate quality of simulations.
-
-\paragraph{More and Less Ordered Populations:} The angles ($\theta$) between the C--H vectors of the last six carbons of the palmitate chain were extracted and the data across all CHOL concentrations was accumulated.
-These data were fit with two Gaussians centered at 0,
-%
-\begin{align}
-    P(\theta)=a_\mathrm{less}\exp\left(-\frac{\theta}{2\sigma_\mathrm{less}}\right)+a_\mathrm{more}\exp\left(-\frac{\theta}{2\sigma_\mathrm{more}}\right),
-    \label{eq:twogauss}
-\end{align}
-%
-where the two standard deviations ($\sigma$) correspond to the more and less ordered components.
-After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations were independently fitted by Eq.~\eqref{eq:twogauss} with the fixed $\sigma$ values.
-These fits provided the pre-factors, whose ratio $a_\mathrm{more}/(a_\mathrm{more}+a_\mathrm{less})$ corresponds to the fraction of the more ordered component in the system.
-
-This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not their natures.
-In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would decrease as a function of CHOL concentration as the membrane gets more ordered.
+All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of CHOL concentration.
+For C--H bond order parameters, also the deviation matrices between simulations and experiments were used to illustrate quality of simulations.
 
@@
 
-\todo{Results: volmaps of chains around cholesterol?}\\
-\todo{Results: SI: RDFs of POPC chains and CHOL around CHOL}\\
-\todo{Unify colours in figures}\\
-
 \subsection{Acyl Chain Ordering Varies Greatly Between the Force Fields}
 
-Cholesterol is known to induce order in cell membranes by increasing the fraction of \textit{anti} conformations in the acyl chains of phospholipids \cite{ferreira13}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}.
-Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.
-Cholesterol ordering effect can be experimentally quantified by measuring the C-H bond order parameters using $^{13}$C or $^{2}$H NMR techniques that can be also directly compared with MD simulations~\cite{ollila16}.
-Simulations generally reproduce the cholesterol ordering effect, but order parameters often deviate from experiments at high cholesterol concentrations~\cite{ferreira13,madej15}.
-Furthermore, it has not been clear how accurately different force field parameters capture the details of lipid-cholesterol interactions and which parameters would give most realistic results for simulations of complex mixtures where such interactions play critical roles.
+CHOL is known to induce order in lipid membranes by increasing the fraction of \textit{anti} conformations in the acyl chains of phospholipids \cite{ferreira13}, which is suggested to play critical role in the phase behaviour of PC--CHOL mixtures \cite{ipsen87}.
+Consequently, the correct CHOL ordering effect is expected to be a necessary condition for a force field used to understand lipid--CHOL phase behaviour.
+CHOL ordering effect can be experimentally quantified by measuring the C--H bond order parameters using $^{13}$C or $^{2}$H NMR techniques, and the result can be also directly compared to MD simulations~\cite{ollila16}.
+Simulations generally reproduce the CHOL ordering effect, but order parameters often deviate from experiments at high CHOL concentrations~\cite{ferreira13,madej15}.
+Furthermore, it has not been clear how accurately different force field parameters capture the details of lipid--CHOL interactions and which force field would give most realistic results for simulations of complex mixtures where such interactions play critical roles.
 
-Here, we evaluate the cholesterol ordering effect in state of the art force field parameters against C-H bond order parameter data from $^{13}$C NMR experiments measured from POPC-cholesterol mixture with cholesterol concentratons ranging between 0-60\%~\cite{ferreira13}.
-To this end, we first interpolated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments.
+Here, we evaluate the CHOL ordering effect in state-of-the-art force field parameters against C--H bond order parameter data from $^{13}$C NMR experiments measured from POPC--CHOL mixture with CHOL concentratons ranging between 0 and 60~mol-\%~\cite{ferreira13}.
+To this end, we first interpolated order parameter maps as a function of acyl chain carbon number and CHOL concentration for both simulations and experiments.
 These maps were then subtracted to obtain the deviation maps between simulations and experiments.
-The deviation maps of different force fields are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate}.
+The deviation maps of different force fields are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} for the palmitate and oleate chains, respectively.
 
-The cholesterol ordering effect is manifested in original profiles in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} as substantial increase in absolute values of acyl chain C--H bond order parameters upon addition of cholesterol in all simulation models and experiments.
-%, which can be explained by increased \textit{anti} conformations in acyl chains \cite{ferreira13,??}. 
-%The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13} exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig.
-\ref{slopes}), which is expected because phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
+The CHOL ordering effect is manifested in original profiles in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} as substantial increase in absolute values of acyl chain C--H bond order parameters upon addition of CHOL in all simulation force fields and experiments.
 The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive view for a quantitative comparison of different force fields with experiments.
 In white regions, the simulation results are considered to fall within the experimental error as the deviations are in the range of [--0.02,0.02].
-Blue indicates that the order parameters are too negative, i.e., the acyl chains are too ordered in simulations, and \textit{vice versa} for red.
-Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
+Blue indicates that the order parameters are too negative, \textit{i.e.}, the acyl chains are too ordered in simulations, and \textit{vice versa} for red.
+Overall, the simulation force fields behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
 
@@
 Negative values indicate that order is too high ($-S_\mathrm{CH}$ values too negative) in the simulations.
-The values that are within the estimated experimental error range of 0.02~\cite{ollila16} are
-  %considered to be ``correct'' and 
-  coloured in white.
-Order parameters of identical hydrogens attached to the same carbons were averaged with the exception of C2 carbon of the oleate chain with the forked order parameters for which differences for both the larger and smaller values were calculated, and the average of this deviation is shown in the C2 column of oleate.
+The values that are within the estimated experimental error range of 0.02~\cite{ollila16} are coloured in white.
+Order parameters of identical hydrogens attached to the same carbons were averaged with the exception of C2 carbon of the oleate chain with the forked order parameters for which differences for both the larger and smaller values were calculated, and the average of this deviation is shown in the C2 column.
   }
@@
 
-In CHARMM36, the both the palmitate and oleate chains get too ordered upon increasing CHOL concentration, yet the oleate C9 carbon next to the double bond is too disordered in the simulation.
-The oleate chain shows best agreement with experiment in Slipids with only slight excess ordering at higher CHOL concentrations.
-The excess ordering is more severe for the palmitate chain.
+In CHARMM36, both the palmitate and oleate chains get too ordered upon increasing CHOL concentration.
+The oleate chain shows best agreement with experiment in Slipids, whereas there is some excess ordering in the palmitate chain.
 Still, the major discrepancy between Slipids and experiment is the drastically too disordered C2 and C3 carbons of the palmitate chain.
-This effect was not present in the original Slipids model \cite{NMRlipidsDatabank}, indicating that it was introduced in the recent reparametrization that improved the head group and glycerol backbone structures of Slipids \cite{grote2020optimization}.
+This effect was introduced in the recent reparametrization of Slipids that improved the head group and glycerol backbone structures of Slipids \cite{grote2020optimization}.
 Lipid17 provides the best overall agreement with experiment, as no segments deviate significantly from experiment at any CHOL concentrations.
-At low CHOL concentrations, the middle of the oleate chain is too disordered.
-%, which agrees with the large APL observed in Fig.~\ref{fig:densmaps}. 
-At large CHOL concentrations, the oleate chain agrees well with experiment, whereas the palmitate chain has a slightly too ordered C9--C12 segment, yet the deviation is significantly smaller than in the other force fields.
-MacRog behaves reasonably well at low CHOL concentrations, yet the C2--C5 carbons in palmitate are slightly too disordered, whereas the C2 of oleate is somewhat too ordered.
-However, at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.
+MacRog behaves reasonably well at low CHOL concentrations, yet at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.
+For the head group and glycerol backbone order parameters, provided in Fig.~\ref{SIfig:headgroups}, CHARMM36 give the best agreement at all cholesterol concentrations in line with previous studies~\cite{botan15,antila2019headgroup}.
 
-%The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}.
-The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL \cite{botan15} largely persist across all studied CHOL concentrations.
-Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.
+\subsection{Cholesterol Effect on Membrane Properties is Manifested Differently in Different Force Fields}
 
-\todo{Cholesterol order parameters? Into the deviation metric?}
+CHOL-induced ordering straightens the acyl chain conformations that leads to the thickening of membranes.
+While acyl chain order and membrane thickness are well correlated~\cite{NMRlipidsDatabank}, lipid bilayer dimensions can be accessed more directly by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16}.
+Electron density profile, area per lipid and bilayer thickness can all be extracted from the form factor using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,doktorova2020molecular}.
+To complement the evaluation of CHOL ordering effect against NMR order parameters, we measured also X-ray scattering form factors from POPC--CHOL mixtures with systematically increasing CHOL concentrations.
+Scattering intensities from experiments are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, form factors from experiments and simulations in Fig.~\ref{SIfig:scattering}, and density profiles in Fig.~\ref{SIfig:densprofs}.
 
-\subsection{Cholesterol Effect on membrane properties is Manifested Differently in Different Force Fields}
+The effect of CHOL on structural properties of bilayers is compared between the SDP model (based on experimental form factors) and MD simulation results in Fig.~\ref{fig:densmaps}.
+All force fields demonstrate increasing thickness upon addition of CHOL that saturates after approximately 30~mol-\% (bottom middle panel in Fig.~\ref{fig:densmaps}).
+MD simulations agree well with the SDP model below 30~mol-\% but overshoot the SDP results at high CHOL concentrations.
+Lipid17 simulations are exception as they predict thinner membranes than SDP model at low CHOL concentrations, and clear saturation of thickness increase is not observed, unlike for other force fields and experiments.
 
-Cholesterol induced ordering straigthens the acyl chain conformations that leads to thickening of membranes.
-While acyl chain order and membrane thickness are well correlated~\cite{NMRlipidsDatabank}, lipid bilayer dimensions can be accessed more directly by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16}.
-Electron density profile, area per lipid and bilayer thickness can be extracted from the form factor using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,doktorova2020molecular}.
-%Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.
-To complement the evaluation of cholesterol ordering effect against NMR order parameters, we measured also X-ray scattering form factors from POPC-cholesterol mixtures with systematically increasing cholesterol concentrations.
-%
-%Thus, we compare here the form factors and the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species or their segments.
-For reference,
-Scattering intensitities from experiments are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, form factors from experiments and simulations in Fig.~\ref{SIfig:scattering}, and density profiles in Fig.~\ref{SIfig:densprofs}.
-
-While the absolute and relative lobe amplitudes depend on simulation box size, minima in form factors correlate with membrane dimensions~\cite{NMRlipidsDatabank}.
-Thus, we interpolated the locations of first two minima in the form factors to the entire studied range of CHOL concentrations in Fig.~\ref{SIfig:ffminima}.
-The plots in Fig.~\ref{SIfig:ffminima} highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields.
-CHARMM36 seems to be off more than the other three force fields.
-Above $\sim$25~mol-\% of CHOL, the location of the minimum shifts to larger wave vector values in the experiment, which is curiously not captured by any of the force fields.
-The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields.
-Slipids and Lipid17 are generally in better agreement with the experiment than MacRog and CHARMM36.
-
-Effect of cholesterol on structural properties of bilayers is compared between the SDP model (based on experimental form factors) and MD simulation results in Fig.~\ref{fig:densmaps}.
-All models demonstrate increasing thickness upon addition of cholesterol that saturates after approximately 30~mol-\% (bottom middle panel in Fig.~\ref{fig:densmaps}).
-MD simulations agree well with the SDP model below 30~mol-\% but overshoot the SDP results at high cholesterol concentrations.
-Lipid17 simulations are exception as they predict thinner membranes than SDP model at low cholesterol concentrations, and clear saturation of thickness increase is not observed.
-%
-%The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}.
-The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
-However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without affecting its total area \cite{javanainen2017two}.
-We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}.
-%
 The dependence of APL on CHOL concentration follows the trends in thickness inversely in general (bottom right panel of Fig.~\ref{fig:densmaps}), yet provide curious differences between the force fields at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
 Lipid17 has the largest APL across the entire CHOL contration range.
-%, and the addition of CHOL always leads to an increase in the membrane area. 
 MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect.
-The profiles for Slipids and CHARMM36 are very similar with the small or zero cholesterol partial area until a concentration of 20~mol-\%.
-%Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained for POPC by scattering and NMR experiments performed by different teams (See Fig.~\ref{fig:densmaps} and its caption).
-Still, all the simulation force fields provide areas per lipid that fall within the experimental spread for pure POPC.
+The profiles for Slipids and CHARMM36 are very similar with the small or zero CHOL partial area until a concentration of 20~mol-\%.
 
@@
   %
-  Bottom left: Electron density map for the experiment.
+  Bottom left: Electron density map for the experiment from the SDP model.
 The scaling is uniform between all simulations and experiment.
@@
   %
-  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids.
-The experimental estimates, interpolated to 298~K, are taken from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$).
-The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR.
-  %
-  The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
+  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids. The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
   }
@@
 
-For more detailed comparison of membrane structure, 
-%Moving on to the electron density profiles extracted from the form factors, 
-we interpolated the changes in electron density profiles along membrane normal (Fig.~\ref{SIfig:densprofs}) as a function of cholesterol concentration to create two-dimensional electron density maps shown in Fig.~\ref{fig:densmaps}.
-Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region containing electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
-However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields.
-%CHOL renders the membrane thicker, resulting in the spreading of the high-density bands with increasing amount of CHOL.
-In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL.
-In Lipid17, however, the membrane gets thicker until the maximum studied CHOL concentration of 47~mol-\%.
-In experiment, the effect of CHOL is generally weaker, and some saturation is evident at high CHOL concentrations.
-These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness extracted from the density peak positions.
-Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating smaller membrane fluctuations that would smear the bands, and the same is true for MacRog at higher cholesterol concentrations.
+For more detailed comparison of membrane structure, we interpolated the changes in electron density profiles along membrane normal (Fig.~\ref{SIfig:densprofs}) as a function of CHOL concentration to create two-dimensional electron density maps shown in Fig.~\ref{fig:densmaps}.
+Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region containing electron-rich phosphorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
+However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields and SPD model.
+CHARMM36 has the sharpest low- and high-density bands among MD simulation results indicating smaller membrane fluctuations that would smear the bands, and the same is true for MacRog at higher CHOL concentrations.
 The less sharp bands for Lipid17 and especially Slipids profiles indicate that the membranes are more flexible with these force fields.
+SDP model gives the most sharp bands (bottom left panel in Fig.~\ref{fig:densmaps} and original electron density profiles in Fig.~\ref{SIfig:densprofs}), suggesting more uniform membrane structure than any MD simulation.
+However, System size plays a role here as  undulations smear out the density profile features with increasing simulation box size as demonstrated in Fig.~\ref{SIfig:densprofssize} by the density maps calculated for the CHARMM36 simulations in three sizes.
+Still, even in the smallest system, the band intensities are less localized than in the SDP model.
+This difference might arise because most undulations are assigned to structure factor rather than form factor in their deconvolution, but it could also indicate different elastic properties between simulation and experiment.
 
-Comparison against experiments (bottom left panel in Fig.~\ref{fig:densmaps}) indicates that the membrane structure is more uniform in the experiment, as the bands are sharper.
-This is also evident in the original electron density profiles in Fig.~\ref{SIfig:densprofs}.
-System size plays a role here, as demonstrated by the density maps calculated for the CHARMM36 simulations in three sizes, shown in Fig.~\ref{SIfig:densprofssize}.
-The larger the system, the larger the bilayer undulations are, which can smear the features of the density profiles.
-Still, even in the smallest system, the band intensities are not as localized as in the experiment.
-This difference likely partially stems from the features of the used SDP model, but it could also indicate different elastic properties between simulation and experiment.
+Nevertheless, minima in X-ray scattering form factors are independent on simulation box size and correlate with membrane dimensions~\cite{NMRlipidsDatabank}.
+For more direct comparison between simulations and experiments, we thus interpolated the locations of first two minima in the form factors to the entire studied range of CHOL concentrations in the bottom right panel of Fig.~\ref{fig:deviation}.
+These curves highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields.
+CHARMM36 seems to be off more than the other three force fields.
+Above $\sim$25~mol-\% of CHOL, the location of the minimum shifts to larger wave vector values in the experiment, which is curiously not captured by any of the force fields.
+The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields.
+Slipids and Lipid17 are generally in better agreement with the experiment than MacRog and CHARMM36.
 
@@
 The comparison between lateral diffusion coefficients extracted from simulation and experiment has been limited due to a box-size dependence observed in simulations performed using periodic boundary conditions \cite{vogele2016divergent,vogele2018hydrodynamics,camley2015strong}.
-Here, we tackle this issue by performing simulations with three system sizes, which allows the extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent} \cite{vogele2018hydrodynamics,vogele2018hydrodynamics}.
+Here, we tackle this issue by performing simulations with three system sizes, which allows the extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent} \cite{vogele2016divergent,vogele2018hydrodynamics}.
 The size-dependence of lipid lateral diffusion from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
@@
   \caption{\label{fig:dynamics}%
-  \textbf{Dynamic properties of the POPC/cholesterol mixtures.}
-  Top: Lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:dpbc}.
+  \textbf{Dynamic properties of the POPC/CHOL mixtures.}
+  Top: Lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:sdlatpbc}.
 Experimental data are taken from NMR measurements on well-hydrated samples \cite{filippov2003influence,filippov2003effect}.
@@
   %
-  Bottom: Shear viscosities obtained from the finite-size correction. Eq.~\eqref{eq:dpbc}.
+  Bottom: Shear viscosities obtained from the finite-size correction, Eq.~\eqref{eq:sdlatpbc}.
   %
@@
 
-The lipid models again show significantly different behavior.
-CHARMM36 and Lipid17 provide drastically too large diffusion coefficients at low CHOL concentrations, yet they converge towards experimental estimates upon the addition of CHOL.
-This similar behavior is somewhat surprising, as they behave very differently in terms of both area per POPC (Fig.~\ref{fig:densmaps}) and acyl chain ordering (Fig.~\ref{fig:OPmaps}).
-For CHARMM36, the excessive ordering observed at high CHOL concentrations counterbalances the too fast dynamics observed at low CHOL concentrations.
-However, the behavior of Lipid17 cannot be explained by the order parameters, as they agree with experiment relatively well across all studied CHOL concentrations.
-The area per POPC is exaggerated at all CHOL concentrations, yet the deviation from experiment decreases towards larger CHOL concentrations.
-Thus, it seems that CHARMM36 dynamics correlate with chain ordering, whereas those of Lipid17 go hand in hand with area per POPC.
-With MacRog, the dynamics of POPC are already slightly too slow compared to experiment, and the excessive ordering at large CHOL concentrations leads to a larger underestimation of the experimental values.
-Curiously, Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms other force fields in terms of lateral dynamics.
+The lipid force fields again show significantly different behavior.
+Lipid17 and CHARMM36 show too fast dynamics for pure POPC, and the slow-down induced by CHOL is exaggerated as compared to experiment.
+Diffusion in MacRog is generally too slow, yet Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms other force fields in terms of lateral dynamics.
+Interestingly, there is no correlation between dynamic and structural properties described earlier when it comes to the deviations from experiments.
+To investigate if the differences in lateral diffusion coefficients could be explained by different Lennard-Jones (LJ) cutoff values, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids using a shorter cutoff of 0.9~nm corresponding to that of Lipid17, while original cutoff for Slipids was 1.4~nm (see Table~\ref{SItab:params}).
+The PBC-corrected diffusion coefficient values with shorter cutoff, shown in the top panel of Fig.~\ref{fig:dynamics} as hollow red circles, are only slightly larger than the original values, indicating that the LJ cutoff alone does not explain the differences between force fields, but they truly spark from differences in the force field interaction parameters.
 
-Slipids has the largest Lennard-Jones (LJ) cutoff of the used force fields of 1.4~nm (see Table~\ref{SItab:params}), and this cutoff is known to affect dynamic properties \cite{leonard2018comparison}.
-Therefore, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids but with a cutoff of 0.9~nm corresponding to that of Lipid17, and the PBC-corrected diffusion coefficient values are shown in the top panel of Fig.~\ref{fig:dynamics} as hollow green circles.
-The decrease in cutoff indeed leads to a minor increase in the lateral diffusion coefficient.
-However, the LJ cutoff alone does not explain the differences between force fields, indicating that they truly spark from differences in the force field interaction parameters.
+When comparing lipid lateral diffusion between simulations and experiments, it is important to note that the PBC correction does not only affect values of diffusion coefficients but also qualitatively changes their trends as a function of cholesterol.
+This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration.
+For example, Fig.~\ref{SIfig:dvschol} shows that Lipid17 and CHARMM36 systemically underestimate the experimental values in systems with all sizes simulated here while the effect of CHOL seems to be qualitatively reproduced.
+However, the PBC corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations leading to values close to experimental ones at 47~mol-\% CHOL.
+With MacRog, also the PBC corrected values fall below the experimental ones, yet the slowdown effect of CHOL again seems stronger than in experiment.
+With Slipids, the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent.
+These results highlight that little can be said about CHOL-dependence of lateral diffusion coefficients without performing simulations with multiple box sizes required for the PBC correction at different CHOL concentrations.
+Therefore, fine-tuning of interaction parameters to correctly reproduce lateral diffusion coefficients can require massive number of simulations.
 
-As the results in Fig.~\ref{SIfig:dvschol} demonstrate, the PBC correction can qualitatively change the trends observed for diffusion coefficients.
-This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration.
-For example, Lipid17 and CHARMM36 underestimate the experimental values in all system sizes used here and the effect of CHOL seems to be well reproduced.
-However, the corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations so that the simulations and experiment almost agree at 47~mol-\% CHOL.
-With MacRog, even the corrected values fall below the experimental ones, yet the slowdown effect of CHOL again seems stronger than in experiment.
-Slipids is a clear outlier, as the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent.
-These results highlight that little can be said about CHOL-dependence of lateral diffusion coefficients without performing multiple simulations required for the PBC correction at multiple CHOL concentrations.
-When fine-tuning interaction parameters, this can lead to a massive number of simulations.
-
-The PBC correction, Eq.~\eqref{eq:sdlatpbc} also provides the shear viscosity of the membrane as a fitting parameter.
-Instead of shear viscosity, the surface viscosity is sometimes reported, yet this is simply the product of shear viscosity and membrane thickness (available in Fig.~\ref{fig:densmaps}).
-The shear viscosity values are in line with the lateral diffusion coefficients.
+We also provide the shear viscosity values of the membranes obtained from the fits of Eq.~\eqref{eq:sdlatpbc} in Fig.~\ref{fig:dynamics} for reference, yet they are not discussed further due to the large scatter of experimental estimates \cite{faizi2022vesicle}.
+Note that the data for CHARMM36 was also discussed in our previous work \cite{fabian2023protein}.
 
@@
 
-The previous sections reveal differences in predictions of 
-%highlight that the agreement of the studied 
-force fields for the effect of cholesterol on membrane properties. 
-%with experiment depends on both the parameter in question or the CHOL concentration. 
-%To capture these features in an easily digestible form, 
-To streamline the selection of force fields that best capture the effect of cholesterol on different membrane properties, we used the interpolated data to calculate the relative deviations from experiments (difference of simulated and experimental values divided by the experimental value) for the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients from all the force fields.
-%The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter.
+To streamline the selection of force fields that best capture the effect of CHOL on different membrane properties, we defined the quantitative quality measures for simulations.
+For this we calculated  relative deviations from experiments (difference of simulated and experimental values divided by the experimental value) using interpolated data for the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients from all the force fields.
+The quality evaluations are shown in Fig.~\ref{fig:deviation} for force fields that predicted different effects of CHOL on membrane properties in previous sections.
 
@@
   \caption{\label{fig:deviation}%
-  \textbf{Total deviation of the force fields from experimental data.} 
+  \textbf{Total relative deviation of the force fields from experimental data.} 
 %
@@
   The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
-The differences in the form factor minima in Fig.~\ref{fig:SIfig:ffminima} between experiment and simulation were calculated, normalized against experimental data, and summed together.
+The differences in the form factor minima---shown in the bottom right panel---between experiment and simulation were calculated, normalized against experimental data, and summed together.
 The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
@@
   Bottom right panel:
-  Deviation of the density profiles calculated from the form factor using the SDS model.
-The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
+  Effecf of CHOL on the location of the first two minima in the form factor.
+The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
   }
@@
 
-The deviations in Fig.~\ref{fig:deviation} 
-%The comparison of force fields in the two leftmost columns of Fig.~\ref{fig:deviation} 
-reveals that Slipids provides overall the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
-%Still, Slipids is the second best force field even in the high CHOL regime.
-Lipid17 provides the worst agreement with experiment at low CHOL concentrations, but improves significantly towards larger concentrations so that at
-Lipid17 provides slightly better agreement with experiment above $\sim$30~mol-\% of CHOL than Slipids, but exhibits major deviation from experiments with low cholesterol concentrations. 
-%CHARMM36 and MacRog never provide the best overall agreement with experiment. 
-%In fact,  all contributions to , and it shows the poorest agreement with the form factor minima. 
+The quality evaluations reveal that Slipids provides overall the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
+Lipid17 provides a slightly better agreement with experiment above $\sim$30~mol-\% of CHOL than Slipids, but exhibits major deviation from experiments at low CHOL concentrations.
 MacRog performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
-The deviation metric for CHARMM36 are significant with all cholesterol concentrations.
-%
+The deviation metric for CHARMM36 are significant at all CHOL concentrations.
+
 Since the contributions from the form factor minima deviations are barely visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
@@
 
-Finally, we also analyzed the deviations in the density profiles in the bottom right panel of Fig.~\ref{fig:deviation}.
-This analysis demonstrates that even the significant deviations as well as the trend as a function of CHOL concentration in the form factor minima are lost as form factors are converted to density profiles.
-This indicates that the form factor minima are a better target value for force field development and refinement.
-
-\subsection{Cholesterol Induces Order in a Non-collective Manner}
-
-CHOL could induce order in a lipid membrane in two ways; it could either increase the average order of all phospholipids (``collective transition''), or increase the prevalence of the ordered lipids (``two state model'').
-To clarify this effect, we extracted the C--H angle distributions with respect to the $xy$ plane for the last six carbons of the palmitate chain from all force fields.
-We first accumulated data at all CHOL concentrations, and observed that these data could be fitted with two Gaussian functions.
-These Gaussians were centered at 0$^\circ$, and with the different standard deviations they corresponding to the more and less ordered acyl chains.
-The obtained standard deviations for the more and less ordered chains only showed minor differences between the force fields (rightmost panel in Fig.~\ref{fig:chdist}).
-We then fitted the C--H angle distributions for all CHOL concentrations with the two Gaussians with the obtained standard deviations and with their means set to 0.
-These Gaussians provided provided excellent fits to the data with $R^2=0.991\pm 0.005$ (see the leftmost panel of Fig.~\ref{fig:chdist} for the fits to the Slipids data).
-The prefactors of the Gaussian functions were considered to represent the fractions of the more and less ordered components in the system.
-In pure POPC, the ordered component was absent, yet its prevalence increased linearly as a function of CHOL, after saturating at high CHOL concentrations (rightmost panel in Fig.~\ref{fig:chdist}).
-The success of the two-Gaussian fits in describing the data indicates that with all force fields the bilayer is composed of lipids ordered by CHOL and lipids corresponding to their state in pure POPC membrane, \textit{i.e.} corresponding to the ``two state model'', and CHOL simply changes the fraction of these two states present.
-Indeed, single Gaussian fits with the standard deviation as a free parameter fail when CHOL is present, indicating that the ordering does not occur \textit{via} a ``collective transition''.
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{../FIGS/CH.pdf}
-    \caption{\label{fig:chdist}%
-    \textbf{Distributions of C--H bond orientations.}
-    Leftmost panel: The probability distributions for the angle of C--H vectors of the last 6 carbons in the palmitate chain of POPC with respect to the $xy$-plane.
-All distributions are centered around 0, but have different shapes.
-The markers show data extracted from Slipids simulations at various CHOL concentrations, and the solid lines show two-Gaussian fits with fixed standard deviations to the data.
-These Gaussians correspond to the more and less ordered palmitate chains.
-The fixed standard deviations were extracted from a fit of two Gaussians to the data accumulated across all CHOL concentrations.
-The fit with the fixed standard deviations to all CHOL concentrations are excellent with $R^2=0.991\pm 0.005$ (mean $\pm$ standard deviation).
-    %
-    Middle panel: The standard deviations of the two Gaussians fitted to the C--H distribution (such as the one in the leftmost panel) for the four force fields. 
-    %
-    Rightmost panel: The fraction of the more ordered component as a function of CHOL concentration extracted simply by dividing the prefactor in the Gaussian distribution for the more ordered component by the sum of the two prefactors.
-Notably, in the CHOL-free system, the ordered component is not present.
-    }
-\end{figure}
-
-\subsection{Molecular Level Structure of PC--CHOL Mixtures}
-
-Molecular dynamics simulations provide a description of the molecular organization of POPC and CHOL in the bilayer, yet the reliability of this description depends on the quality of the used force field.
-In the previous section we established that Slipids and Lipid17 provide the best agreement with experimental properties at low and high CHOL concentrations, respectively.
-Thus, it is reasonable to assume that the molecular pictures provided by these two force fields at the corresponding CHOL concentration regimes are the most realistic.
-
-Earlier studies have analyzed the organization of CHOL and DSPC chains and observed five and three density peaks for DSPC chains and other CHOL molecules around CHOL molecules \cite{martinez2010cholesterol}.
-
-\textcolor{red}{2D density maps}
-\textcolor{red}{RDFs to SI?}
-
-Apart from the lateral organization of the molecules, their orientations might also differ between the force fields.
-However, we observed only small differences in the tilt of CHOL molecules, defined by the angle between a vector spanning the rigid ring structure as well as the $z$ axis (membrane lies in the $x,y$ plane).
-The mean angles and their standard deviations are shown in Fig.~\ref{SIfig:choltilt}.
-Lipid17 and Slipids show very consistent values.
-CHARMM36 has slightly smaller tilt angles than them across all CHOL concentrations.
-MacRog has the most tilted CHOL molecules at low CHOL concentrations, but their tilt angle is the smallest at large CHOL concentrations.
-Still, the differences are overall minor.
-
 \section{Conclusions}
 
-The tested all-atom force fields captured the most important general effects of cholesterol on membrane properties: increased acyl chain ordering, concomitant increase in bilayer thickness, and reduced lateral diffusion rate of lipids.
+The tested all-atom force fields captured the most important general effects of CHOL on membrane properties: increased acyl chain ordering, concomitant increase in bilayer thickness, and reduced lateral diffusion rate of lipids.
 However, quantitative comparison reveals differences between force fields and their qualities evaluated against NMR and X-ray scattering data.
-Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results up to 20~mol-\% of cholesterol, but overestimate acyl chain ordering and membrane thickness after further addition of cholesterol.
-Apparent exception to this are Lipid17 simulations, yet their seemingly better agreement with experiments can be explained by underestimated acyl chain order with low cholesterol concentrations that approaches experiments upon addition of cholesterol and finally overshoots the experimental results.
-In conclusion, unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the cholesterol ordering effect, particularly above 20~mol-\% of cholesterol.
-Previously published comparison~\cite{ferreira13} suggests the same conclusion for historically relevant Berger/H{\"o}ltje force field parameters that were not included in this work.
+Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results up to 20~mol-\% of CHOL, but overestimate acyl chain ordering and membrane thickness after further addition of CHOL.
+An apparent exception to this is the Lipid17 force field, yet their seemingly better agreement with experiments can be explained by the over-correction of the originally underestimated acyl chain order at low CHOL concentrations upon CHOL addition.
+In conclusion, a unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the CHOL ordering effect, particularly above 20~mol-\% of CHOL.
+Previously published comparison~\cite{ferreira13} suggests the same conclusion for historically relevant Berger/H{\"o}ltje force field parameters \cite{Berger97,Holtje01} that were not included in this work.
 
-Besides the ordering effect, effect of cholesterol on membrane properties has been discussed also in terms of the cholesterol condensing effect which 
-%The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}.
-The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL).
-However, condensation could also
-refers to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol (negative partial area) \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
-%Area per lipid increases less than what would be expected if the areas of POPC and CHOL were additive, manifesting a condensation effect of CHOL.
-However, beyond these qualitative features, the force fields behave very differently.
-At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, only MacRog predicts negative partial area for cholesterol, while CHARMM36 predicts zero partial area, and Slipids and Lipid17 predict small positive partial area.
-Considering that Slipids and Lipid17 perform best in our quality evaluation against experiments, yet still slightly overestimating the cholesterol condensing effect, our results suggests that cholesterol has positive but small partial area.
-%In addition to these trends, the absolute values of area per lipid and bilayer thickness vary substantially.
-At larger CHOL concentrations all force fields predict a too thick membrane, indicating an excessive ordering effect of CHOL.
-This is indeed verified by the order parameter maps, which suggest that the palmitate chain is overly ordered in all force fields to a varying degree.
+Besides the ordering effect, effect of CHOL on membrane properties has been discussed also in terms of the CHOL condensing effect which refers to a decrease in the \emph{area per phospholipid} upon the addition of CHOL (negative partial area) \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
+At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, only MacRog predicts negative partial area for CHOL, while CHARMM36 predicts zero partial area, and Slipids and Lipid17 predict small positive partial areas.
+Considering that Slipids and Lipid17 perform best in our quality evaluation against experiments, yet still slightly overestimating the CHOL condensing effect, our results suggests that CHOL has a positive but small partial area.
 
-%The best agreement with experiment is achieved with Slipids at low CHOL concentration and with Lipid17 at high CHOL concentration.
-In terms of dynamics, the lateral diffusion in pure POPC is too fast in Lipid17 and CHARMM36, and the addition of CHOL slows them both down more than in experiments.
-MacRog has overall too slow dynamics, whereas Slipids agrees quantitatively with the experiment, indicating that it provides the best description of lipid diffusion.
-
-%None of the force fields unambigously provides the best agreement with experiment. 
-Considering also lateral dynamics, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid-cholesterol interactions play major role.
-%and reasonable behavior in terms of membrane thickness and acyl chain ordering, especially at lower and intermediate CHOL concentrations.
-On the other hand, Lipid17 provides a good description for the palmitate chain ordering, but is otherwise in a relatively poor agreement with experiment except for higher CHOL concentrations.
-Thus, the choice of force field depends not only on the choice of membrane properties (strutural vs.
-dynamic) relevant to the problem, but also on the CHOL concentration range of interest.
-At low concentrations, Slipids is a safe choice, but at higher concentrations Lipid17 might provide the most realistic description.
-Nevertheless, all tested parameters capture qualitative effects of cholesterol on membrane properties relatively well and differences between force fields are clearly smaller than, for example, in the case of PC-PS lipid mixtures~\cite{antila2022emerging}.
-Therefore, quality of selected force field for other molecules, such as,
-%Since many research questions will involve 
-other lipids, proteins, sugars, or drugs, 
-%so the quality of the force fields describing these additional molecules 
-together with force field compatibility might be more relevant decisive factor for simulations of complex systems.
+Considering also lateral dynamics, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid--CHOL interactions play major role.
+Nevertheless, all tested parameters capture qualitative effects of CHOL on membrane properties relatively well and differences between force fields are clearly smaller than, for example, in the case of PC--PS lipid mixtures~\cite{antila2022emerging}.
+Therefore, the quality of selected force field for other molecules, such as, other lipids, proteins, sugars, or drugs, together with force field compatibility might be more relevant decisive factor for simulations of complex systems.
 Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behavior, yet we believe that evaluating the magnitude of these effects is well beyond the scope of this work.
@@
 Form factor plots from simulations and experiments with the traced minima at various cholesterol concentrations.
-%
-Interpolated 1\textsuperscript{st} and 2\textsuperscript{nd} minima of the form factors as a function of cholesterol concentration from simulation and experiment.
 %
@@
 %
-JJM thanks the Amber community (Benjamin Madej from the Walker lab and Callum Dickson from the Gould lab) for technical assistance related to the lipid14 force field parameters. 
+%JJM thanks the Amber community (Benjamin Madej from the Walker lab and Callum Dickson from the Gould lab) for technical assistance related to the lipid14 force field parameters. 
 %
-MJ thanks the Academy of Finland (Postdoctoral researcher grant no. 338160) and the Emil Aaltonen foundation for funding.
+MJ (grant no. 338160) and OHSO (grant nos. 315596 \& 319902) thank the Academy of Finland for funding.
+
 \end{acknowledgement}
@@
 
-%\newpage
-%\section{APPENDIX: The NMR results reported by Tiago Ferreira}
-
-\listoftodos
-
 \end{document}
-
-\begin{table*}[]
-  \centering
-  \caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
-    $^a$ The number of lipid molecules
-    $^b$ The number of cholesterol molecules
-    $^c$ Cholesterol concentration (mol\%)
-    $^d$ The number of water molecules
-    $^e$ Simulation temperature
-    $^f$ The total simulation time
-    $^g$ Time frames used in the analysis
-    $^h$ Reference link for the downloadable simulation files
-    $^i$ Reference for the full simulation details
-  }\label{systemsCHOL}
-  \begin{tabular}{c c c c c c c c c c c}
-    %\hline
-    Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
-    \hline
-    Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
-    /H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
-    &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
-    &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
-    &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
-    &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
-    &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
-CHARMM36\cite{klauda10,lim12}  & POPC   & 200& 0& 0\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 22& 10\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 50& 20\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 86& 30\% & 9000  & 310  & ? & 100 & [\citenum{charmm36gromacs5chol0-30}]$^*$  &  SI   \\
-                               & POPC   & 200& 134& 40\% & 15030  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
-                               & POPC   & 200& 200& 50\% & 18000  & 310  & 109 & 100 & [\citenum{charmm36gromacs5chol40-50}]$^*$  &  SI   \\
-Slipids\cite{jambeck12b,jambeck12,jambeck13b}  & POPC   & 200 & 0 & 0\%   & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 512 & 0 & 0\%   & 23943  & 310 & 170 & 100 & [\citenum{slipidsCHOL0T298}]$^*$ & SI  \\
-                                               & POPC   & 200 & 22 & 10\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 200 & 50 & 20\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 200 & 86 & 30\% & ?  & 310 & ? & 100 & [\citenum{slipidsCHOL0-30T310}]$^*$ & SI  \\
-                                               & POPC   & 358 & 154 & 30\% & 21183  & 298 & 170 & 100 & [\citenum{slipidsCHOL30T298}]$^*$ & SI  \\
-                                               & POPC   & 200 & 134 & 40\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
-                                               & POPC   & 200 & 200 & 50\% & ?  & 310 & 109 & 100 & [\citenum{slipidsCHOL40-50T310}]$^*$ & SI  \\
-                                               & POPC   & 256 & 256 & 50\% & 20334  & 298 & 170 & 100 & [\citenum{slipidsCHOL50T298}]$^*$ & SI  \\ 
-     MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}] \\ 
-                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & [\citenum{botan15}]    \\
-     Lipid14\cite{walker14,walker15}     & POPC   & 120 & 8 & 6\% &  3968 & 303 & 200 & 200 & [\citenum{lipid14sys1}]$^*$ & SI \\ 
-                          & POPC   & 108  & 20  & 16\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys2}]$^*$ & SI    \\
-                          & POPC   &  84  & 44 &  34\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys3}]$^*$ & SI    \\
-                             & POPC   & 64  & 64 & 50\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys4}]$^*$ & SI    \\
-                             & POPC   &  52  & 76 & 59\% & 3968  & 303  & 200 & 200 & [\citenum{lipid14sys5}]$^*$ & SI    \\
-\end{tabular}
-\end{table*}
-
-\begin{figure}[]
- \centering
- \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
- \caption{\label{FormFactors}
-   Form factors from simulations and experiments.
- }
- \todo{Details about form factor calculation code is discussed in issues
-   https://github.com/NMRLipids/MATCH/issues/56 and https://github.com/NMRLipids/MATCH/issues/50.
-   Once the code is finalized, we should recalculate and check the form factors.
-   CURRENTLY, 10\%-40\% cholesterol concetrations are calculated with a code which gives possibly incorrect
-   heigths for the maxima.
- } \\
- \todo{The y-axis scale cannot be explicitly measured in experiments.
-   Therefore, the y-axis of the experimental form factor is typically scaled to match with simulations.
-   This currently not done, because we have several simulations which give inequal form factors, and therefore
-   it is not clear against which simulation results we should scale the experimental results.
-   I have created a issue for this discussion: https://github.com/NMRLipids/NmrLipidsCholXray/issues/18
- } \\
- \todo{Not all experimental and simulation data is here (maybe?).}
-\end{figure}
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/apls.eps}
-  \caption{\label{apls}
-    Area per molecules calculated from different simulation models
-    as a function of cholesterol concentration. The solid symbols are area per
-    total amount of molecules (chol+PC) and the empty symbols area per PC headgroups.
-    Top figure shows absolute values and bottom figure shows changes respect
-    to pure lipid system.
-  }
-\end{figure}
-
-Visual inspection of form factors (Fig. \ref{FormFactors}) and quantitative quality estimation (Fig. \ref{FormFactorsFITNESS})
-suggests that the Berge/Holtje force field gives the best agreement with experiments with large
-cholesterol concentration. This is surprising because the same model overestimated acyl chain
-order parameter increase upon addition of cholesterol more than CHARMM36 or Slipids simulations
-(Figs. \ref{OrderParametersCHOL} and \ref{OrderParametersCHOLchanges}).
-The Berger/Holtje model also exhibits the most pronounced decrease in the area per molecule
-upon addition of cholesterol (Fig. \ref{alps}).
-On the other hand, the electron density profiles from Berger/Holtje simulations are closer
-to the result from the SDP model then other MD simulation models for both pure POPC bilayer
-and equimolar POPC/cholesterol mixture, especially in the middle of the bilayer
-where SDP model gives substantially lower density than MD simulations (Fig. \ref{densities}).
-Therefore, the good quality of form factor Berger/Holtje model with large amounts of cholesterol
-may rather indicate the importance of this low electron density in the bilayer center,
-rather than better quality of lipid-cholesterol interactions in this model.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/FFfitness.eps}
-  \caption{\label{FormFactorsFITNESS}
-    Fitness of form factors between simulations and experiments.
-  }
-    \todo{Figure to be done properly when the fitness code is finished.
-    This is discussed in issues
-    https://github.com/NMRLipids/MATCH/issues/65,
-    https://github.com/NMRLipids/MATCH/issues/64,
-    https://github.com/NMRLipids/MATCH/issues/56, and
-    https://github.com/NMRLipids/MATCH/issues/50
-  }
-\end{figure}
-
-The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
-with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8cm]{../FIGS/slopes.eps}
-  \caption{\label{OrderParametersCHOLchanges}
-    Slopes of order parameters as a function of cholesterol.
-    Determined by fitting equation $S_{\rm CH}(C_{\rm chol})=k_{\rm OP}C_{\rm chol}+S_{\rm CH}(0)$
-    to the data in figures \ref{slopes}-\ref{slopesslipids310K}.
-  }
-  \todo{It would be good to plot these slopes from also other experimental data in the literature.}
-\end{figure}
-
-% \begin{figure}[]
-%   \centering
-%   \includegraphics[width=8cm]{../FIGS/densitiesEXTREMES.eps}
-%   \caption{\label{densities}
-%     Electron density profiles from MD simulations and SDP model.
-%   }
-% \end{figure}
-
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
-  \caption{\label{OrderParametersCHOLfitness}
-    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-  }
-  \todo{Figure to be done properly when the fitness code is finished.
-    This is discussed in issues
-    https://github.com/NMRLipids/MATCH/issues/65,
-    https://github.com/NMRLipids/MATCH/issues/64,
-    https://github.com/NMRLipids/MATCH/issues/56, and
-    https://github.com/NMRLipids/MATCH/issues/50
-  }
-\end{figure*}
-
-% \begin{figure*}[]
-%   \centering
-%   \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
-%   \caption{\label{OrderParametersCHOL}
-%     Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-%   }
-%   \todo{Lipid14 results?} 
-% \end{figure*}
-
-\subsection{Presenting the quality of lipid bilayer MD simulation with binary mixtures}
-
-All the tested MD simulation models qualitatively reproduced experimentally observed condenced and
-ordering effects of cholesterol. However, the acyl chain ordering was overestimated and quality
-of form factors reduced with increasing amount of cholesterol. Although the biological importance
-of these details is unclear, overestimated ordering may promote formation of liquid ordered phases and
-inaccuate form factors complicate interpretation of scattering experiment with MD simulations.
-Therefore, MD simulation force fields correctly reproducing the order effects upon addition of the
-cholesterol would be invaluable.
-
-To facilitate the comparison of the quality of complex lipid bilayers,
-we introduce the quantitative quality measures which measure the quality of
-intermolecular interactions in binary lipid bilayers from different perspectives.
-The simple difference between experimental and simulated C-H bond order parameters
-give a very detailed picture about the quality of molecular conformations in all
-positions of lipid bilayer (Fig. \ref{OrderParametersCHOLfitness}).
-However, complexity of such presentation increases with increasing amount of data
-and it may be hard to present in practical format readable for human or machine.
-Using root mean square deviation across regions in the molecule simplifies the presentation,
-but compromizes the spatial resolution. Here, we divide the lipid into headgroup, glycerol
-backbone, sn-1 and sn-2 chains to present quality of binary lipid bilayer in more convenient
-format (Fig. \ref{OrderParametersCHOLfitness}).
-\todo{Discussion to be finished once we have the fitness function defined.}

### 2023-05-02 MANUSCRIPT/manuscript.tex
Update

@@
 \author{Matti Javanainen}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, CZ-16000 Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, FI-00790 Helsinki, Finland}
 \email{matti.javanainen@helsinki.fi}
@@
 \author{Josef Melcr}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, CZ-16000 Prague 6, Czech Republic}
 %
@@
 \alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
-
 %
 \author{O. H. Samuli Ollila}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, CZ-16000 Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, FI-00790 Helsinki, Finland}
 \email{samuli.ollila@helsinki.fi}
@@
 \author{Thomas Piggot}
-\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
+\affiliation{School of Chemistry, University of Southampton, Southampton SO17 1BJ, United Kingdom}
 \alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
@@
 Here, we define quantitative quality measures for binary lipid mixtures in membranes against C--H bond order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering.
-We then perform a systematic evaluation on how well the commonly used force fields describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
+Based on these measures, we then perform a systematic evaluation on the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
 None of the tested force fields clearly outperforms the others across the tested properties and conditions, yet Slipids parameters provide the best overall performance.

### 2023-05-02 MANUSCRIPT/manuscriptSUPPL.tex
Update

@@
 
-% \author{Fernando Favela-Rosales}
-% \affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
+\author{Fernando Favela-Rosales}
+\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
 %
@@
 %
-% \author{Jesper J. Madsen}
-% \affiliation{Department of Global Health, College of Public Health}
-% \alsoaffiliation{University of South Florida}
+\author{Jesper J. Madsen}
+\affiliation{Department of Global Health, College of Public Health}
+\alsoaffiliation{University of South Florida}
 %
-% \author{Josef Melcr}
-% \affiliation{Institute of Organic Chemistry and Biochemistry,
-% Academy of Sciences of the Czech Republic, 
-% Prague 6, Czech Republic}
+\author{Josef Melcr}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
 %
@@
 %
-% \author{Thomas Piggot}
-% \affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
-% \alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
+\author{Thomas Piggot}
+\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
+\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
 

### 2023-05-03 MANUSCRIPT/manuscript.tex

@@
 \author{Peter Heftberger}
-\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, AT-8010 Graz, Austria}
 %
@@
 \affiliation{Department of Chemistry, University of Bergen, Norway}
-\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
+\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, NO-5008 Bergen, Norway}
 %
@@
 \author{Georg Pabst}
-\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, AT-8010 Graz, Austria}
 \alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
@@
 None of the tested force fields clearly outperforms the others across the tested properties and conditions, yet Slipids parameters provide the best overall performance.
-
+%
 The quality evaluation metrics introduced in this work will particularly foster the future force field development and refinement for multi-component membranes with automated approaches.

### 2023-05-03 CoverLetter/cover.tex
Update

\documentclass[10pt]{article}

\usepackage{charter} % Use the Charter font for the document text
\usepackage[left=1.15in,right=1.15in,top=1in,bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{enumitem}

\setlength{\parindent}{0pt}% No paragraph indent
\setlength{\parskip}{.5\baselineskip}
\pagestyle{empty}% No headers/footers
\setlength{\tabcolsep}{0.5cm}

\begin{document}

\begin{flushright}
\includegraphics[height=2.5cm]{CoverLetter/logo_uh.pdf}
\end{flushright}

\bigskip
\bigskip
\bigskip
Dear Editor,
\bigskip

Please find enclosed our original manuscript \emph{``Quantitative Comparison With Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions''} by Peter Heftberger, Jesper J. Madsen, Josef Melcr, Markus Miettinen, Georg Pabst, and the signees to be considered for publication in \emph{Journal of Chemical Theory and Computation}.

Atomistic molecular dynamics simulations are being applied to ever more complex problems in biosciences, yet the quality of these studies relies on the accuracy of the used force fields. Membrane proteins are cellular factories that reside in complex lipid environments in the plasma membrane and the membranes of intracellular organelles. While the models used in protein simulations have been validated thoroughly throughout the recent decades, the lipid ones have not been under similar scrutiny. This is surprising, considering that we have recently started to appreciate lipids as key regulators of membrane protein function through both specific interactions and membrane-mediated effects. Of most interest is cholesterol, which is known to bind to specific sites of GPCRs as well as to alter membrane properties.

Here, we evaluate the ability of present force fields to capture the phospholipid--cholesterol interactions that are crucial for the accurate modeling of membrane-mediated lipid--protein interactions. Probe-free experimental techniques, NMR and X-ray scattering, provide a direct and quantitative metrics for force field quality. Importantly, by an exhaustive simulation approach with multiple system sizes, we eliminate the finite-size effects, and provide the first benchmark of the effect of cholesterol on lateral dynamics among force fields and in comparison to NMR experiments. All in all, our comparison based on both structural and dynamic properties reveals significant limitations in the present force fields. We hope that our findings will guide the refinement of the next generation of atomistic lipid force fields, and thereby improve the quality of membrane protein simulations.

All the used simulation inputs and outputs are openly available online in the Zenodo repository (https://zenodo.org), and they can be programmatically accessed through the NMRlipids Databank (https://databank.nmrlipids.fi), allowing other scientists to verify our findings and to extend the presented analysis. We believe that \emph{Journal of Chemical Theory and Computation} together with these open access strategies provides the best platform for the rapid dissemination of our work. We hope to be able to improve our manuscript \textit{via} the peer review process.

\vspace{2cm}
\begin{center}
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r}
	Dr. Matti Javanainen,     & Dr. Samuli Ollila, \\ 
	Postdoctoral Researcher & Academy Research Fellow\\
    & \\
    \multicolumn{2}{c}{Institute of Biotechnology, University of Helsinki} \\ 
	\multicolumn{2}{c}{Viikinkaari 5, FI-00790 Helsinki, Finland} \\
	& \\
    matti.javanainen@helsinki.fi & samuli.ollila@helsinki.fi \\
\end{tabular*}
\end{center}

\end{document}

### 2023-05-03 MANUSCRIPT/manuscript.tex
Update

@@
 
-\author{Fernando Favela-Rosales}
-\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
-%
 \author{Peter Heftberger}
@@
 %
-\author{Thomas Piggot}
-\affiliation{School of Chemistry, University of Southampton, Southampton SO17 1BJ, United Kingdom}
-\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
 

### 2023-05-03 MANUSCRIPT/manuscriptSUPPL.tex
Update

@@
 %
-\author{Thomas Piggot}
-\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
-\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}
 

### 2023-05-18 MANUSCRIPT/manuscript.tex
Update manuscript.tex

@@
 Based on these measures, we then perform a systematic evaluation on the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
-None of the tested force fields clearly outperforms the others across the tested properties and conditions, yet Slipids parameters provide the best overall performance.
+None of the tested force fields clearly outperforms the others across the tested properties and conditions but Slipids parameters provide the best overall performance.
 %
@@
 
-Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
-A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}.
+Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
+A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane, viral envelopes) \cite{van2008membrane}.
 CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}.
@@
 
-While simulations of complex membranes with CHOL have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge.
+While simulations of complex membranes with CHOL have become relatively straightforward task, estimating the trustworthiness of MD simulations for membranes remains a challenge, especially as the number of components and complexity increase.
 Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C--H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}.
@@
 An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both simulations and experiments, and were thus treated separately in our analyses.
-The C--H order parameters for the POPC head group are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.
+The C--H order parameters for the POPC head group are shown separately for all CH$_2$ groups, \textit{i.e.}, no averaging was performed.
 
@@
 2D matrices were created by interpolation for the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains and CHOL concentration) using the \texttt{interp2} function in Matlab based on linear interpolation.
-Similar matrices were also generated for the electron density profiles (as a function of normal distance from bilayer center and CHOL concentration) but with the \texttt{scatteredInterpolant} function in Matlab based on Delanuay triangulation \cite{amidror2002scattered}.
+Similar matrices were also generated for the electron density profiles (as a function of normal distance from bilayer center and CHOL concentration) but with the \texttt{scatteredInterpolant} function in Matlab based on Delaunay triangulation \cite{amidror2002scattered}.
 Linear 1D interpolation with the \texttt{interp1} function in Matlab was used for the CHOL-dependence of the 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations and diffusion coefficients.
@@
 All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of CHOL concentration.
-For C--H bond order parameters, also the deviation matrices between simulations and experiments were used to illustrate quality of simulations.
+For C--H bond order parameters, the deviation matrices between simulations and experiments were also used to illustrate quality of simulations.
 
@@
 Consequently, the correct CHOL ordering effect is expected to be a necessary condition for a force field used to understand lipid--CHOL phase behaviour.
-CHOL ordering effect can be experimentally quantified by measuring the C--H bond order parameters using $^{13}$C or $^{2}$H NMR techniques, and the result can be also directly compared to MD simulations~\cite{ollila16}.
+CHOL ordering effect can be experimentally quantified by measuring the C--H bond order parameters using $^{13}$C or $^{2}$H NMR techniques, and the result can also be directly compared to MD simulations~\cite{ollila16}.
 Simulations generally reproduce the CHOL ordering effect, but order parameters often deviate from experiments at high CHOL concentrations~\cite{ferreira13,madej15}.
@@
 MD simulations agree well with the SDP model below 30~mol-\% but overshoot the SDP results at high CHOL concentrations.
-Lipid17 simulations are exception as they predict thinner membranes than SDP model at low CHOL concentrations, and clear saturation of thickness increase is not observed, unlike for other force fields and experiments.
+Lipid17 simulations are the exception as they predict thinner membranes than SDP model at low CHOL concentrations, and clear saturation of thickness increase is not observed, unlike for other force fields and experiments.
 
@@
 SDP model gives the most sharp bands (bottom left panel in Fig.~\ref{fig:densmaps} and original electron density profiles in Fig.~\ref{SIfig:densprofs}), suggesting more uniform membrane structure than any MD simulation.
-However, System size plays a role here as  undulations smear out the density profile features with increasing simulation box size as demonstrated in Fig.~\ref{SIfig:densprofssize} by the density maps calculated for the CHARMM36 simulations in three sizes.
+However, system size plays a role here as  undulations smear out the density profile features with increasing simulation box size as demonstrated in Fig.~\ref{SIfig:densprofssize} by the density maps calculated for the CHARMM36 simulations in three sizes.
 Still, even in the smallest system, the band intensities are less localized than in the SDP model.
@@
 To investigate if the differences in lateral diffusion coefficients could be explained by different Lennard-Jones (LJ) cutoff values, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids using a shorter cutoff of 0.9~nm corresponding to that of Lipid17, while original cutoff for Slipids was 1.4~nm (see Table~\ref{SItab:params}).
-The PBC-corrected diffusion coefficient values with shorter cutoff, shown in the top panel of Fig.~\ref{fig:dynamics} as hollow red circles, are only slightly larger than the original values, indicating that the LJ cutoff alone does not explain the differences between force fields, but they truly spark from differences in the force field interaction parameters.
+The PBC-corrected diffusion coefficient values with shorter cutoff, shown in the top panel of Fig.~\ref{fig:dynamics} as hollow red circles, are only slightly larger than the original values, indicating that differences arise from a combination of force field interaction parameters and are not explained by the LJ cutoff alone.
 
@@
 For example, Fig.~\ref{SIfig:dvschol} shows that Lipid17 and CHARMM36 systemically underestimate the experimental values in systems with all sizes simulated here while the effect of CHOL seems to be qualitatively reproduced.
-However, the PBC corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations leading to values close to experimental ones at 47~mol-\% CHOL.
-With MacRog, also the PBC corrected values fall below the experimental ones, yet the slowdown effect of CHOL again seems stronger than in experiment.
+However, the PBC corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations, leading to values close to experimental ones at 47~mol-\% CHOL.
+With MacRog, also the PBC corrected values fall below the experimental ones, yet the slowdown effect of CHOL again appears stronger than in experiment.
 With Slipids, the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent.
@@
 
-Since the contributions from the form factor minima deviations are barely visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
+Since the contributions from the form factor minima deviations barely are visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
 This panel demonstrates that apart from CHARMM36, all force fields capture the location of the 1\textsuperscript{st} minimum relatively well at low CHOL concentrations, but deviate significantly at larger CHOL concentrations.
@@
 Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results up to 20~mol-\% of CHOL, but overestimate acyl chain ordering and membrane thickness after further addition of CHOL.
-An apparent exception to this is the Lipid17 force field, yet their seemingly better agreement with experiments can be explained by the over-correction of the originally underestimated acyl chain order at low CHOL concentrations upon CHOL addition.
+An apparent exception to this is the Lipid17 force field, yet its seemingly better agreement with experiments can be explained by the over-correction of the originally underestimated acyl chain order at low CHOL concentrations upon CHOL addition.
 In conclusion, a unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the CHOL ordering effect, particularly above 20~mol-\% of CHOL.
@@
 
-Besides the ordering effect, effect of CHOL on membrane properties has been discussed also in terms of the CHOL condensing effect which refers to a decrease in the \emph{area per phospholipid} upon the addition of CHOL (negative partial area) \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
+Besides the CHOL-induced ordering, effects on membrane properties has been discussed also in terms of the CHOL condensing effect which refers to a decrease in the \emph{area per phospholipid} upon the addition of CHOL (negative partial area) \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
 At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, only MacRog predicts negative partial area for CHOL, while CHARMM36 predicts zero partial area, and Slipids and Lipid17 predict small positive partial areas.
-Considering that Slipids and Lipid17 perform best in our quality evaluation against experiments, yet still slightly overestimating the CHOL condensing effect, our results suggests that CHOL has a positive but small partial area.
+Considering that Slipids and Lipid17 perform best in our quality evaluation against experiments, yet still slightly overestimating the CHOL condensing effect, our results suggest that CHOL has a positive but small partial area.
 

### 2023-05-18 MANUSCRIPT/manuscript.tex
Update manuscript.tex

@@
 \author{Jesper J. Madsen}
-\affiliation{Department of Global Health, College of Public Health}
-\alsoaffiliation{University of South Florida}
+\affiliation{Global and Planetary Health, College of Public Health}
+\alsoaffiliation{Department of Molecular Medicine, Morsani College of Medicine}
+\alsoaffiliation{University of South Florida, Tampa, Florida, 33612, United States of America}
 %
@@
 
+\section*{Author contributions statement}
+
+J.J.M. contributed to early testing (debugging, setup, and running) of simulations with the Amber force field and manuscript editing.
+
 \bibliography{refs.bib}

### 2023-06-10 MANUSCRIPT/manuscript.tex
Update manuscript

@@
 }
+
 \myexternaldocument{manuscriptSUPPL}
@@
 \author{Peter Heftberger}
-\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, AT-8010 Graz, Austria}
+\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
 %
 \author{Matti Javanainen}
-\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, CZ-16000 Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki, FI-00790 Helsinki, Finland}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, 16000 Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
 \email{matti.javanainen@helsinki.fi}
@@
 %
-\author{Josef Melcr}
-\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, CZ-16000 Prague 6, Czech Republic}
-%
-\author{Markus Miettinen}
-\affiliation{Department of Chemistry, University of Bergen, Norway}
-\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, NO-5008 Bergen, Norway}
+\author{Markus S. Miettinen}
+\affiliation{Fachbereich Physik, Freie Universit\"at Berlin, 14195 Berlin, Germany}
+\alsoaffiliation{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
+\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
 %
 \author{O. H. Samuli Ollila}
-\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, CZ-16000 Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki, FI-00790 Helsinki, Finland}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, 16000 Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
 \email{samuli.ollila@helsinki.fi}
@@
 \author{Georg Pabst}
-\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, AT-8010 Graz, Austria}
-\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
+\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
+\alsoaffiliation{BioTechMed-Graz, 8010 Graz, Austria}
+\alsoaffiliation{Field of Excellence BioHealth—University of Graz, 8010 Graz, Austria}
 %
@@
 %\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
-\title{Quantitative Comparison With Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions}
+\title{Quantitative Comparison Against Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions}
 
@@
 Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, renders the membrane stiffer, and drives domain formation.
-Molecular dynamics simulations have played a crucial role in resolving these effects at molecular level, yet it has recently become evident that different simulation force fields predict quantitatively different behavior.
-Although easily neglected, identifying such limitations are becoming increasingly important as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions where it is crucial to accurately model the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol.
+Molecular dynamics (MD) simulations have played a crucial role in resolving these effects at the molecular level, yet it has recently become evident that different MD force fields predict quantitatively different behavior.
+Although easily neglected, identifying such limitations is increasingly important as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions: Pertinent multi-component simulations must capture accurately the interactions between their fundamental building blocks, such as phospholipids and cholesterol.
 %
-Here, we define quantitative quality measures for binary lipid mixtures in membranes against C--H bond order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering.
-Based on these measures, we then perform a systematic evaluation on the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
-None of the tested force fields clearly outperforms the others across the tested properties and conditions but Slipids parameters provide the best overall performance.
+Here, we define quantitative quality measures for simulations of binary lipid mixtures in membranes against C--H bond order parameters and lateral diffusion coefficients from NMR spectroscopy as well as form factors from X-ray scattering.
+Based on these measures, we perform a systematic evaluation on the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
+None of the tested force fields clearly outperforms the others across the tested properties and conditions, but the Slipids parameters provide the best overall performance.
 %
-The quality evaluation metrics introduced in this work will particularly foster the future force field development and refinement for multi-component membranes with automated approaches.
+The quality-evaluation metrics introduced in this work will, particularly, foster future force field development and refinement for multi-component membranes using automated approaches.
 \end{abstract}
@@
 Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}.
-A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane, viral envelopes) \cite{van2008membrane}.
-CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}.
+A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations from $\sim$10~mol-\% (endoplasmic reticulum) up to $\sim$50~mol-\% (plasma membrane, viral envelopes) \cite{van2008membrane}.
+CHOL has the unique ability to order neighbouring lipids and thus induce the liquid-ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}.
 In the cellular setting, the interaction between other lipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}.
@@
 
-While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations have been employed to obtain a detailed view on the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}.
-Such efforts are further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable the simulations of ever more complex and thus realistic membranes \cite{lorent2020plasma}.
+While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atom-resolution molecular dynamics (MD) simulations have been employed to obtain a detailed view of the lateral organization driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}.
+Further MD efforts are facilitated by the growing availability of force fields with compatible lipid and protein parameters---enabling simulations of ever more complex and thus biologically relevant membranes \cite{lorent2020plasma}.
 
-The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the force field by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig16}, respectively.
-Notably, simulations using CHARMM36, Lipid17, and Slipids can now be set up using CHARMM-GUI for multiple simulation engines, which enables the rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.
+The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have sizeable libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the force field by Maciejewski and R\'og (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig16}, respectively.
+Notably, simulations using CHARMM36, Lipid17, and Slipids can these days be set up for multiple simulation engines with the CHARMM-GUI online tool; this decreases greatly the manual work needed to run complex membrane simulations \cite{lee16,lee2020charmm}.
 
-While simulations of complex membranes with CHOL have become relatively straightforward task, estimating the trustworthiness of MD simulations for membranes remains a challenge, especially as the number of components and complexity increase.
-Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C--H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}.
-This approach has been useful for finding the best force fields for the description of the headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE)~\cite{bacle2021inverse} and phosphatidylglycerol (PG)~lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}.
-Furthermore, quantitative quality measures based on C--H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}.
+While simulating complex membranes with CHOL has become a relatively straightforward task, estimating the trustworthiness of MD simulations remains a challenge, especially as the complexity and number of membrane components increase.
+Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against the C--H bond order parameters from NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}.
+This approach has been useful for finding the best force fields to describe the headgroups of phosphatidylcholine (PC)~\cite{botan15,kurki2022structure}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE)~\cite{bacle2021inverse} and phosphatidylglycerol (PG)~lipids \cite{bacle2021inverse}, to evaluate and improve membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and to find simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}.
+Furthermore, quantitative quality measures based on C--H bond order parameters have been recently defined and used to rank simulations in the NMRlipids Databank~\cite{NMRlipidsDatabank}.
 However, such automatic quality evaluation is limited to simulations for which experimental data at a corresponding composition and temperature are available.
-Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automatized in the NMRlipids databank.
+Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automatized in the NMRlipids Databank.
 
-Here, we demonstrate how simulations of binary POPC--CHOL mixtures can be evaluated against experimental NMR and X-ray scattering data by interpolating through multiple CHOL concentrations.
-Because effect of CHOL on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously \cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in CHOL induced lateral membrane heterogeneity.
-We also evaluated the dependence of lateral diffusion coefficients on CHOL against pulsed field gradient (PFG) NM experiments \cite{filippov2003effect,filippov2003influence} using the recent theoretical framework that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}.
-With the structural and dynamic comparisons established, we estimated the quality of popular force fields at different CHOL concentrations.
-While we focus here on POPC-CHOL mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems.
+Here, we demonstrate how simulations of binary PC--CHOL mixtures can be evaluated against experimental NMR spectroscopy and X-ray scattering data by interpolating through multiple CHOL concentrations.
+As the effect of CHOL on lipid headgroup (and its independence from acyl chain ordering) has been discussed previously \cite{botan15,antila22b}, we focus here on the acyl chains; these are expected to play a larger role than the headgroup in CHOL-induced lateral membrane heterogeneity.
+We also evaluate the dependence of lateral diffusion coefficients on CHOL against pulsed field gradient (PFG) NMR experiments \cite{filippov2003effect,filippov2003influence} using the recent theoretical framework that allows a quantitative comparison with experiment by eliminating finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}.
+With the structural and dynamic comparisons established, we then estimate the quality of four popular force fields at different CHOL concentrations.
+While we focus here on a PC--CHOL mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in any binary or multi-component system.
 
@@
 Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high}.
-This avoids the precipitation of CHOL crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% CHOL content.
+This avoids the precipitation of CHOL crystallites at high concentration \cite{buboltz1999novel}, yielding non-phase separated samples up to 50~mol-\% CHOL content.
 Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification.
@@
 The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtained both contributions.
-The electron density profiles were modelled from form factors using the SDP model where volume distribution functions are modelled by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}.
-CHOL was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}.
-Membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
-The electron density maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
+The electron density profiles were modelled from form factors using the SDP model, where volume distribution functions are described by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}.
+CHOL was described using two Gaussians, as detailed in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}.
+The membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
+The electron density maxima were extracted by the \texttt{findpeaks} function in Matlab.
 
@@
 We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%.
-Systems were simulated using four commonly used force fields, namely CHARMM36 (often abbreviated ``C36'' in figure legends in this work) \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig16} force fields.
-In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256 or 1024 POPC molecules).
+Systems were simulated using four commonly used force fields: CHARMM36 (often abbreviated ``C36'' in figure legends in this work) \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and all-atom OPLS compatible MacRog \cite{kulig14,kulig15,Kulig16}.
+To eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256, or 1024 POPC molecules).
 The number of POPC molecules was kept constant across the different CHOL concentrations.
 All membranes were solvated by 50 waters per lipid (POPC or CHOL).
-The small membranes (with 64 lipids) were first generated using CHARMM-GUI and equilibrated using the standard protocols for CHARMM36, Slipids, and Lipid17 for which inputs are readily available from CHARMM-GUI \cite{lee16,lee2020charmm}.
+The small membranes (with 64 lipids) were first generated using CHARMM-GUI and equilibrated using the standard protocols for CHARMM36, Slipids, and Lipid17, for which inputs are readily available from CHARMM-GUI \cite{lee16,lee2020charmm}.
 Then, the atomic coordinates were replicated in the membrane plane to create the 4- and 16-fold larger simulation systems.
-Since CHARMM-GUI does not support MacRog, the production simulations were initiated from equilibrated CHARMM36 structures, since they share the atom ordering.
+Since CHARMM-GUI does not support MacRog, the production simulations were initiated from equilibrated CHARMM36 structures, since the two force fields share the atom ordering.
 All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}.
 The first 10~ns of each simulation were omitted from analyses.
-The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
-All simulations were stored to the NMRlipids databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}.
+The simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749  (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).
+All simulations were added to the NMRlipids Databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}.
 The CHARMM36 simulations have been previously analyzed for their dynamic properties in Ref.~\citenum{fabian2023protein}.
@@
     \caption{\label{tab:simulations}%
-    Details of the simulation systems provided for (small/medium/large) system sizes).
-The box dimensions in the membrane plane ($x/y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
+    Details of the (small/medium/large) simulation systems.
+The box dimensions in the membrane plane ($x$ and $y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
     }
@@
     \toprule
-    $[$CHOL$]$ & POPC & CHOL & Water & $x/y$ (nm) & $z$ (nm) \\
+    $[$CHOL$]$ & POPC & CHOL & Water & $x$ and $y$ (nm) & $z$ (nm) \\
     \midrule
@@
 
-\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles, automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank}, were used.
+\paragraph{Structural properties:} The C--H bond order parameters, form factors, and electron density profiles, automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank}, were used.
 Similarly to the experimental X-ray scattering data, membrane thickness was defined as twice the distance from the electron density maximum to the membrane center.
-Locations of maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.
-Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.
+Locations of maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data with the \texttt{smooth} function (5-point moving interval) in Matlab.
+Area per phospholipid was obtained by dividing the area of the bilayer (area of the simulation box) by the number of phospholipids in one leaflet.
 
-To simplify the interpolation of order parameter data to intermediate CHOL concentrations (see \textit{``Quantitative quality evaluation of the effect of CHOL on membrane properties''} below), C--H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged.
+To simplify the interpolation of order parameter data to intermediate CHOL concentrations (see \textit{``Quantitative quality evaluation of the effect of CHOL on membrane properties''} below), C--H bond order parameters of the 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged.
 These groups rotate freely and thus the order parameters are essentially identical for both (all) hydrogens in experiments and simulations.
@@
 The water viscosity value of $\mu_\mathrm{f}= 0.3228$~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).
-The simulation box dimensions, membrane edge length $L$ and the dimension normal to the membrane plane ($z$) $L_z$, were taken from the final configuration of each simulation.
+The simulation box dimensions (membrane edge length $L$ and the dimension normal to the membrane plane ($z$) $L_z$) were taken from the final configuration of each simulation.
 $L_z$ was only needed in the calculation of $H$ (see below).
@@
 
-CHOL is known to induce order in lipid membranes by increasing the fraction of \textit{anti} conformations in the acyl chains of phospholipids \cite{ferreira13}, which is suggested to play critical role in the phase behaviour of PC--CHOL mixtures \cite{ipsen87}.
-Consequently, the correct CHOL ordering effect is expected to be a necessary condition for a force field used to understand lipid--CHOL phase behaviour.
-CHOL ordering effect can be experimentally quantified by measuring the C--H bond order parameters using $^{13}$C or $^{2}$H NMR techniques, and the result can also be directly compared to MD simulations~\cite{ollila16}.
-Simulations generally reproduce the CHOL ordering effect, but order parameters often deviate from experiments at high CHOL concentrations~\cite{ferreira13,madej15}.
-Furthermore, it has not been clear how accurately different force field parameters capture the details of lipid--CHOL interactions and which force field would give most realistic results for simulations of complex mixtures where such interactions play critical roles.
+CHOL is known to induce order in lipid membranes by increasing the fraction of \textit{anti} conformations in the acyl chains of phospholipids \cite{ferreira13}, which is suggested to play a critical role in the phase behaviour of PC--CHOL mixtures \cite{ipsen87}.
+Consequently, the correct CHOL-induced ordering is expected to be a necessary condition for a force field used to understand lipid--CHOL phase behaviour.
+The CHOL-induced ordering can be experimentally quantified by measuring the C--H bond order parameters using $^{13}$C or $^{2}$H NMR, and the results can also be directly compared to MD simulations~\cite{ollila16}.
+Simulations generally show some CHOL-induced ordering, but the order parameter values often deviate from the experimental ones at high CHOL concentrations~\cite{ferreira13,madej15}.
+Furthermore, it has not been clear how accurately different force fields capture the details of lipid--CHOL interactions, and which force field would give most realistic results for simulations of complex mixtures, where such interactions play critical roles.
 
-Here, we evaluate the CHOL ordering effect in state-of-the-art force field parameters against C--H bond order parameter data from $^{13}$C NMR experiments measured from POPC--CHOL mixture with CHOL concentratons ranging between 0 and 60~mol-\%~\cite{ferreira13}.
+Here, we evaluate the CHOL-induced ordering in state-of-the-art force fields against C--H bond order parameter data from $^{13}$C NMR experiments measured from POPC--CHOL mixtures with CHOL concentrations ranging between 0 and 60~mol-\%~\cite{ferreira13}.
 To this end, we first interpolated order parameter maps as a function of acyl chain carbon number and CHOL concentration for both simulations and experiments.
-These maps were then subtracted to obtain the deviation maps between simulations and experiments.
+These maps were then subtracted to obtain deviation maps between simulations and experiments.
 The deviation maps of different force fields are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} for the palmitate and oleate chains, respectively.
 
-The CHOL ordering effect is manifested in original profiles in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} as substantial increase in absolute values of acyl chain C--H bond order parameters upon addition of CHOL in all simulation force fields and experiments.
-The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive view for a quantitative comparison of different force fields with experiments.
-In white regions, the simulation results are considered to fall within the experimental error as the deviations are in the range of [--0.02,0.02].
-Blue indicates that the order parameters are too negative, \textit{i.e.}, the acyl chains are too ordered in simulations, and \textit{vice versa} for red.
+In all simulation force fields and experiments, the CHOL-induced ordering is manifested in the original profiles (Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate}) as a substantial increase in the absolute values of acyl chain C--H bond order parameters upon addition of CHOL.
+The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive view for a quantitative comparison of different force fields against experiments: In white regions, the simulation results are considered to fall within the experimental error as the deviations are in the range of [--0.02,0.02]; blue indicates that the order parameters are too negative, \textit{i.e.}, the acyl chains are too ordered in simulations, and \textit{vice versa} for red.
 Overall, the simulation force fields behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.
@@
 Data are shown for palmitate (top row) and oleate (bottom row) and for the four force fields (columns).
-Negative values indicate that order is too high ($-S_\mathrm{CH}$ values too negative) in the simulations.
-The values that are within the estimated experimental error range of 0.02~\cite{ollila16} are coloured in white.
-Order parameters of identical hydrogens attached to the same carbons were averaged with the exception of C2 carbon of the oleate chain with the forked order parameters for which differences for both the larger and smaller values were calculated, and the average of this deviation is shown in the C2 column.
+Negative values indicate that order is too high ($S_\mathrm{CH}$ values too negative) in the simulations.
+The values that are within the estimated experimental error range of $\pm$0.02~\cite{ollila16} are coloured white.
+Statistical error in simulations is not considered here because it is approximately order of magnitude smaller than the experimental error.
+Order parameters of hydrogens attached to the same carbon were averaged, except for the C2 carbon of the oleate chain (whose order parameters forked) for which differences for both the larger and smaller values were calculated, and the average of these differences is shown in the C2 column.
   }
@@
 Lipid17 provides the best overall agreement with experiment, as no segments deviate significantly from experiment at any CHOL concentrations.
-MacRog behaves reasonably well at low CHOL concentrations, yet at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.
-For the head group and glycerol backbone order parameters, provided in Fig.~\ref{SIfig:headgroups}, CHARMM36 give the best agreement at all cholesterol concentrations in line with previous studies~\cite{botan15,antila2019headgroup}.
+MacRog behaves reasonably well at low CHOL concentrations, yet at larger CHOL concentrations the chains become too ordered, leading to the largest overall deviations from experiment.
+For the head group and glycerol backbone order parameters, provided in Fig.~\ref{SIfig:headgroups}, CHARMM36 gives the best agreement at all cholesterol concentrations, in line with previous studies~\cite{botan15,antila22b}.
 
@@
 
-CHOL-induced ordering straightens the acyl chain conformations that leads to the thickening of membranes.
-While acyl chain order and membrane thickness are well correlated~\cite{NMRlipidsDatabank}, lipid bilayer dimensions can be accessed more directly by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16}.
-Electron density profile, area per lipid and bilayer thickness can all be extracted from the form factor using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,doktorova2020molecular}.
-To complement the evaluation of CHOL ordering effect against NMR order parameters, we measured also X-ray scattering form factors from POPC--CHOL mixtures with systematically increasing CHOL concentrations.
+CHOL-induced ordering straightens the acyl chain conformations, which leads to membrane thickening.
+While acyl chain order and membrane thickness are well correlated~\cite{NMRlipidsDatabank}, lipid bilayer dimensions can be accessed more directly by measuring the X-ray form factor, which is related to the electron density along membrane normal \textit{via} Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16}.
+Electron density profile, area per phospholipid, and bilayer thickness can all be extracted from the form factor using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,doktorova2020molecular}.
+To complement the evaluation of CHOL-induced ordering against NMR order parameters, we measured also the X-ray form factors from POPC--CHOL mixtures at systematically increasing CHOL concentrations.
 Scattering intensities from experiments are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, form factors from experiments and simulations in Fig.~\ref{SIfig:scattering}, and density profiles in Fig.~\ref{SIfig:densprofs}.
@@
 The effect of CHOL on structural properties of bilayers is compared between the SDP model (based on experimental form factors) and MD simulation results in Fig.~\ref{fig:densmaps}.
-All force fields demonstrate increasing thickness upon addition of CHOL that saturates after approximately 30~mol-\% (bottom middle panel in Fig.~\ref{fig:densmaps}).
-MD simulations agree well with the SDP model below 30~mol-\% but overshoot the SDP results at high CHOL concentrations.
-Lipid17 simulations are the exception as they predict thinner membranes than SDP model at low CHOL concentrations, and clear saturation of thickness increase is not observed, unlike for other force fields and experiments.
+All force fields demonstrate increasing thickness upon addition of CHOL that tends to saturate after approximately 30~mol-\% (bottom middle panel in Fig.~\ref{fig:densmaps}).
+Most MD simulations agree well with the SDP model below 30~mol-\%, but overshoot the SDP results at high CHOL concentrations.
+Lipid17 simulations are the exception: They predict thinner membranes than the SDP model at low CHOL concentrations, and a clear saturation of thickness increase is not observed, unlike for other force fields and experiments.
 
-The dependence of APL on CHOL concentration follows the trends in thickness inversely in general (bottom right panel of Fig.~\ref{fig:densmaps}), yet provide curious differences between the force fields at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
+The dependence of the area per phospholipid (APL) on CHOL concentration follows the trends in thickness inversely in general (bottom right panel of Fig.~\ref{fig:densmaps}), yet provides curious differences between force fields at the physiologically relevant CHOL concentration range \cite{van2008membrane}.
 Lipid17 has the largest APL across the entire CHOL contration range.
-MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect.
-The profiles for Slipids and CHARMM36 are very similar with the small or zero CHOL partial area until a concentration of 20~mol-\%.
+MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration, indicating a particularly strong condensation effect.
+The profiles for Slipids and CHARMM36 are very similar, with a small or zero partial area of CHOL until a concentration of 20~mol-\%.
 
@@
   \caption{\label{fig:densmaps}%
-  \textbf{Electron density profiles, thickness and area per lipid as a function of CHOL concentration.}
+  \textbf{Electron density profiles, thickness, and area per phospholipid as a function of CHOL concentration.}
   %
-  Top: Electron density maps for the simulations using four different force fields.
-  %
-  Bottom left: Electron density map for the experiment from the SDP model.
-The scaling is uniform between all simulations and experiment.
+  Top: Electron density maps for the simulations using four different force fields and for the experiment (from the SDP model).
+The colorbar is common for all maps.
 The original electron density profiles are shown in Fig.~\ref{SIfig:densprofs}.
@@
   %
-  Bottom center: Bilayer thickness.
+  Bottom left: Bilayer thickness.
 Thickness is defined as twice the distance from the peak in electron density to the membrane core.
 Experimental data are extracted in a similar manner from electron density profiles obtained with X-ray scattering.
+The bin width used in the profiles is used as the error estimate.
   %
-  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids. The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
+  Bottom middle: area per phospholipid measured by dividing the total membrane area by the number of phospholipids.
+The size-dependency of the area per phospholipid is shown in Fig.~\ref{SIfig:aplvssize}.
+  %
+  Bottom right: Effect of CHOL on the location of the first two minima in the form factor.
+The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
   }
@@
 
-For more detailed comparison of membrane structure, we interpolated the changes in electron density profiles along membrane normal (Fig.~\ref{SIfig:densprofs}) as a function of CHOL concentration to create two-dimensional electron density maps shown in Fig.~\ref{fig:densmaps}.
-Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region containing electron-rich phosphorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
-However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields and SPD model.
-CHARMM36 has the sharpest low- and high-density bands among MD simulation results indicating smaller membrane fluctuations that would smear the bands, and the same is true for MacRog at higher CHOL concentrations.
-The less sharp bands for Lipid17 and especially Slipids profiles indicate that the membranes are more flexible with these force fields.
-SDP model gives the most sharp bands (bottom left panel in Fig.~\ref{fig:densmaps} and original electron density profiles in Fig.~\ref{SIfig:densprofs}), suggesting more uniform membrane structure than any MD simulation.
+For more detailed comparison of membrane structure, we interpolated the changes in electron density profiles along membrane normal (Fig.~\ref{SIfig:densprofs}) as a function of CHOL concentration to create two-dimensional electron density maps (Fig.~\ref{fig:densmaps}).
+Overall, all electron density profiles share the same features across all CHOL concentrations: High-densities band corresponding to the tightly-packed interfacial regions containing the electron-rich phosphorus, low density at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
+However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields and the SPD model.
+CHARMM36 has the sharpest low- and high-density bands among MD simulation results, indicating smaller membrane undulations that would smear the bands; the same is true for MacRog at higher CHOL concentrations.
+The less sharp bands for Lipid17 and especially Slipids profiles indicate that their membranes are more flexible.
+The SDP model gives the sharpest bands (bottom left panel in Fig.~\ref{fig:densmaps} and original electron density profiles in Fig.~\ref{SIfig:densprofs}), suggesting more uniform membrane structure than any MD simulation.
 However, system size plays a role here as  undulations smear out the density profile features with increasing simulation box size as demonstrated in Fig.~\ref{SIfig:densprofssize} by the density maps calculated for the CHARMM36 simulations in three sizes.
 Still, even in the smallest system, the band intensities are less localized than in the SDP model.
-This difference might arise because most undulations are assigned to structure factor rather than form factor in their deconvolution, but it could also indicate different elastic properties between simulation and experiment.
+This might indicate different elastic properties between simulation and experiment, but we cannot fully exclude the role of models used to interpret scattering data.
 
-Nevertheless, minima in X-ray scattering form factors are independent on simulation box size and correlate with membrane dimensions~\cite{NMRlipidsDatabank}.
-For more direct comparison between simulations and experiments, we thus interpolated the locations of first two minima in the form factors to the entire studied range of CHOL concentrations in the bottom right panel of Fig.~\ref{fig:deviation}.
-These curves highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields.
-CHARMM36 seems to be off more than the other three force fields.
+Nevertheless, the minima in X-ray  form factors are independent of the simulation box size and correlate with membrane dimensions~\cite{NMRlipidsDatabank}.
+For a more direct comparison between simulations and experiments, we thus interpolated the locations of first two minima in the form factors over the entire studied range of CHOL concentrations (bottom right panel of Fig.~\ref{fig:densmaps}).
+Because differences between experiments and simulations for the 1\textsuperscript{st} minimum location are barely visible for some force fields, we have highlighted the relative deviations in Fig.~\ref{SIfig:firstminima}.
+These curves highlight that at first the addition of CHOL shifts the first minima to smaller wave vector values in the experiment; this is reasonably well captured by the simulation force fields, although CHARMM36 seems to be off more than the other three force fields.
 Above $\sim$25~mol-\% of CHOL, the location of the minimum shifts to larger wave vector values in the experiment, which is curiously not captured by any of the force fields.
-The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields.
-Slipids and Lipid17 are generally in better agreement with the experiment than MacRog and CHARMM36.
+For the second minimum, the experiment shows a steady shift to smaller wave vector values; this is reproduced by all simulation force fields, but Slipids and Lipid17 are generally in better agreement with the experiment than MacRog and CHARMM36.
 
@@
 
-Apart from the ordering effect on the bilayer structure, CHOL is also known to make them stiffer and less dynamic \cite{rog2009ordering,filippov2003effect,filippov2003influence}.
+Apart from the ordering effect on the bilayer structure, CHOL is also known to make bilayers stiffer and less dynamic \cite{rog2009ordering,filippov2003effect,filippov2003influence}.
 The comparison between lateral diffusion coefficients extracted from simulation and experiment has been limited due to a box-size dependence observed in simulations performed using periodic boundary conditions \cite{vogele2016divergent,vogele2018hydrodynamics,camley2015strong}.
 Here, we tackle this issue by performing simulations with three system sizes, which allows the extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent} \cite{vogele2016divergent,vogele2018hydrodynamics}.
-The size-dependence of lipid lateral diffusion from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
-The PBC-corrected lateral diffusion coefficients are shown in the top panel of Fig.~\ref{fig:dynamics} together with experimental values from $^1$H pulsed field gradient NMR diffusion measurements on label-free macroscopically aligned bilayers~\cite{filippov2003influence,filippov2003effect}.
+The size-dependence of phospholipid lateral diffusion from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
+The PBC-corrected lateral diffusion coefficients of POPC are shown in the top panel of Fig.~\ref{fig:dynamics} together with experimental values from $^1$H pulsed field gradient NMR diffusion measurements on label-free macroscopically aligned bilayers~\cite{filippov2003influence,filippov2003effect}.
 
@@
   \textbf{Dynamic properties of the POPC/CHOL mixtures.}
-  Top: Lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:sdlatpbc}.
-Experimental data are taken from NMR measurements on well-hydrated samples \cite{filippov2003influence,filippov2003effect}.
+  Top: POPC lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:sdlatpbc}.
+Experimental data are taken from NMR measurements of well-hydrated samples \cite{filippov2003influence,filippov2003effect}.
 The hollow circles show data extracted for Slipids using a shorter Lennard-Jones cutoff (see text).
+The error bars show the standard deviations of the diffusion coefficients obtained from 10,000 fits to the values $D\pm\Delta D$, where $\Delta D$ are sampled from normal distributions whose standard deviations are equal to the error estimates of the corresponding values of $D$.
+For these error estimates, we used the differences of the diffusion coefficients extracted from the two membrane leaflets.
   %
@@
 Lipid17 and CHARMM36 show too fast dynamics for pure POPC, and the slow-down induced by CHOL is exaggerated as compared to experiment.
-Diffusion in MacRog is generally too slow, yet Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms other force fields in terms of lateral dynamics.
-Interestingly, there is no correlation between dynamic and structural properties described earlier when it comes to the deviations from experiments.
-To investigate if the differences in lateral diffusion coefficients could be explained by different Lennard-Jones (LJ) cutoff values, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids using a shorter cutoff of 0.9~nm corresponding to that of Lipid17, while original cutoff for Slipids was 1.4~nm (see Table~\ref{SItab:params}).
-The PBC-corrected diffusion coefficient values with shorter cutoff, shown in the top panel of Fig.~\ref{fig:dynamics} as hollow red circles, are only slightly larger than the original values, indicating that differences arise from a combination of force field interaction parameters and are not explained by the LJ cutoff alone.
+Diffusion in MacRog is generally too slow, yet Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms the other force fields in terms of lateral dynamics.
+Interestingly, there is no correlation in the deviations from experiments between the lateral dynamics and the structural properties described earlier.
+To investigate if the differences in lateral diffusion coefficients could be explained by different Lennard-Jones (LJ) cutoff values, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids using a shorter cutoff of 0.9~nm (corresponding to that of Lipid17, while original cutoff for Slipids was 1.4~nm, see Table~\ref{SItab:params}).
+The PBC-corrected diffusion coefficient values with shorter cutoff, shown in the top panel of Fig.~\ref{fig:dynamics} as hollow red circles, are only slightly larger than the original values, indicating that differences arise from a combination of force field interaction parameters, and are not explained by the LJ cutoff alone.
 
@@
 This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration.
-For example, Fig.~\ref{SIfig:dvschol} shows that Lipid17 and CHARMM36 systemically underestimate the experimental values in systems with all sizes simulated here while the effect of CHOL seems to be qualitatively reproduced.
-However, the PBC corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations, leading to values close to experimental ones at 47~mol-\% CHOL.
-With MacRog, also the PBC corrected values fall below the experimental ones, yet the slowdown effect of CHOL again appears stronger than in experiment.
+For example, Fig.~\ref{SIfig:dvschol} would indicate that Lipid17 and CHARMM36 systemically underestimate the experimental values in systems with all sizes simulated here, while the effect of CHOL seems to be qualitatively reproduced.
+However, their PBC-corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations, leading to values close to experimental ones at 47~mol-\% CHOL.
+With MacRog, also the PBC-corrected values fall below the experimental ones, yet the slowdown effect of CHOL still appears stronger than in experiment.
 With Slipids, the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent.
-These results highlight that little can be said about CHOL-dependence of lateral diffusion coefficients without performing simulations with multiple box sizes required for the PBC correction at different CHOL concentrations.
-Therefore, fine-tuning of interaction parameters to correctly reproduce lateral diffusion coefficients can require massive number of simulations.
+These results underline that little can be said about the CHOL-dependence of lateral diffusion coefficients without the PBC correction that requires performing simulations with multiple box sizes at different CHOL concentrations.
+Consequently, fine-tuning of interaction parameters to correctly reproduce lateral diffusion coefficients can require massive number of simulations.
 
-We also provide the shear viscosity values of the membranes obtained from the fits of Eq.~\eqref{eq:sdlatpbc} in Fig.~\ref{fig:dynamics} for reference, yet they are not discussed further due to the large scatter of experimental estimates \cite{faizi2022vesicle}.
-Note that the data for CHARMM36 was also discussed in our previous work \cite{fabian2023protein}.
+We also provide the shear viscosity values of the membranes obtained from the fits of Eq.~\eqref{eq:sdlatpbc} in Fig.~\ref{fig:dynamics}.
+Direct comparison of these to experimental values is complicated due to the large scatter of experimental estimates \cite{faizi2022vesicle}, yet values from Slipids are expected to be  most realistic because its lateral diffusion coefficients are closest to experiments.
+Note that the viscosities from CHARMM36 simulations are further  discussed in our previous work \cite{fabian2023protein}.
 
@@
 
-To streamline the selection of force fields that best capture the effect of CHOL on different membrane properties, we defined the quantitative quality measures for simulations.
-For this we calculated  relative deviations from experiments (difference of simulated and experimental values divided by the experimental value) using interpolated data for the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients from all the force fields.
-The quality evaluations are shown in Fig.~\ref{fig:deviation} for force fields that predicted different effects of CHOL on membrane properties in previous sections.
+To streamline the selection of force fields that best capture the effect of CHOL on different membrane properties, we defined  quantitative quality measures for simulations.
+For this we calculated for all the force fields their relative deviations from experiments (difference between simulated and experimental values divided by the experimental value) using interpolated data for the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients.
+The quality evaluations are shown in Fig.~\ref{fig:deviation}.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{../FIGS/deviation.pdf}
+  \includegraphics[width=0.8\linewidth]{../FIGS/deviation.pdf}
   \caption{\label{fig:deviation}%
-  \textbf{Total relative deviation of the force fields from experimental data.} 
+  \textbf{Total relative deviation of force fields from experimental data.} 
 %
   Two leftmost columns:
-  The relative deviations from the first two minima in the form factor, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
+  The relative deviations in the first two form factor minima, palmitate and oleate chain order parameters, and diffusion coefficients shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
   %
   The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
-The differences in the form factor minima---shown in the bottom right panel---between experiment and simulation were calculated, normalized against experimental data, and summed together.
+For the form factor minima (shown in the bottom right panel) the deviation was obtained by calculating the difference between experiment and simulation for both minima, normalizing each against experimental data, and summing the two together.
 The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
@@
   The zoomed-in view to the deviation in the location of the first minimum in the form factor.
-%
-  Bottom right panel:
-  Effecf of CHOL on the location of the first two minima in the form factor.
-The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
   }
@@
 
-The quality evaluations reveal that Slipids provides overall the best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
+The quality evaluations reveal that Slipids provides the overall best agreement with experiment, yet its quality decreases slightly at higher CHOL concentrations.
 Lipid17 provides a slightly better agreement with experiment above $\sim$30~mol-\% of CHOL than Slipids, but exhibits major deviation from experiments at low CHOL concentrations.
 MacRog performs relatively well at low CHOL concentrations, but its quality deteriorates significantly upon the addition of CHOL.
-The deviation metric for CHARMM36 are significant at all CHOL concentrations.
-
-Since the contributions from the form factor minima deviations barely are visible for some force fields in the leftmost columns of Fig.~\ref{fig:deviation}, we have highlighted the error in the 1\textsuperscript{st} minimum location in the top right panel of Fig.~\ref{fig:deviation}.
-This panel demonstrates that apart from CHARMM36, all force fields capture the location of the 1\textsuperscript{st} minimum relatively well at low CHOL concentrations, but deviate significantly at larger CHOL concentrations.
+The deviation for CHARMM36 is significant at all CHOL concentrations.
 
@@
 
-The tested all-atom force fields captured the most important general effects of CHOL on membrane properties: increased acyl chain ordering, concomitant increase in bilayer thickness, and reduced lateral diffusion rate of lipids.
+The tested all-atom force fields captured the most important general effects of CHOL on membrane properties: increased acyl chain ordering, concomitant increase in bilayer thickness, and reduced lateral diffusion rate of phospholipids.
 However, quantitative comparison reveals differences between force fields and their qualities evaluated against NMR and X-ray scattering data.
-Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results up to 20~mol-\% of CHOL, but overestimate acyl chain ordering and membrane thickness after further addition of CHOL.
-An apparent exception to this is the Lipid17 force field, yet its seemingly better agreement with experiments can be explained by the over-correction of the originally underestimated acyl chain order at low CHOL concentrations upon CHOL addition.
+Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results until up to 20~mol-\% of CHOL, but overestimate the acyl chain ordering and membrane thickness after further addition of CHOL.
+An apparent exception to this is the Lipid17 force field, yet its seemingly better agreement with experiments results from compensation of intially (at low CHOL concentrations) underestimated acyl chain order by the overly strong response to the CHOL addition.
 In conclusion, a unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the CHOL ordering effect, particularly above 20~mol-\% of CHOL.
-Previously published comparison~\cite{ferreira13} suggests the same conclusion for historically relevant Berger/H{\"o}ltje force field parameters \cite{Berger97,Holtje01} that were not included in this work.
+A previously published comparison~\cite{ferreira13} suggested the same conclusion for the historically relevant Berger/H{\"o}ltje force field parameters \cite{Berger97,Holtje01} that were not included in this work.
 
-Besides the CHOL-induced ordering, effects on membrane properties has been discussed also in terms of the CHOL condensing effect which refers to a decrease in the \emph{area per phospholipid} upon the addition of CHOL (negative partial area) \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
+Besides the CHOL-induced ordering, effects on membrane properties have been discussed also in terms of (i) the CHOL condensing effect, which refers to a decrease in the \emph{area per phospholipid} upon the addition of CHOL (negative partial area) \cite{edholm2005areas}, or (ii) CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without effecting its total area (zero partial area) \cite{javanainen2017two}.
 At the physiological CHOL concentrations in the range from 0 to 30~mol-\%, only MacRog predicts negative partial area for CHOL, while CHARMM36 predicts zero partial area, and Slipids and Lipid17 predict small positive partial areas.
@@
 
-Considering also lateral dynamics, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid--CHOL interactions play major role.
+Considering also lateral dynamics and previous evaluation of rotational dynamics against NMR data~\cite{antila21using}, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid--CHOL interactions play major role.
 Nevertheless, all tested parameters capture qualitative effects of CHOL on membrane properties relatively well and differences between force fields are clearly smaller than, for example, in the case of PC--PS lipid mixtures~\cite{antila2022emerging}.
-Therefore, the quality of selected force field for other molecules, such as, other lipids, proteins, sugars, or drugs, together with force field compatibility might be more relevant decisive factor for simulations of complex systems.
-Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behavior, yet we believe that evaluating the magnitude of these effects is well beyond the scope of this work.
+Therefore, the quality of the force field selected for other molecules, such as lipids other than phosphatidylcholine, proteins, sugars, or drugs together with force field compatibility might be a more relevant decisive factor for simulations of complex systems.
+Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behavior, yet we believe that evaluating the magnitude of such effects is well beyond the scope of this work.
 
 Our results demonstrate that quality evaluation of lipid mixture simulations against experimental NMR and X-ray scattering data give consistent results for how accurately force field parameters capture intermolecular interactions.
-The interpolation approach introduced here extends the NMRlipids databank quality metrics~\cite{NMRlipidsDatabank} beyond individual systems enabling automatic ranking of not only lipid mixtures but also membrane mixed with other molecules such as ions.
-Such tools will strongly support the emerging endeavours for automatic improvements of force field parameters~\cite{antila2022emerging}.
+The interpolation approach introduced here extends the NMRlipids Databank quality metrics~\cite{NMRlipidsDatabank} beyond individual systems: this enables automatic ranking of not only lipid mixtures but also of membranes mixed with other molecules such as ions.
+Such tools will strongly support the emerging endeavours for automatic improvement of force field parameters~\cite{antila2022emerging}.
 
@@
 %
-Effect of simulation box size on the area per lipid across the studied cholesterol concentrations.
+Effect of simulation box size on the area per phospholipid across the studied cholesterol concentrations.
 %
@@
 %JJM thanks the Amber community (Benjamin Madej from the Walker lab and Callum Dickson from the Gould lab) for technical assistance related to the lipid14 force field parameters. 
-%
 MJ (grant no. 338160) and OHSO (grant nos. 315596 \& 319902) thank the Academy of Finland for funding.
@@
 
-J.J.M. contributed to early testing (debugging, setup, and running) of simulations with the Amber force field and manuscript editing.
+MJ performed all simulations.
+MJ and OHSO conceptualized the manuscript, analyzed all simulations and wrote the manuscript.
+PH performed and analyzed all SAXS experiments.
+GP designed SAXS experiments and revised the manuscript.
+JJM contributed to early testing (debugging, setup, and running) of simulations with the Amber force field and manuscript editing.
+MSM contributed to initial planning of the work and revised the manuscript.
 

### 2023-06-10 MANUSCRIPT/manuscriptSUPPL.tex
UPdate
Update manuscript

@@
-\documentclass[journal=jpcbfk]{achemso}
+\documentclass[journal=jpcbfk,manuscript=suppinfo]{achemso}
 
@@
 \author{Peter Heftberger}
-\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
+\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
 %
 \author{Matti Javanainen}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, 16000 Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
 \email{matti.javanainen@helsinki.fi}
@@
 \author{Jesper J. Madsen}
-\affiliation{Department of Global Health, College of Public Health}
-\alsoaffiliation{University of South Florida}
+\affiliation{Global and Planetary Health, College of Public Health}
+\alsoaffiliation{Department of Molecular Medicine, Morsani College of Medicine}
+\alsoaffiliation{University of South Florida, Tampa, Florida, 33612, United States of America}
 %
-\author{Josef Melcr}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-%
-\author{Markus Miettinen}
-\affiliation{Department of Chemistry, University of Bergen, Norway}
-\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
-
+\author{Markus S. Miettinen}
+\affiliation{Fachbereich Physik, Freie Universit\"at Berlin, 14195 Berlin, Germany}
+\alsoaffiliation{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
+\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
 %
 \author{O. H. Samuli Ollila}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, 16000 Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
 \email{samuli.ollila@helsinki.fi}
@@
 \author{Georg Pabst}
-\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
-\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
-%
+\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
+\alsoaffiliation{BioTechMed-Graz, 8010 Graz, Austria}
+\alsoaffiliation{Field of Excellence BioHealth—University of Graz, 8010 Graz, Austria}
 
@@
 
-\renewcommand{\thetable}{S\arabic{table}}%
-\renewcommand{\thefigure}{S\arabic{figure}}%
-\renewcommand{\thesection}{S\arabic{section}}%
-\renewcommand{\thepage}{S\arabic{page}}%
-
-\title{SUPPLEMENTARY INFORMATION FOR\\ 
-    NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Resolution Molecular Dynamics Simulations}
+\title{Quantitative Comparison Against Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions}
 
@@
 
-The force field specific simulation parameters used for each force field are listed in Table~\ref{SItab:params}.
+The force-field-specific simulation parameters used for each force field are listed in Table~\ref{SItab:params}.
 The parameter input files (\texttt{.mdp}) are provided in the Zenodo portal (links in the main text).
@@
 The P-LINCS constraint algorithm \cite{hess97,hess07} was used for the bonds noted in Table~\ref{SItab:params}.
-NMRlipids databank~\cite{NMRlipidsDatabank} ID numbers of simulations are given in table~\ref{IDtable}.
+NMRlipids Databank~\cite{NMRlipidsDatabank} ID numbers of simulations are given in Table~\ref{IDtable}.
 
@@
 \begin{table}[]
+\begin{small}
 \begin{tabular}{ccc}
@@
 \midrule
+64POPC\_3200SOL\_298K  &  678 \\ 
+256POPC\_12800SOL\_298K  &  710 \\ 
+1024POPC\_51200SOL\_298K  &  701 \\ 
+\midrule
+64POPC\_8CHOL\_3600SOL\_298K  &  109 \\ 
+256POPC\_32CHOL\_14400SOL\_298K  &  119 \\ 
 1024POPC\_128CHOL\_57600SOL\_298K  &  426 \\ 
-64POPC\_3200SOL\_298K  &  546 \\ 
+\midrule
+64POPC\_16CHOL\_4000SOL\_298K  &  525 \\ 
+256POPC\_64CHOL\_16000SOL\_298K  &  72 \\ 
+1024POPC\_256CHOL\_64000SOL\_298K  &  88 \\ 
+\midrule
+64POPC\_26CHOL\_4500SOL\_298K  &  393 \\ 
+256POPC\_104CHOL\_18000SOL\_298K  &  298 \\ 
+1024POPC\_416CHOL\_72000SOL\_298K  &  412 \\ 
+\midrule
+64POPC\_40CHOL\_5200SOL\_298K  &  620 \\ 
+256POPC\_160CHOL\_20800SOL\_298K  &  275 \\ 
+1024POPC\_640CHOL\_83200SOL\_298K  &  550 \\ 
+\midrule
+64POPC\_56CHOL\_6000SOL\_298K  &  91 \\ 
+256POPC\_224CHOL\_24000SOL\_298K  &  166 \\ 
 1024POPC\_896CHOL\_96000SOL\_298K  &  543 \\ 
-64POPC\_40CHOL\_5200SOL\_298K  &  620 \\ 
-64POPC\_56CHOL\_6000SOL\_298K  &  91 \\ 
-1024POPC\_416CHOL\_72000SOL\_298K  &  412 \\ 
-1024POPC\_256CHOL\_64000SOL\_298K  &  88 \\ 
-256POPC\_64CHOL\_16000SOL\_298K  &  72 \\ 
-64POPC\_8CHOL\_3600SOL\_298K  &  109 \\ 
-256POPC\_104CHOL\_18000SOL\_298K  &  298 \\ 
-64POPC\_16CHOL\_4000SOL\_298K  &  525 \\ 
-64POPC\_26CHOL\_4500SOL\_298K  &  393 \\ 
-1024POPC\_640CHOL\_83200SOL\_298K  &  550 \\ 
-256POPC\_160CHOL\_20800SOL\_298K  &  275 \\ 
-256POPC\_224CHOL\_24000SOL\_298K  &  166 \\ 
-256POPC\_32CHOL\_14400SOL\_298K  &  119 \\ 
+\bottomrule
 \end{tabular}
@@
 \midrule
-Slipids\_POPC\_CHOL20\_L  &  691 \\ 
-Slipids\_POPC\_CHOL47\_M  &  712 \\ 
-Slipids\_POPC\_CHOL29\_S  &  693 \\ 
+Slipids\_POPC\_S  &  664 \\ 
+Slipids\_POPC\_M  &  708 \\ 
+Slipids\_POPC\_L  &  696 \\ 
+\midrule
+Slipids\_POPC\_CHOL11\_S  &  672 \\ 
+Slipids\_POPC\_CHOL11\_M  &  681 \\ 
+Slipids\_POPC\_CHOL11\_L  &  697 \\ 
+\midrule
 Slipids\_POPC\_CHOL20\_S  &  692 \\ 
 Slipids\_POPC\_CHOL20\_M  &  687 \\ 
-Slipids\_POPC\_CHOL11\_S  &  672 \\ 
-Slipids\_POPC\_CHOL11\_M  &  681 \\ 
-Slipids\_POPC\_L  &  696 \\ 
+Slipids\_POPC\_CHOL20\_L  &  691 \\ 
+\midrule
+Slipids\_POPC\_CHOL29\_S  &  693 \\ 
+Slipids\_POPC\_CHOL29\_M  &  659 \\ 
+Slipids\_POPC\_CHOL29\_L  &  709 \\ 
+\midrule
 Slipids\_POPC\_CHOL38\_S  &  661 \\ 
-Slipids\_POPC\_M  &  708 \\ 
+Slipids\_POPC\_CHOL38\_M  &  668 \\ 
 Slipids\_POPC\_CHOL38\_L  &  670 \\ 
-Slipids\_POPC\_S  &  664 \\ 
-Slipids\_POPC\_CHOL29\_L  &  709 \\ 
-Slipids\_POPC\_CHOL29\_M  &  659 \\ 
-Slipids\_POPC\_CHOL11\_L  &  697 \\ 
-Slipids\_POPC\_CHOL38\_M  &  668 \\ 
+\midrule
 Slipids\_POPC\_CHOL47\_S  &  703 \\ 
+Slipids\_POPC\_CHOL47\_M  &  712 \\ 
 Slipids\_POPC\_CHOL47\_L  &  682 \\
+\bottomrule
 \end{tabular}
@@
 \midrule
+Lipid17\_POPC\_S  &  715 \\ 
+Lipid17\_POPC\_M  &  657 \\ 
+Lipid17\_POPC\_L  &  684 \\ 
+\midrule
+Lipid17\_POPC\_CHOL11\_S  &  680 \\ 
+Lipid17\_POPC\_CHOL11\_M  &  689 \\
+Lipid17\_POPC\_CHOL11\_L  &  663 \\ 
+\midrule
+Lipid17\_POPC\_CHOL20\_S  &  662 \\ 
+Lipid17\_POPC\_CHOL20\_M  &  666 \\ 
+Lipid17\_POPC\_CHOL20\_L  &  667 \\ 
+\midrule
+Lipid17\_POPC\_CHOL29\_S  &  686 \\ 
 Lipid17\_POPC\_CHOL29\_M  &  700 \\ 
+Lipid17\_POPC\_CHOL29\_L  &  683 \\ 
+\midrule
+Lipid17\_POPC\_CHOL38\_S  &  688 \\ 
+Lipid17\_POPC\_CHOL38\_M  &  694 \\ 
+Lipid17\_POPC\_CHOL38\_L  &  673 \\ 
+\midrule
+Lipid17\_POPC\_CHOL47\_S  &  707 \\ 
+Lipid17\_POPC\_CHOL47\_M  &  714 \\ 
 Lipid17\_POPC\_CHOL47\_L  &  717 \\ 
-Lipid17\_POPC\_CHOL11\_L  &  663 \\ 
-Lipid17\_POPC\_CHOL29\_L  &  683 \\ 
-Lipid17\_POPC\_CHOL20\_S  &  662 \\ 
-Lipid17\_POPC\_CHOL47\_M  &  714 \\ 
-Lipid17\_POPC\_CHOL38\_S  &  688 \\ 
-Lipid17\_POPC\_L  &  684 \\ 
-Lipid17\_POPC\_S  &  715 \\ 
-Lipid17\_POPC\_CHOL20\_M  &  666 \\ 
-Lipid17\_POPC\_M  &  657 \\ 
-Lipid17\_POPC\_CHOL29\_S  &  686 \\ 
-Lipid17\_POPC\_CHOL38\_L  &  673 \\ 
-Lipid17\_POPC\_CHOL47\_S  &  707 \\ 
-Lipid17\_POPC\_CHOL11\_S  &  680 \\ 
-Lipid17\_POPC\_CHOL20\_L  &  667 \\ 
-Lipid17\_POPC\_CHOL38\_M  &  694 \\ 
-Lipid17\_POPC\_CHOL11\_M  &  689 \\
+\bottomrule
 \end{tabular}
@@
 \midrule
+MacRog\_POPC\_S  &  674 \\ 
+MacRog\_POPC\_M  &  675 \\ 
+MacRog\_POPC\_L  &  658 \\ 
+\midrule
+MacRog\_POPC\_CHOL11\_S  &  665 \\ 
+MacRog\_POPC\_CHOL11\_M  &  695 \\ 
+MacRog\_POPC\_CHOL11\_L  &  660 \\ 
+\midrule
+MacRog\_POPC\_CHOL20\_S  &  702 \\
+MacRog\_POPC\_CHOL20\_M  &  706 \\ 
+MacRog\_POPC\_CHOL20\_L  &  677 \\ 
+\midrule
+MacRog\_POPC\_CHOL29\_S  &  690 \\ 
+MacRog\_POPC\_CHOL29\_M  &  655 \\ 
+MacRog\_POPC\_CHOL29\_L  &  705 \\ 
+\midrule
 MacRog\_POPC\_CHOL38\_S  &  716 \\ 
-MacRog\_POPC\_CHOL29\_L  &  705 \\ 
+MacRog\_POPC\_CHOL38\_M  &  679 \\ 
+MacRog\_POPC\_CHOL38\_L  &  698 \\ 
+\midrule
+MacRog\_POPC\_CHOL47\_S  &  704 \\ 
 MacRog\_POPC\_CHOL47\_M  &  699 \\ 
 MacRog\_POPC\_CHOL47\_L  &  671 \\ 
-MacRog\_POPC\_CHOL20\_L  &  677 \\ 
-MacRog\_POPC\_S  &  674 \\ 
-MacRog\_POPC\_CHOL20\_M  &  706 \\ 
-MacRog\_POPC\_CHOL38\_L  &  698 \\ 
-MacRog\_POPC\_CHOL38\_M  &  679 \\ 
-MacRog\_POPC\_M  &  675 \\ 
-MacRog\_POPC\_L  &  658 \\ 
-MacRog\_POPC\_CHOL29\_M  &  655 \\ 
-MacRog\_POPC\_CHOL11\_L  &  660 \\ 
-MacRog\_POPC\_CHOL11\_S  &  665 \\ 
-MacRog\_POPC\_CHOL29\_S  &  690 \\ 
-MacRog\_POPC\_CHOL11\_M  &  695 \\ 
-MacRog\_POPC\_CHOL47\_S  &  704 \\ 
-MacRog\_POPC\_CHOL20\_S  &  702 \\
+\bottomrule
 \end{tabular}
-\caption{NMRlipids databank IDs of performed simulations.}\label{IDtable}
+\caption{NMRlipids Databank IDs of performed simulations.}\label{IDtable}
+\end{small}
 \end{table}
@@
     \includegraphics[width=0.45\linewidth]{../FIGS/scatt25.pdf}
-    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the next figure.}
+    \caption{\textbf{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol.} More data are shown in the next figure.}
     \label{SIfig:scattering1}
@@
     \includegraphics[width=0.45\linewidth]{../FIGS/scatt50.pdf}
-    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the previous figure.}
+    \caption{\textbf{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol.} More data are shown in the previous figure.}
     \label{SIfig:scattering2}
@@
     \caption{\label{SIfig:scattering}%
-     \textbf{Scattering intensities as a function of scattering vector from experiments and simulations.} Each of the profiles is shifted vertically with respect to the previous one, by 1 for the experimental profiles and by 50 for the computational ones.
+     \textbf{Absolute values of X-ray scattering form factors.} Each of the profiles is shifted vertically with respect to the previous one, by 1 for the experimental profiles and by 50 for the computational ones.
 The minima are marked by filled circles to guide the eye.
+The numerical data for these plots can be found from the entries in the NMRlipids databank with the ID numbers listed in Table~\ref{IDtable}.
     }
@@
     Each of the profiles is shifted vertically with respect to the previous one by 100 units.
+The numerical data for these plots can be found from the entries in the NMRlipids databank with the ID numbers listed in Table~\ref{IDtable}.
+    }
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=8.7cm]{../FIGS/firstminima.pdf}
+    \caption{\label{SIfig:firstminima}%
+     \textbf{Deviation of the first form factor minima.} 
     }
@@
      The legend at the top corresponds to all simulations, and the one at the bottom to the experiments.
+Error bars in simulations are smaller than symbols.
+Error of the experimental data is estimated to be $\pm$0.02~\cite{Ollila16}.
+The numerical data for these plots can be found from the entries in the NMRlipids databank with the ID numbers listed in Table~\ref{IDtable}.
     }
@@
 Since the order parameters measured for the two hydrogens bound to the C2 carbon differ, they are both shown in the plots.
+Stereospecific labeling is not done for these hydrogens but the one with larger value is shown first.
+Error bars in simulations are smaller than symbols.
+Error of the experimental data is estimated to be $\pm$0.02~\cite{Ollila16}.
+The numerical data for these plots can be found from the entries in the NMRlipids databank with the ID numbers listed in Table~\ref{IDtable}.
     }
@@
     \caption{\label{SIfig:dvssize}%
-    \textbf{Dependence of lateral diffusion coefficients on simulation box size.} The values calculated for the lipid centres of mass with \texttt{gmx msd} after eliminating leaflet drift.
+    \textbf{Dependence of POPC lateral diffusion coefficients on the simulation box size.} The values calculated for the lipid centres of mass with \texttt{gmx msd} after eliminating leaflet drift.
 The values for the three system sizes are shown as markers together with fits of Eq.~\eqref{eq:sdlatpbc}.
@@
     \caption{\label{SIfig:dvschol}%
-     \textbf{Dependence of lateral diffusion coefficients on cholesterol concentration.} Data are shown for all system sizes; small (S), medium (M), and large (L).
+     \textbf{Dependence of POPC lateral diffusion coefficients on cholesterol concentration.} Data are shown for all system sizes; small (S), medium (M), and large (L).
 The values extrapolated to infinity are shown as well ($\infty$).
-Experimental measured at two hydration levels, 30 m-\% and 55-\% of water \cite{filippov2003effect,filippov2003influence}.
+Experimental data were measured at a hydration level of 55~wt-\% of water \cite{filippov2003effect,filippov2003influence}, whereas simulations have a hydration level of~$\sim$54~wt-\%.
     }
@@
     \caption{\label{SIfig:aplvssize}%
-     \textbf{Dependence of area per lipid on simulation box size.} Area per lipid is calculated by dividing the membrane area by the number of lipids in one leaflet.
+     \textbf{Dependence of area per lipid on simulation size.} Area per lipid is calculated by dividing the box area by the number of lipids in one leaflet.
 Error bars show standard error extracted using block averaging in \texttt{gmx analyze}.

### 2023-06-15 MANUSCRIPT/manuscript.tex
Update for the first submission

@@
 
-\author{Peter Heftberger}
-\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
-%
 \author{Matti Javanainen}
@@
 \email{matti.javanainen@helsinki.fi}
+\author{Peter Heftberger}
+\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
 %
 \author{Jesper J. Madsen}
-\affiliation{Global and Planetary Health, College of Public Health}
-\alsoaffiliation{Department of Molecular Medicine, Morsani College of Medicine}
-\alsoaffiliation{University of South Florida, Tampa, Florida, 33612, United States of America}
+\affiliation{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\alsoaffiliation{Center for Global Health and Infectious Diseases Research, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\alsoaffiliation{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
 %
@@
 %
-\author{O. H. Samuli Ollila}
-\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, 16000 Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
-\email{samuli.ollila@helsinki.fi}
 %
@@
 %
+\author{O. H. Samuli Ollila}
+\affiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
+\alsoaffiliation{VTT Technical Research Centre of Finland, Espoo, Finland}
+\email{samuli.ollila@helsinki.fi}
 
@@
 %
-Here, we define quantitative quality measures for simulations of binary lipid mixtures in membranes against C--H bond order parameters and lateral diffusion coefficients from NMR spectroscopy as well as form factors from X-ray scattering.
+Here, we define quantitative quality measures for simulations of binary lipid mixtures in membranes against the C--H bond order parameters and lateral diffusion coefficients from NMR spectroscopy as well as the form factors from X-ray scattering.
 Based on these measures, we perform a systematic evaluation on the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol.
@@
 This approach has been useful for finding the best force fields to describe the headgroups of phosphatidylcholine (PC)~\cite{botan15,kurki2022structure}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE)~\cite{bacle2021inverse} and phosphatidylglycerol (PG)~lipids \cite{bacle2021inverse}, to evaluate and improve membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and to find simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}.
-Furthermore, quantitative quality measures based on C--H bond order parameters have been recently defined and used to rank simulations in the NMRlipids Databank~\cite{NMRlipidsDatabank}.
+Furthermore, quantitative quality measures based on the C--H bond order parameters have been recently defined and used to rank simulations in the NMRlipids Databank~\cite{NMRlipidsDatabank}.
 However, such automatic quality evaluation is limited to simulations for which experimental data at a corresponding composition and temperature are available.
@@
 Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high}.
-This avoids the precipitation of CHOL crystallites at high concentration \cite{buboltz1999novel}, yielding non-phase separated samples up to 50~mol-\% CHOL content.
+This avoids the precipitation of CHOL crystallites at high concentration \cite{buboltz1999novel}, yielding non-phase-separated samples up to 50~mol-\% CHOL content.
 Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification.
@@
 We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%.
-Systems were simulated using four commonly used force fields: CHARMM36 (often abbreviated ``C36'' in figure legends in this work) \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and all-atom OPLS compatible MacRog \cite{kulig14,kulig15,Kulig16}.
+Systems were simulated using four commonly used force fields: CHARMM36 (often abbreviated ``C36'' in figure legends in this work) \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and all-atom-OPLS-compatible MacRog \cite{kulig14,kulig15,Kulig16}.
 To eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256, or 1024 POPC molecules).
@@
     Details of the (small/medium/large) simulation systems.
-The box dimensions in the membrane plane ($x$ and $y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
+The box dimensions in the membrane plane ($x$ and $y$) and normal to the membrane ($z$) are provided for the Slipids simulations; the values vary slightly between the force fields.
     }
@@
 This resulted in 1D deviation vectors as a function of CHOL concentration.
-For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
+For diffusion coefficients and form factor minima (strickly speaking: minima of the absolute value of the form factor), the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration.
 All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of CHOL concentration.
@@
 
-\subsection{Cholesterol Effect on Membrane Properties is Manifested Differently in Different Force Fields}
+\subsection{Cholesterol Effect on Membrane Properties Manifests Differently in Different Force Fields}
 
-CHOL-induced ordering straightens the acyl chain conformations, which leads to membrane thickening.
+CHOL-induced ordering straightens the acyl chains, which leads to membrane thickening.
 While acyl chain order and membrane thickness are well correlated~\cite{NMRlipidsDatabank}, lipid bilayer dimensions can be accessed more directly by measuring the X-ray form factor, which is related to the electron density along membrane normal \textit{via} Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16}.
@@
 For more detailed comparison of membrane structure, we interpolated the changes in electron density profiles along membrane normal (Fig.~\ref{SIfig:densprofs}) as a function of CHOL concentration to create two-dimensional electron density maps (Fig.~\ref{fig:densmaps}).
-Overall, all electron density profiles share the same features across all CHOL concentrations: High-densities band corresponding to the tightly-packed interfacial regions containing the electron-rich phosphorus, low density at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
+Overall, all electron density profiles share the same features across all CHOL concentrations: High densities corresponding to the tightly-packed interfacial regions containing the electron-rich phosphorus, low density at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase.
 However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields and the SPD model.
@@
 The less sharp bands for Lipid17 and especially Slipids profiles indicate that their membranes are more flexible.
-The SDP model gives the sharpest bands (bottom left panel in Fig.~\ref{fig:densmaps} and original electron density profiles in Fig.~\ref{SIfig:densprofs}), suggesting more uniform membrane structure than any MD simulation.
-However, system size plays a role here as  undulations smear out the density profile features with increasing simulation box size as demonstrated in Fig.~\ref{SIfig:densprofssize} by the density maps calculated for the CHARMM36 simulations in three sizes.
+The SDP model gives the sharpest bands (top right panel in Fig.~\ref{fig:densmaps} and original electron density profiles in Fig.~\ref{SIfig:densprofs}), suggesting more uniform membrane structure than any MD simulation.
+However, system size plays a role here as  undulations smear out the density profile features with increasing simulation box size, as demonstrated in Fig.~\ref{SIfig:densprofssize} by the density maps calculated for the CHARMM36 simulations in three sizes.
 Still, even in the smallest system, the band intensities are less localized than in the SDP model.
-This might indicate different elastic properties between simulation and experiment, but we cannot fully exclude the role of models used to interpret scattering data.
+This might indicate different elastic properties between simulation and experiment, but we cannot fully exclude the role of models used to interpret the scattering data.
 
 Nevertheless, the minima in X-ray  form factors are independent of the simulation box size and correlate with membrane dimensions~\cite{NMRlipidsDatabank}.
-For a more direct comparison between simulations and experiments, we thus interpolated the locations of first two minima in the form factors over the entire studied range of CHOL concentrations (bottom right panel of Fig.~\ref{fig:densmaps}).
+For a more direct comparison between simulations and experiments, we thus interpolated the locations of the first two minima in the form factors over the entire studied range of CHOL concentrations (bottom right panel of Fig.~\ref{fig:densmaps}).
 Because differences between experiments and simulations for the 1\textsuperscript{st} minimum location are barely visible for some force fields, we have highlighted the relative deviations in Fig.~\ref{SIfig:firstminima}.
@@
 
-\subsection{The Force Fields Predict Very Different Lateral Mobilities}
+\subsection{Force Fields Predict Very Different Lateral Mobilities}
 
 Apart from the ordering effect on the bilayer structure, CHOL is also known to make bilayers stiffer and less dynamic \cite{rog2009ordering,filippov2003effect,filippov2003influence}.
-The comparison between lateral diffusion coefficients extracted from simulation and experiment has been limited due to a box-size dependence observed in simulations performed using periodic boundary conditions \cite{vogele2016divergent,vogele2018hydrodynamics,camley2015strong}.
-Here, we tackle this issue by performing simulations with three system sizes, which allows the extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent} \cite{vogele2016divergent,vogele2018hydrodynamics}.
-The size-dependence of phospholipid lateral diffusion from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependence of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
+The comparison between lateral diffusion coefficients extracted from simulation and experiment has been limited due to a box-size dependence observed in simulations performed using periodic boundary conditions (PBCs)\cite{vogele2016divergent,vogele2018hydrodynamics,camley2015strong}.
+Here, we tackle this issue by performing simulations with three system sizes, which allows extrapolation of the results to an infinite system with the theoretical description developed by \citeauthor{vogele2016divergent} \cite{vogele2016divergent,vogele2018hydrodynamics}.
+The size-dependence of phospholipid lateral diffusion from simulations together with the fit of Eq.~\eqref{eq:sdlatpbc} are shown in Fig.~\ref{SIfig:dvssize}, whereas the CHOL-dependences of these values in systems with different sizes are shown in Fig.~\ref{SIfig:dvschol}.
 The PBC-corrected lateral diffusion coefficients of POPC are shown in the top panel of Fig.~\ref{fig:dynamics} together with experimental values from $^1$H pulsed field gradient NMR diffusion measurements on label-free macroscopically aligned bilayers~\cite{filippov2003influence,filippov2003effect}.
@@
   Bottom: Shear viscosities obtained from the finite-size correction, Eq.~\eqref{eq:sdlatpbc}.
+The distributions of the viscosity values from the 10,000 fits (see above) were often skewed, and thus could not be fitted by a single Gaussian, complicating the error estimation.
+The results from double-Gaussian fits to the data are presented in Fig.~\ref{SIfig:viscoerrors}.
   %
@@
 
-The lipid force fields again show significantly different behavior.
+The lipid force fields again show significantly different behaviors.
 Lipid17 and CHARMM36 show too fast dynamics for pure POPC, and the slow-down induced by CHOL is exaggerated as compared to experiment.
-Diffusion in MacRog is generally too slow, yet Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms the other force fields in terms of lateral dynamics.
+Diffusion in MacRog is generally too slow.
+Yet Slipids provides an essentially quantitative agreement with experiment across the studied CHOL concentrations, and thus significantly outperforms the other force fields in terms of lateral dynamics.
 Interestingly, there is no correlation in the deviations from experiments between the lateral dynamics and the structural properties described earlier.
@@
 This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration.
-For example, Fig.~\ref{SIfig:dvschol} would indicate that Lipid17 and CHARMM36 systemically underestimate the experimental values in systems with all sizes simulated here, while the effect of CHOL seems to be qualitatively reproduced.
+For example, Fig.~\ref{SIfig:dvschol} would indicate that Lipid17 and CHARMM36 systematically underestimate the experimental values in systems with all sizes simulated here, while the effect of CHOL seems to be qualitatively reproduced.
 However, their PBC-corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations, leading to values close to experimental ones at 47~mol-\% CHOL.
@@
 We also provide the shear viscosity values of the membranes obtained from the fits of Eq.~\eqref{eq:sdlatpbc} in Fig.~\ref{fig:dynamics}.
-Direct comparison of these to experimental values is complicated due to the large scatter of experimental estimates \cite{faizi2022vesicle}, yet values from Slipids are expected to be  most realistic because its lateral diffusion coefficients are closest to experiments.
+Direct comparison of these to experimental values is complicated due to the large scatter of experimental estimates \cite{faizi2022vesicle}, yet values from Slipids are expected to be  most realistic, because its lateral diffusion coefficients are closest to experiments.
 Note that the viscosities from CHARMM36 simulations are further  discussed in our previous work \cite{fabian2023protein}.
@@
   The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data.
-For the form factor minima (shown in the bottom right panel) the deviation was obtained by calculating the difference between experiment and simulation for both minima, normalizing each against experimental data, and summing the two together.
+For the form factor minima (shown bottom right in Fig.~\ref{fig:densmaps}) the deviation was obtained by calculating the difference between experiment and simulation for both minima, normalizing each against experimental data, and summing the two together.
 The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
 %
-  Top right panel: 
-  The zoomed-in view to the deviation in the location of the first minimum in the form factor.
+%  Top right panel: 
+%  The zoomed-in view to the deviation in the location of the first minimum in the form factor.
   }
@@
 However, quantitative comparison reveals differences between force fields and their qualities evaluated against NMR and X-ray scattering data.
-Comparison with NMR order parameters and X-ray scattering form factors propose that simulations reproduce experimental results until up to 20~mol-\% of CHOL, but overestimate the acyl chain ordering and membrane thickness after further addition of CHOL.
-An apparent exception to this is the Lipid17 force field, yet its seemingly better agreement with experiments results from compensation of intially (at low CHOL concentrations) underestimated acyl chain order by the overly strong response to the CHOL addition.
-In conclusion, a unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the CHOL ordering effect, particularly above 20~mol-\% of CHOL.
+Comparison with NMR order parameters and X-ray scattering form factors proposes that simulations reproduce experimental results until up to 20~mol-\% of CHOL, but overestimate the acyl chain ordering and membrane thickness after further addition of CHOL.
+An apparent exception to this is the Lipid17 force field, yet its seemingly better agreement with experiments results from compensation of the initially (at low CHOL concentrations) underestimated acyl chain order by the overly strong response to CHOL addition.
+In conclusion, a unified picture emerging from comparison with NMR and X-ray scattering data suggests that all the tested force fields overestimate the CHOL-induced ordering effect, particularly above 20~mol-\% of CHOL.
 A previously published comparison~\cite{ferreira13} suggested the same conclusion for the historically relevant Berger/H{\"o}ltje force field parameters \cite{Berger97,Holtje01} that were not included in this work.
@@
 
-Considering also lateral dynamics and previous evaluation of rotational dynamics against NMR data~\cite{antila21using}, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid--CHOL interactions play major role.
+Considering also lateral dynamics and previous evaluation of rotational dynamics against NMR data~\cite{antila21using}, Slipids is overall closest to experiments among the parameters tested here, and is therefore probably the best choice for studies where lipid--CHOL interactions play a major role.
 Nevertheless, all tested parameters capture qualitative effects of CHOL on membrane properties relatively well and differences between force fields are clearly smaller than, for example, in the case of PC--PS lipid mixtures~\cite{antila2022emerging}.
-Therefore, the quality of the force field selected for other molecules, such as lipids other than phosphatidylcholine, proteins, sugars, or drugs together with force field compatibility might be a more relevant decisive factor for simulations of complex systems.
-Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behavior, yet we believe that evaluating the magnitude of such effects is well beyond the scope of this work.
+Therefore, the quality of the force field selected for other molecules---such as proteins, sugars, drugs, or lipids other than phosphatidylcholine---and the force field compatibility might be more relevant decisive factors for simulations of complex systems.
+Finally, it must be noted that while all simulations were performed with their suggested simulation parameters, the different simulation engines might provide slightly different behaviour, yet we believe that evaluating the magnitude of such effects is well beyond the scope of this work.
 
 Our results demonstrate that quality evaluation of lipid mixture simulations against experimental NMR and X-ray scattering data give consistent results for how accurately force field parameters capture intermolecular interactions.
-The interpolation approach introduced here extends the NMRlipids Databank quality metrics~\cite{NMRlipidsDatabank} beyond individual systems: this enables automatic ranking of not only lipid mixtures but also of membranes mixed with other molecules such as ions.
+The interpolation approach introduced here extends the NMRlipids Databank quality metrics~\cite{NMRlipidsDatabank} beyond individual systems: This enables automatic ranking of not only lipid mixtures but also of membranes mixed with other molecules such as ions.
 Such tools will strongly support the emerging endeavours for automatic improvement of force field parameters~\cite{antila2022emerging}.
@@
 MJ (grant no. 338160) and OHSO (grant nos. 315596 \& 319902) thank the Academy of Finland for funding.
+MSM acknowledges support from the Volkswagen Foundation (86110) and the Trond Mohn Foundation (BFS2017TMT01).
+GP acknowledges funding through the Austrian Science Fund FWF (grant. no. P 24459).
 
@@
 PH performed and analyzed all SAXS experiments.
-GP designed SAXS experiments and revised the manuscript.
 JJM contributed to early testing (debugging, setup, and running) of simulations with the Amber force field and manuscript editing.
 MSM contributed to initial planning of the work and revised the manuscript.
+GP designed SAXS experiments and revised the manuscript.
 

### 2023-06-15 MANUSCRIPT/manuscriptSUPPL.tex

@@
 \email{matti.javanainen@helsinki.fi}
+%
+\author{Peter Heftberger}
+\affiliation{Biophysics, Institute of Molecular Biosciences, NAWI Graz, University of Graz, 8010 Graz, Austria}
 %
@@
 %
-\author{O. H. Samuli Ollila}
-\affiliation{Institute of Organic Chemistry and Biochemistry, Academy of Sciences of the Czech Republic, 16000 Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
-\email{samuli.ollila@helsinki.fi}
-%
 \author{Georg Pabst}
@@
 \alsoaffiliation{Field of Excellence BioHealth—University of Graz, 8010 Graz, Austria}
+%
+\author{O. H. Samuli Ollila}
+\affiliation{Institute of Biotechnology, University of Helsinki, 00790 Helsinki, Finland}
+\alsoaffiliation{VTT Technical Research Centre of Finland, Espoo, Finland}
+\email{samuli.ollila@helsinki.fi}
+%
 
@@
     \includegraphics[width=0.45\linewidth]{../FIGS/scatt25.pdf}
-    \caption{\textbf{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol.} More data are shown in the next figure.}
+    \caption{\textbf{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol.} Solid lines correspond to best fits using the SDP-GAP model.
+More data are shown in the next figure.}
     \label{SIfig:scattering1}
@@
     \includegraphics[width=0.45\linewidth]{../FIGS/scatt50.pdf}
-    \caption{\textbf{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol.} More data are shown in the previous figure.}
+    \caption{\textbf{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol.} Solid lines correspond to best fits using the SDP-GAP model.
+More data are shown in the previous figure.}
     \label{SIfig:scattering2}
@@
 
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.8\linewidth]{../FIGS/viscoerrors.pdf}
+    \caption{\label{SIfig:viscoerrors}%
+    \textbf{Error estimation for viscosities.}
+    The error in $D_\infty$ was estimated by performing 10,000 fits of Eq.~\eqref{eq:sdlatpbc} with each $D$ value assigned a random error (see the caption of Fig.~\ref{fig:dynamics} in the main text).
+These fits revealed a skewed Gaussian distribution of the viscosity values, which could be fitted by two Gaussians.
+The mean values and the corresponding standard deviations are shown here for all simulations.
+For some simulations, the distribution is close to unimodal and the two markers overlap, whereas for some they deviate significantly.
+The data for different force fields are slightly shifted horizonatally for readability.
+    }
+\end{figure}
+
 \clearpage
